---
title: "Basal_PSM_tests"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(MatchIt)
library(dplyr)
library(ggplot2)
library(UBL)
```

### 1. Load Basal cells seurat obj 

```{r}
obj <- readRDS(here("00_data_ingest", "04_rds_generated", "seurat_obj", "CD4_processed.rds"))
obj@meta.data$old_cluster <- obj@meta.data$res.0.6


obj@meta.data$Stage <- as.character(obj@meta.data$Stage)
obj@meta.data[obj@meta.data$Batch == 3 & obj@meta.data$Stage == "", "Stage"] <- "Unknown"
obj@meta.data[obj@meta.data$Stage == "", "Stage"] <- "Benign"
```


### 2. calculate Propensity Score Matching
First of all, find out whether other variables have huge difference based on stage,

#### Step1. construct a counts table based on different variables
```{r}
meta = as.data.frame(obj@meta.data %>% select(PatientID, Stage, Disease) %>% group_by(PatientID))

meta$Stage <- as.data.frame(meta %>% select(Stage, PatientID) %>% group_by(PatientID) %>% add_tally())$n
# meta$Sex <- as.data.frame(meta %>% select(Stage, Sex) %>% group_by(Stage) %>% add_tally())$n
meta$Disease <- as.data.frame(meta %>% select(PatientID, Disease) %>% group_by(PatientID) %>% add_tally())$n

meta$PatientID <- as.integer(gsub("[^\\d]", "", meta$PatientID, perl = T))
# 
# 
# meta$Stage <- sapply(meta$Stage, function(i) {
#     if(i == "Benign") {
#         return(0)
#     } else if (i == "I") {
#         return(1)
#     } else if(i == "II") {
#         return(2)
#     } else if(i == "III") {
#         return(3)
#     } else if(i == "IV") {
#         return(4)
#     }
#     return(5)
# })
# 
# 
# meta$Disease <- sapply(meta$Disease, function(i) {
#     if(i == "Normal") {
#         return(0)
#     } else if (i == "Adenocarcinoma") {
#         return(1)
#     } else if(i == "Squamous cell carcinoma") {
#         return(2)
#     } else if(i == "Large cell neuroendocrine carcinoma") {
#         return(3)
#     } else if(i == "Lymphoid epithelioid carcinoma") {
#         return(4)
#     } else if(i == "") {
#         return(6)
#     }
#     return(5)
# })
# 
# 
# meta$Sex <- sapply(meta$Sex, function(i) {
#     if(i == "") {
#         return(0)
#     } else if (i == "Male") {
#         return(1)
#     } else if(i == "Female") {
#         return(2)
#     }
# })
# 

meta <- na.omit(meta)
```

#### Step2. calculate the Propensity Score based on different variables
```{r}

ident = "PatientID"

cov <- colnames(meta)[colnames(meta) != ident]

command <- paste('meta %>% group_by(', ident, ') %>% select(one_of(cov)) %>% summarise_all(funs(mean(., na.rm = T)))', sep = '')

stats = as.data.frame(eval(parse(text = command)))
temp = stats[, !colnames(stats) %in% c(ident)]
p.values = data.frame(Var=colnames(temp), pValue=apply(temp, 2, function(x) {return(t.test(x)$p.value)}))

temp = meta
max_stage = max(temp[, ident])
for(i in 1:ncol(meta)) {
    temp[, i] <- temp[, i] / max(temp[, i])
}

m_ps <- glm(Stage ~ PatientID + Disease, family = binomial(), data = temp)

prs_df <- data.frame(pr_score = predict(m_ps, type = "response"), PatientID = m_ps$model$PatientID * max_stage)
```


```{r}
summary(m_ps)
```

#### check the distribution of score
```{r fig.height=16, fig.width=4}
prs_df$pr_score <- as.numeric(prs_df$pr_score)

ggplot(prs_df, aes(x = pr_score)) + geom_histogram(stat = "density") + facet_grid(PatientID~.)
```


#### TOMEK
```{r}
m = TomekClassif(PatientID ~ Stage, meta)
```

#### SMOTE
```{r}
m = SmoteClassif(Stage ~ PatientID + Disease, meta)
```

```{r}
ggplot(as.data.frame(table(obj@meta.data[, "Stage"])), aes(x=Var1, y = Freq)) + geom_bar(stat = "identity")

ggplot(as.data.frame(table(obj@meta.data[rownames(m), "Stage"])), aes(x=Var1, y = Freq)) + geom_bar(stat = "identity")
```
```{r}
ggplot(as.data.frame(table(obj@meta.data[, "PatientID"])), aes(x=Var1, y = Freq)) + geom_bar(stat = "identity")

ggplot(as.data.frame(table(obj@meta.data[rownames(m), "PatientID"])), aes(x=Var1, y = Freq)) + geom_bar(stat = "identity")
```

#### match it, to select cells
```{r}
mod_match <- matchit(PatientID ~ Stage + Disease, method = "nearest", data = temp)

dta_m <- match.data(mod_match)
dim(dta_m)
```


---

### 1. Load Basal cells seurat obj 

```{r}
obj <- readRDS(here("00_data_ingest", "04_rds_generated", "seurat_obj", "Basal_cells_processed.rds"))
obj@meta.data$old_cluster <- obj@meta.data$res.0.6


obj@meta.data$Stage <- as.character(obj@meta.data$Stage)
obj@meta.data[obj@meta.data$Batch == 3 & obj@meta.data$Stage == "", "Stage"] <- "Unknown"
obj@meta.data[obj@meta.data$Stage == "", "Stage"] <- "Benign"
```


### 2. calculate Propensity Score Matching
First of all, find out whether other variables have huge difference based on stage,

#### Step1. construct a counts table based on different variables
```{r}
meta = as.data.frame(obj@meta.data %>% select(PatientID, Sex, Stage, Disease) %>% group_by(PatientID))

meta$Stage <- as.data.frame(meta %>% select(Stage, Sex) %>% group_by(Sex) %>% add_tally())$n
# meta$Sex <- as.data.frame(meta %>% select(Stage, Sex) %>% group_by(Stage) %>% add_tally())$n
meta$Disease <- as.data.frame(meta %>% select(Sex, Disease) %>% group_by(Sex) %>% add_tally())$n

meta$PatientID <- as.integer(gsub("[^\\d]", "", meta$PatientID, perl = T))
# 
# 
# meta$Stage <- sapply(meta$Stage, function(i) {
#     if(i == "Benign") {
#         return(0)
#     } else if (i == "I") {
#         return(1)
#     } else if(i == "II") {
#         return(2)
#     } else if(i == "III") {
#         return(3)
#     } else if(i == "IV") {
#         return(4)
#     }
#     return(5)
# })
# 
# 
# meta$Disease <- sapply(meta$Disease, function(i) {
#     if(i == "Normal") {
#         return(0)
#     } else if (i == "Adenocarcinoma") {
#         return(1)
#     } else if(i == "Squamous cell carcinoma") {
#         return(2)
#     } else if(i == "Large cell neuroendocrine carcinoma") {
#         return(3)
#     } else if(i == "Lymphoid epithelioid carcinoma") {
#         return(4)
#     } else if(i == "") {
#         return(6)
#     }
#     return(5)
# })
# 
# 
meta$Sex <- sapply(meta$Sex, function(i) {
    if(i == "") {
        return(0)
    } else if (i == "Male") {
        return(1)
    } else if(i == "Female") {
        return(0)
    }
})
# 

meta <- na.omit(meta)
```

#### Step2. calculate the Propensity Score based on different variables
```{r}

ident = "Sex"

cov <- colnames(meta)[colnames(meta) != ident]

command <- paste('meta %>% group_by(', ident, ') %>% select(one_of(cov)) %>% summarise_all(funs(mean(., na.rm = T)))', sep = '')

stats = as.data.frame(eval(parse(text = command)))
temp = stats[, !colnames(stats) %in% c(ident)]
p.values = data.frame(Var=colnames(temp), pValue=apply(temp, 2, function(x) {return(t.test(x)$p.value)}))

temp = meta
max_stage = max(temp[, ident])
for(i in 1:ncol(meta)) {
    temp[, i] <- temp[, i] / max(temp[, i])
}

m_ps <- glm(Sex ~ Stage + PatientID + Disease, family = binomial(), data = temp)

prs_df <- data.frame(pr_score = predict(m_ps, type = "response"), Sex = m_ps$model$PatientID * max_stage)
```


```{r}
summary(m_ps)
```

#### check the distribution of score
```{r fig.height=16, fig.width=4}
prs_df$pr_score <- as.numeric(prs_df$pr_score)

ggplot(prs_df, aes(x = pr_score)) + geom_histogram(stat = "density") + facet_grid(Sex~.)
```


#### TOMEK
```{r}
m = TomekClassif(PatientID ~ Stage, meta)
```

#### SMOTE
```{r}
m = SmoteClassif(Stage ~ PatientID + Disease, meta)
```

```{r}
ggplot(as.data.frame(table(obj@meta.data[, "Stage"])), aes(x=Var1, y = Freq)) + geom_bar(stat = "identity")

ggplot(as.data.frame(table(obj@meta.data[rownames(m), "Stage"])), aes(x=Var1, y = Freq)) + geom_bar(stat = "identity")
```
```{r}
ggplot(as.data.frame(table(obj@meta.data[, "PatientID"])), aes(x=Var1, y = Freq)) + geom_bar(stat = "identity")

ggplot(as.data.frame(table(obj@meta.data[rownames(m), "PatientID"])), aes(x=Var1, y = Freq)) + geom_bar(stat = "identity")
```

#### match it, to select cells
```{r}
mod_match <- matchit(Sex ~ PatientID + Disease, method = "nearest", data = temp)

dta_m <- match.data(mod_match)
dim(dta_m)
```

---

#### check the matched cells
```{r}
fn_bal <- function(dta, variable) {
  dta$variable <- dta[, variable]
  if (variable == 'w3income') dta$variable <- dta$variable / 10^3
  dta$Stage <- as.factor(dta$Stage)
  support <- c(min(dta$variable), max(dta$variable))
  ggplot(dta, aes(x = distance, y = variable, color = Stage)) +
    geom_point(alpha = 0.2, size = 1.3) +
    geom_smooth(method = "loess", se = F) +
    xlab("Propensity score") +
    ylab(variable) +
    theme_bw() +
    ylim(support)
}

library(gridExtra)
grid.arrange(
   fn_bal(dta_m, "Batch"),
   fn_bal(dta_m, "PatientID") + theme(legend.position = "none"),
   # fn_bal(dta_m, "Age"),
   # fn_bal(dta_m, "Sex") + theme(legend.position = "none"),
   fn_bal(dta_m, "Disease"),
   nrow = 3, widths = c(1, 0.8)
)
```


```{r}
table(obj@meta.data[rownames(dta_m), "Stage"])
```

```{r}
table(obj@meta.data[rownames(dta_m), "PatientID"])
```


```{r}
obj <- readRDS('/mnt/raid62/Personal_data/huang_fei/single_cell/Lung_cancer/result/Basal_monocle3_process.rds')

pdata <- obj@phenoData@data
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
