---
title: "LungCancer10X_analysis"
author: "Yiming Zhang"
date: "2019/3/5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
root.dir = "/mnt/raid62/Lung_cancer_10x/"
knitr::opts_knit$set(root.dir = root.dir)

cols = 3
size = 10
dpi = 600
filenames = list()
```

```{r cars}
load_packages <- function() {
    library(here)
    library(gridExtra)
    library(dplyr)
    library(grid)
    library(monocle)
    library(reshape2)
    library(pheatmap)
    library(wesanderson)
    library(cowplot)
    library(AUCell)
}

suppressPackageStartupMessages(load_packages())
```

## Read seurat obj

```{r}
# obj <- readRDS(here("00_data_ingest", "04_rds_generated", "seurat_obj_selected_patients_harmony.rds"))
# 
# markers <- read.xlsx(here("00_data_ingest", "03_annotation_csv", "20190410_gene_markers.xlsx"), sheet = 5)

# markers <- read.xlsx("/mnt/raid62/Lung_cancer_10x/00_data_ingest/03_annotation_csv/20190410_gene_markers.xlsx", sheet = 5)

cell_names <- read.xlsx(here("00_data_ingest", "03_annotation_csv", "20190410_gene_markers.xlsx"), sheet = 4)

# cell_names <- read.xlsx("/mnt/raid62/Lung_cancer_10x/00_data_ingest/03_annotation_csv/20190410_gene_markers.xlsx", sheet = 4)

obj@meta.data$cluster_id <- obj@meta.data$res.0.6
obj@meta.data$cell_name <- obj@meta.data$res.0.6

for (i in 1:nrow(cell_names)) {
  idx = obj@meta.data$res.0.6 == cell_names[i, 1]
  
  obj@meta.data$cluster_id[idx] <- rep(paste(cell_names[i, 2], cell_names[i, 1], sep = "_"), sum(idx))
  
  obj@meta.data$cell_name[idx] <- rep(as.character(cell_names[i, 2]), sum(idx))
}

write.csv(obj@meta.data, "/mnt/raid62/Lung_cancer_10x/00_data_ingest/03_annotation_csv/20190412_seurat_meta.csv")

```


## 统计每种不同细胞系中，不同疾病类型的细胞数量
```{r}

# function to calculate freq or percentage by two select elements
# :param object: seurat obj
# :param first_ele: str, first select element, main elements
# :param second_ele: str, second select element
# :return dataframe of three columns, var1 -> first element, var2 -> second element, value -> percentage
make_perc_plot <- function(object, first_ele, second_ele, percentage = FALSE) {
    data <- object[, c("cluster_id", first_ele, second_ele)] 
    
    res = data.frame("var1"=NA, "var2"=NA, "value"=NA)
    for (i in unique(data[, first_ele])) {
        temp <- data[data[, first_ele] == i, ]
        
        uniq_second_ele = unique(as.vector(data[, second_ele]))
        temp_res = as.data.frame(matrix(NA, nrow = length(uniq_second_ele), ncol=3))

        for (j in 1:length(uniq_second_ele)) {
            temp_val <- sum(data[, second_ele] == uniq_second_ele[j] & data[, first_ele] == i)
            
            if (percentage) {
                temp_val <- temp_val / nrow(temp)
            }
            
            temp_res[j, ] <- c(i, uniq_second_ele[j], temp_val) 
        }
        
        colnames(temp_res) <- colnames(res)
        
        res <- rbind(temp_res, res)
    }
    
    res <- na.omit(res)
    res <- res[res[,3] > 0,]
    res[, 3] <- as.numeric(res[, 3])
    return(res)
}

```

#### 1. 简单看下每种细胞中不同Stage的细胞数量多少，便于后续筛选用于比对的细胞数

```{r fig.height=12, fig.width=8}
# function to draw barplot and heatmap
# :param data: Seurat obj@meta.data
make_heatmap_of_cells <- function(data, title="Total") {
    
    # make bar plot
    uniq_stage <- unique(data[, c("PatientID", "Stage")])
    uniq_stage <- as.data.frame(table(uniq_stage$Stage))
    
    p1 <- ggplot(data = uniq_stage, aes(x=Var1, y=Freq)) + geom_bar(stat = "identity") + labs(x="Stage", y="Num.", title = title)
    
    # make heatmap
    data <- make_perc_plot(data, first_ele = "cell_name", second_ele = "Stage")
    data <- acast(data, var1~var2)
    data[is.na(data)] <- 0
    data <- data[, order(colnames(data))]
    p2 <- pheatmap(log10(data + 1), cluster_cols = FALSE, silent = TRUE)
    
    filenames[[paste0("first_", i)]] <- here("03_analysis_plots", paste("heatmap_", i, ".png", sep = ""))
    
    png(filename = filenames[[paste0("first_", i)]], width = 8, height = 12, res = dpi, units = "in")
    grid.arrange(grobs = list(p1, p2[[4]]))
    dev.off()
    
    knitr::include_graphics(filenames[[paste0("first_", i)]])
}
```


#### 2. 统计下每中细胞所涉及到的cluster中不同的类别的细胞比例是多少

```{r fig.height=12, fig.width=8}

for(cell_name in unique(obj@meta.data$cell_name)) {
    
    print(cell_name)
    meta <- obj@meta.data[obj@meta.data$cell_name == cell_name,]
    patient <- make_perc_plot(meta, first_ele = "cluster_id", second_ele = "PatientID", percentage = TRUE)
    disease <- make_perc_plot(meta, first_ele = "cluster_id", second_ele = "Disease", percentage = TRUE)
    cells <- make_perc_plot(meta, first_ele = "cluster_id", second_ele = "cell_name")
    
    # make levels for order of plots
    levels <- c()
    for (i in sort(as.numeric(unique(meta$cluster_id)), decreasing = TRUE)) {
        levels <- c(levels, i)
    }
    
    patient$var1 <- factor(patient$var1, levels)
    p1 <- ggplot(patient, aes(x=var1, y=value, fill = var2)) +
        geom_bar(stat = "identity", position = position_stack()) + 
        coord_flip() +
        labs(x=cell_name, y = "Fraction of cells") +
        theme(legend.position = "top", axis.title.y = element_text(angle = 0)) + 
        guides(fill=guide_legend(
            title="Patient", 
            ncol = 4,
            title.position = "top"
            )) +
        scale_fill_manual(values=wes_palette("Zissou1", length(unique(patient$var2)), type = "continuous"))
    
    disease$var1 <- factor(disease$var1, levels)
    p2 <- ggplot(disease, aes(x=var1, y=value, fill = var2)) +
        geom_bar(stat = "identity", position = position_stack()) + 
        coord_flip() +
        labs(x="", y = "Fraction of cells") +
        theme(
            legend.position = "top", 
            axis.title.y = element_text(angle = 0),
            legend.justification = "center"
            ) + 
        guides(fill=guide_legend(
            title="Disease", 
            nrow = 10,
            title.position = "top"
            )) +
        scale_fill_manual(values=wes_palette("Darjeeling1", length(unique(disease$var2)), type = "continuous"))
    
    cells$var1 <- factor(cells$var1, levels)
    cells$var2 <- ""
    p3 <- ggplot(cells, aes(x=var1, y=log10(value + 1), fill = var2)) +
        geom_bar(stat = "identity") + 
        coord_flip() +
        labs(x="", y = "log10(Number of cells + 1)") +
        theme(legend.position = "top", axis.title.y = element_text(angle = 0)) + 
        guides(fill=guide_legend(
            title="", 
            nrow = 10,
            title.position = "top"
            )) +
        scale_fill_manual(values=wes_palette("GrandBudapest1", nrow(cells), type = "continuous"))
    
    meta$cluster_id = factor(meta$cluster_id, levels)
    p4 <- ggplot(meta, aes(x = cluster_id, y = log10(nGene + 1), fill = cluster_id)) + 
        geom_boxplot() + 
        coord_flip() +
        labs(x="") +
        theme(legend.position = "top", axis.title.y = element_text(angle = 0)) + 
        guides(fill=guide_legend(title="Clusters", nrow = 10, ncol=5)) +
        scale_fill_manual(values=wes_palette("Moonrise3", length(unique(meta$cluster_id)), type = "continuous"))
    
    phage <- meta[!is.na(meta$Phage) & meta$Phage != "", ]
    phage %>% select(cluster_id, Phage) %>% group_by(cluster_id) %>% count(Phage) -> phage
    p5 <- ggplot(phage, aes(x = cluster_id, y = log10(n), fill = Phage)) + 
        geom_bar(stat = "identity", position = position_stack()) +
        geom_text(aes(label=n), position = position_stack(vjust = 0.5)) +
        coord_flip() + 
        scale_fill_manual(values=wes_palette("Royal2", length(unique(phage$Phage)), type = "continuous")) +
        theme(legend.position = "top", axis.title.y = element_text(angle = 0)) + 
        guides(fill=guide_legend(title="", nrow = 10, ncol=5))

    cell_name = gsub("\\s", "_", cell_name)
    filenames[[paste("barplot", cell_name, sep = "_")]] <- here("03_analysis_plots", paste("barplot_", cell_name, ".png", sep = ""))
    png(filename=filenames[[paste("barplot", cell_name, sep = "_")]], res = dpi, width = 20, height = 15, units = "in")
    grid.newpage()
    p <- cbind(ggplotGrob(p1), ggplotGrob(p2), ggplotGrob(p5), ggplotGrob(p3), ggplotGrob(p4), size = "last")
    grid.draw(p)
    dev.off()
    
    # knitr::include_graphics(filenames[[paste("barplot", cell_name, sep = "_")]])
}

```


#### 3. 每个Stage随机选取每种细胞各1000个（不足1000全部选用），看其tSNE分布是否均匀

```{r}

for(i in unique(obj@meta.data$Stage)) {
    print(i)
    meta <- data.frame()
    
    for (j in unique(obj@meta.data$cluster_id)) {
        temp <- obj@meta.data[obj@meta.data$Stage == i & obj@meta.data$cluster_id == j, ]
        
        if (nrow(temp) > 1000) {
            meta <- rbind(meta, temp[sample(1:nrow(temp), 1000, replace = FALSE),])
        } else {
            meta <- rbind(meta, temp)
        }
    }
    
    p <- DimPlot(
        obj,
        reduction.use = "tSNE",
        pt.size = 0.01,
        cells.use = rownames(meta),
        group.by = "PatientID",
        do.label = TRUE,
        no.legend = FALSE,
        plot.title = i,
        do.return = TRUE
    )
    ggsave(filename = here("03_analysis_plots", paste("tSNE_", gsub("\\s+", "_", i), ".png", sep = "")), plot = p, width = 8, height = 6, dpi = dpi)
}

```


#### 4. 腺癌几个不同时期的细胞数量和分布情况
```{r}

for(i in unique(obj@meta.data$Phage)) {
    print(i)
    if ( i == "" || is.na(i) ) {
        
    } else {
        meta <- data.frame()
    
        for (j in unique(obj@meta.data$cluster_id)) {
            temp <- obj@meta.data[obj@meta.data$Phage == i & obj@meta.data$cluster_id == j, ]
            
            if (nrow(temp) > 1000) {
                meta <- rbind(meta, temp[sample(1:nrow(temp), 1000, replace = FALSE),])
            } else {
                meta <- rbind(meta, temp)
            }
        }
        
        p <- DimPlot(
            obj,
            reduction.use = "tSNE",
            pt.size = 0.01,
            cells.use = rownames(meta),
            group.by = "cell_name",
            do.label = TRUE,
            no.legend = TRUE,
            plot.title = i,
            do.return = TRUE
        )
        ggsave(filename = here("03_analysis_plots", paste("tSNE_phage_", gsub("\\s+", "_", i), ".png", sep = "")), plot = p, width = 6, height = 6, dpi = dpi)
    }
}

```



-----

### Fig 1. tSNE的重合图

```{r fig.height=6, fig.width=6}
temp <- obj@meta.data

temp <- cbind(temp, obj@dr$tsne@cell.embeddings[rownames(temp),])

p1 <- ggplot(temp, aes(x=tSNE_1, y=tSNE_2, color = Tissue)) + 
    geom_point(size = 0.01, alpha = .1) + 
    xlim(-18, 18) +
    ylim(-18, 18) +
    guides(colour = guide_legend(override.aes = list(size=5, alpha=1), ncol = 4)) +
    theme(legend.position = c(0.15, 0.03), legend.title = element_blank()) +
    labs(title = "Sample origin")

p1
```

```{r fig.height=6, fig.width=6}
p2 <- ggplot(temp, aes(x=tSNE_1, y=tSNE_2, color = PatientID)) + 
    geom_point(size = 0.01, alpha = .1) + 
    guides(colour = guide_legend(override.aes = list(size=10, alpha=1))) +
    labs(title = "Patient") + 
    theme(legend.position = "none", legend.title = element_blank()) +
    xlim(-18, 18) +
    ylim(-18, 18)
p2
```

```{r fig.height=6, fig.width=6}
temp1 <- as.data.frame(temp %>% dplyr::select(tSNE_1, tSNE_2, cell_name) %>% group_by(cell_name) %>% summarize(x = median(x = tSNE_1), y = median(x = tSNE_2)) %>% unique())


p3 <- ggplot(temp, aes(x=tSNE_1, y=tSNE_2, color = cell_name)) + 
    geom_point(size = 0.01, alpha = .1) + 
    geom_text(data = temp1, aes(label = cell_name, x = x, y = y), colour = "black", check_overlap = FALSE) + 
    theme(legend.position = "none") +
    labs(title = "Cell type") +
    xlim(-18, 18) +
    ylim(-18, 18)

p3

```


```{r fig.height=6, fig.width=6}
exp <- Matrix::colSums(obj@raw.data)
temp$count <- exp[rownames(temp)]

p4 <- ggplot(temp, aes(x=tSNE_1, y=tSNE_2, colour = log10(count + 1))) + 
    geom_point(size = 0.01) + 
    scale_color_gradientn(colours = c("grey", "blue")) +
    labs(title = "Transcript count") +
    xlim(-18, 18) +
    ylim(-18, 18) +
    theme(legend.position = c(0.01, 0.18), legend.title = element_blank()) +
    annotate(geom = "text", label="log10(count + 1)", x = -15, y = -17.8)

p4
```

```{r fig.height=6, fig.width=24}
p <- plot_grid(p1, p2, p3, p4, labels = "", align = 'h', rel_widths = c(1, 1, 1, 1), ncol = 4)

ggsave(filename = "/mnt/raid62/Lung_cancer_10x/01_figures/analysis/01_tsne.png", plot = p, width = 24, height = 6, units = "in", dpi = dpi)
```


---

### Fig 2. 不同事件的百分比图

1. Sample origin
```{r fig.height=18, fig.width=6}
temp1 <- as.data.frame(temp %>% dplyr::select(Tissue, res.0.6, cell_name) %>% group_by(res.0.6, Tissue) %>% add_tally() %>% unique() %>% group_by(res.0.6) %>% mutate(freq = n / sum(n)))


# plot_order <- temp1[order(temp1$cell_name, as.numeric(temp1$res.0.6)), "res.0.6"]
# 
# 
# temp1$plot_order <- factor(temp1$res.0.6, levels = unique(plot_order))
# 
# 
# ggplot(temp1, aes(x = plot_order, y = freq, fill = Tissue)) +
#     geom_bar(stat = "identity") +
#     facet_wrap(cell_name~., scales = "free_y", nrow = 1) +
#     coord_flip()

cell_names <- unique(temp1$cell_name)

plots = list()
rel_height = c()
for (i in 1:length(cell_names)) {
    temp2 <- temp1[temp1$cell_name == cell_names[i], ]
    rel_height[i] <- length(unique(temp2$res.0.6))
    if (i == 1) {
        plots[[i]] <- ggplot(temp2, aes(x = res.0.6, y = freq, fill = Tissue)) +
            geom_bar(stat = "identity")  +
            coord_flip() +
            labs(x="", y="") +
            theme(
                legend.position = "top",
                axis.title.x=element_blank(),
                axis.text.x=element_blank(),
                axis.ticks.x=element_blank()
            )
    } else if (i == length(cell_names)) {
          plots[[i]] <- ggplot(temp2, aes(x = res.0.6, y = freq, fill = Tissue)) +
            geom_bar(stat = "identity")  +
            coord_flip() +
            labs(x="", y="Fraction of cells") +
            theme(
                legend.position = "none"
            )
    } else {
        plots[[i]] <- ggplot(temp2, aes(x = res.0.6, y = freq, fill = Tissue)) +
            geom_bar(stat = "identity")  +
            coord_flip() +
            labs(x="", y="") +
            theme(
                legend.position = "none",
                axis.title.x=element_blank(),
                axis.text.x=element_blank(),
                axis.ticks.x=element_blank()
            )
    }
}

p1 <- plot_grid(plotlist=plots, labels = "", ncol = 1, rel_heights = rel_height)

```

```{r fig.height=30, fig.width=6}
temp1 <- as.data.frame(temp %>% dplyr::select(PatientID, res.0.6, cell_name) %>% group_by(res.0.6,PatientID) %>% add_tally() %>% unique() %>% group_by(res.0.6) %>% mutate(freq = n / sum(n)))


cell_names <- unique(temp1$cell_name)

plots = list()
for (i in 1:length(cell_names)) {
    temp2 <- temp1[temp1$cell_name == cell_names[i], ]

    if (i == 1) {
        plots[[i]] <- ggplot(temp2, aes(x = res.0.6, y = freq, fill = PatientID)) + 
            geom_bar(stat = "identity")  + 
            coord_flip() +
            labs(x="", y="") +
            theme(
                legend.position = "top", 
                axis.title.x=element_blank(),
                axis.text.x=element_blank(),
                axis.ticks.x=element_blank()
            )
    } else if (i == length(cell_names)) {
          plots[[i]] <- ggplot(temp2, aes(x = res.0.6, y = freq, fill = PatientID)) + 
            geom_bar(stat = "identity")  + 
            coord_flip() +
            labs(x="", y="Fraction of cells") +
            theme(
                legend.position = "none"
            ) 
    } else {
        plots[[i]] <- ggplot(temp2, aes(x = res.0.6, y = freq, fill = PatientID)) + 
            geom_bar(stat = "identity")  + 
            coord_flip() +
            labs(x="", y="") +
            theme(
                legend.position = "none", 
                axis.title.x=element_blank(),
                axis.text.x=element_blank(),
                axis.ticks.x=element_blank()
            )
    }
}
    
p2 <- plot_grid(plotlist=plots, labels = "", ncol = 1, rel_heights = rel_height)
```

```{r fig.height=18, fig.width=6}
temp1 <- as.data.frame(temp %>% dplyr::select(res.0.6, cell_name) %>% group_by(res.0.6) %>% add_tally() %>% unique())

plots = list()
for (i in 1:length(cell_names)) {
    temp2 <- temp1[temp1$cell_name == cell_names[i], ]

    if (i == 1) {
        plots[[i]] <- ggplot(temp2, aes(x = res.0.6, y = n)) + 
            geom_bar(stat = "identity")  + 
            coord_flip() +
            labs(x="", y="") +
            theme(
                legend.position = "top", 
                axis.title.x=element_blank(),
                axis.text.x=element_blank(),
                axis.ticks.x=element_blank()
            )
    } else if (i == length(cell_names)) {
          plots[[i]] <- ggplot(temp2, aes(x = res.0.6, y = n)) + 
            geom_bar(stat = "identity")  + 
            coord_flip() +
            labs(x="", y="Number of cells") +
            theme(
                legend.position = "none"
            ) 
    } else {
        plots[[i]] <- ggplot(temp2, aes(x = res.0.6, y = n)) + 
            geom_bar(stat = "identity")  + 
            coord_flip() +
            labs(x="", y="") +
            theme(
                legend.position = "none", 
                axis.title.x=element_blank(),
                axis.text.x=element_blank(),
                axis.ticks.x=element_blank()
            )
    }
}
    
p3 <- plot_grid(plotlist=plots, labels = "", ncol = 1, rel_heights = rel_height)
```

```{r fig.height=18, fig.width=6}
plots = list()
for (i in 1:length(cell_names)) {
    temp2 <- temp[temp$cell_name == cell_names[i], ]

    if (i == 1) {
        plots[[i]] <- ggplot(temp2, aes(x = res.0.6, y = nGene)) + 
            geom_boxplot() + 
            scale_y_log10() +
            coord_flip() +
            labs(x="", y="") +
            theme(
                legend.position = "top", 
                axis.title.x=element_blank(),
                axis.text.x=element_blank(),
                axis.ticks.x=element_blank()
            )
    } else if (i == length(cell_names)) {
          plots[[i]] <- ggplot(temp2, aes(x = res.0.6, y = nGene)) + 
            geom_boxplot() + 
            scale_y_log10() +
            coord_flip() +
            labs(x="", y="Number of genes (log10 scale)") +
            theme(
                legend.position = "none"
            ) 
    } else {
        plots[[i]] <- ggplot(temp2, aes(x = res.0.6, y = nGene)) + 
            geom_boxplot() + 
            scale_y_log10() +
            coord_flip() +
            labs(x="", y="") +
            theme(
                legend.position = "none", 
                axis.title.x=element_blank(),
                axis.text.x=element_blank(),
                axis.ticks.x=element_blank()
            )
    }
}

p4 <- plot_grid(plotlist=plots, labels = "", ncol = 1, rel_heights = rel_height)
```

```{r fig.height=30, fig.width=24}
plot_grid(plotlist = list(p1, p2, p3, p4), labels = "", nrow = 1, rel_heights = c(1.1, 1.5, 1, 1))
```

```{r}
temp <- cbind(temp, obj@dr$umap@cell.embeddings[rownames(temp), ])
```

---
### Fig3 AUCell

```{r}
cells_rankings <- readRDS(here("00_data_ingest", "04_rds_generated", "AUCell_cells_rankings_all.rds"))

cells_assignment <- readRDS(here("00_data_ingest", "04_rds_generated", "AUCell_cells_assignment_all.rds"))

cells_AUC <- readRDS(here("00_data_ingest", "04_rds_generated", "AUCell_cells_auc_al.rds"))

expMatrix <- readRDS(here("00_data_ingest", "00_raw_data", "raw_counts_clean.rds"))

tsne <- readRDS(here("00_data_ingest", "04_rds_generated", "seurat_harmony_tsne.rds"))

```

```{r}
for (i in 1:nrow(cells_AUC)) {
    print(cells_AUC[i, ]@NAMES)
    png(here("01_figures/analysis/aucell", paste(cells_AUC[i, ]@NAMES, ".png", sep = "")), res = 600, width = 6, height = 6, units = "in")
    AUCell_plotTSNE(tSNE=tsne$Y, cellsAUC=cells_AUC[i, ], exprMat = expMatrix, plots = c("binaryAUC"))
    dev.off()
}
```


---

### Fig4 各cluster的数量热图
```{r fig.height=15, fig.width=30}
temp <- as.data.frame(obj@meta.data %>% dplyr::select(cluster_id, PatientID) %>% dplyr::group_by(cluster_id, PatientID) %>% add_tally())

temp <- reshape2::acast(temp, cluster_id~PatientID)

pheatmap::pheatmap(temp, cluster_rows = F, cluster_cols = F, display_numbers = T, fontsize=15, number_format = "%d")
```

```{r fig.height=15, fig.width=15}
temp <- obj@meta.data
temp$new <- as.character(temp$Tissue)

temp$new[temp$Tissue == "Tumor"] <- as.character(temp$Disease[as.character(temp$Tissue) == "Tumor"])

temp$new[temp$new == ""] <- "Unknown"

temp <- as.data.frame(temp %>% dplyr::select(cluster_id, new) %>% dplyr::group_by(cluster_id, new) %>% add_tally())

temp <- reshape2::acast(temp, cluster_id~new)

pheatmap::pheatmap(temp, cluster_rows = F, cluster_cols = F, display_numbers = T, fontsize=15, number_format = "%d")
```

---

### Split whole cells into different seurat object

```{r}
output_dir = "/mnt/raid62/Lung_cancer_10x/00_data_ingest/04_rds_generated/seurat_obj"

meta <- read.csv("/mnt/raid62/Lung_cancer_10x/00_data_ingest/03_annotation_csv/20190412_seurat_meta.csv", row.names = 1)


for (i in unique(meta$cell_name)) {
    print(i)
    
    temp <- meta[meta$cell_name == i, ]
    
    temp <- CreateSeuratObject(raw.data=expMatrix[, rownames(temp)], project = i, meta.data = temp)
    
    saveRDS(temp, paste(output_dir, gsub(" ", "_", i), sep = "/"))
}
```

---

### 先拿Basal开刀


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
















