---
title: "R Notebook"
output: html_notebook
---

---
title: "LungCancer10X_sctransform"
output: html_document
---

Note: sctransform is integrated with Seurat 3.0

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
detach("package:Seurat", character.only = TRUE)

library(devtools)
library(withr)

libpath = "/home/zhangyiming/R/x86_64-pc-linux-gnu-library/3.5_seurat3.0/"


load_packages <- function() {
    library(Matrix)
    library(Seurat, lib.loc = libpath)
    library(here)
    library(data.table)
    library(gridExtra)
    library(dplyr)
    library(ggpubr)
    library(sctransform, lib.loc = libpath)
}

suppressPackageStartupMessages(load_packages())
options(future.globals.maxSize= 1000000000)

cols = 3
size = 10
dpi = 300
filenames = list()


custom_colors <- c(
    "#1B62A3",
    "#9CB8DD",
    "#EE6719",
    "#F8A966",
    "#278F36",
    "#8CC873",
    "#C70F1F",
    "#EF7E82",
    "#7D509A",
    "#B59BC7",
    "#76423A",
    "#B48880",
    "#CB60A1",
    "#F1A3C5",
    "#7E7F83",
    "#BDC3C4",
    "#ABB126",
    "#D1D57A",
    "#1DAFC3",
    "#8BCFDA"
)
```


### load Rds that pre-processed in harmony pipeline
```{r}
obj = readRDS(here("00_data_ingest", "04_rds_generated", "seurat_obj_selected_patients_base.rds"))

obj <- CreateSeuratObject(counts = obj@raw.data, meta.data = obj@meta.data)
```

```{r}
obj[["percent.mt"]] <- PercentageFeatureSet(object = obj, pattern = "^MT-")


filenames[["vln"]] <- here("01_figures", "sctransform", "VlnPlot.png")
png(filename = filenames[["vln"]], width = 18, height = 6, res = dpi, units = "in")
VlnPlot(object = obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
dev.off()
```

```{r}
filenames[["par"]] <- here("01_figures", "sctransform", "GenePlot.png")

png(filename = filenames[["par"]], width = 12, height = 6, res = dpi, units = "in")

plot1 <- FeatureScatter(object = obj, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(object = obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))

dev.off()
```


```{r}
obj <- NormalizeData(object = obj, normalization.method = "LogNormalize", scale.factor = 10000)

obj <- FindVariableFeatures(object = obj, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(x = VariableFeatures(object = obj), 10)

# plot variable features with and without labels
filenames[["par1"]] <- here("01_figures", "sctransform", "GenePlot1.png")

png(filename = filenames[["par1"]], width = 12, height = 6, res = dpi, units = "in")

plot1 <- VariableFeaturePlot(object = obj)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
CombinePlots(plots = list(plot1, plot2))

dev.off()
```

```{r}
obj <- ScaleData(object = obj, vars.to.regress = c("percent.mt", "nCount_RNA"))

obj <- RunPCA(object = obj, features = VariableFeatures(object = obj), npcs = 100)
```


```{r}
obj <- SCTransform(object = obj, verbose = TRUE)
```

```{r}
obj <- RunPCA(object = obj, verbose = TRUE, npcs = 100)
```

```{r}
filenames[["pc"]] <- here("01_figures", "sctransform", "ElbowPlot.png")
png(filename = filenames[["pc"]], width = 12, height = 6, res = dpi, units = "in")
ElbowPlot(object = obj, ndims=100)
dev.off()
```
```{r}
knitr::include_graphics(filenames[["pc"]])
```

```{r}

n.pcs = 50
obj <- RunHarmony(obj, c("batch"))
obj <- RunTSNE(object = obj, dims = 1:n.pcs, verbose = FALSE, reduction = "pca")
obj <- RunUMAP(object = obj, dims = 1:n.pcs, verbose = TRUE, reduction = "pca")
obj <- FindNeighbors(object = obj, dims = 1:n.pcs, verbose = FALSE, reduction = "pca")
obj <- FindClusters(object = obj, dims = 1:n.pcs, verbose = FALSE, reduction = "pca")
```

```{r}
p1 <- DimPlot(obj, reduction = "tsne", group.by = "SCT_snn_res.0.8", legend = TRUE, label = TRUE, do.return = TRUE, pt.size = 0.1) + 
  theme(legend.position = "none")
p2 <- DimPlot(obj, reduction = "tsne", group.by = "batch", legend = FALSE, label = FALSE, do.return = TRUE, pt.size = 0.1)

p3 <- DimPlot(obj, reduction = "umap", group.by = "SCT_snn_res.0.8", legend = TRUE, label = TRUE, do.return = TRUE, pt.size = 0.1) + 
  theme(legend.position = "none")
p4 <- DimPlot(obj, reduction = "umap", group.by = "batch", legend = FALSE, label = FALSE, do.return = TRUE, pt.size = 0.1)

p <- ggarrange(p1, p2, p3, p4, nrow=2, ncol=2)

filenames[["harmony_tsne"]] <- here("01_figures", "sctransform", "sctransform_tsne.png")
ggsave(filenames[["harmony_tsne"]], plot = p, width = 12, height = 12, dpi = dpi, units = "in")
```


```{r}
saveRDS(obj, here("00_data_ingest", "04_rds_generated", "seurat_obj_selected_patients_sctransform.rds"))
```

```{r}
markers <- read.xlsx(here("00_data_ingest", "03_annotation_csv", "20190326_gene_markers.xlsx"), sheet = 3)

genes <- intersect(unique(markers$Markers), rownames(x=obj))

p <- DotPlot(obj, features = genes, split = "SCT_snn_res.0.8", x.lab.rot = TRUE, do.return = TRUE) + coord_flip() + labs(title = "Harmony")

filenames[["dotplot"]] <- here("01_figures", "sctransform", "sctransform_dotplot.png")
ggsave(filename = filenames[["dotplot"]], plot = p, width = 10, height = 20, units = "in", dpi = dpi)
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

