---
title: "R Notebook"
output: html_notebook
---

The huge memory usage is not affordable, therefore, we use the loom version Seurat instead.

Version 

```{r}
# install.packages("onlinePCA")
# devtools::install_github(repo = "hhoeflin/hdf5r")
# devtools::install_github(repo = "mojaveazure/loomR", ref = "develop")
# devtools::install_github(repo = "satijalab/seurat", ref = "loom")

load_packages <- function() {
    library("loomR")
    library("Seurat")
    library("here")
}

suppressPackageStartupMessages(load_packages())

filenames = list()
image_dir = "04_cca_plots/loom"
dir.create(here(image_dir), showWarnings = FALSE)
dpi = 300

sessionInfo()
```


## 1. Convert seurat obj to loom
```
obj <- readRDS(here("00_data_ingest", "04_rds_generated", "seurat_obj_cca_by_batch_3.rds"))

batch1 = CreateSeuratObject(
    raw.data = obj@raw.data[,rownames(obj@meta.data[obj@meta.data$Batch == 1, ])], 
    meta.data = obj@meta.data, 
    project="first"
)
batch2 = CreateSeuratObject(
    raw.data = obj@raw.data[,rownames(obj@meta.data[obj@meta.data$Batch == 2, ])], 
    meta.data = obj@meta.data, 
    project="second"
)
batch3 = CreateSeuratObject(
    raw.data = obj@raw.data[,rownames(obj@meta.data[obj@meta.data$Batch == 3, ])], 
    meta.data = obj@meta.data, 
    project="nature_medicine"
)

batch1 <- Convert(
    from = batch1,
    to = "loom",
    filename = here("00_data_ingest", "04_rds_generated", "seurat_obj_cca_batch_1.loom"),
    display.progress = TRUE
)

batch2 <- Convert(
    from = batch2, 
    to = "loom", 
    filename = here("00_data_ingest", "04_rds_generated", "seurat_obj_cca_batch_2.loom"), 
    display.progress = TRUE
)

batch3 <- Convert(
    from = batch3, 
    to = "loom", 
    filename = here("00_data_ingest", "04_rds_generated", "seurat_obj_cca_batch_3.loom"), 
    display.progress = TRUE
)

rm(obj)
```

```{r}
batch1 <- as.SingleCellExperiment(batch1)
batch2 <- as.SingleCellExperiment(batch2)
batch3 <- as.SingleCellExperiment(batch3)

correct <- fastMNN(batch1, batch2, batch3)
```

```{r}
batch1 <- connect(filename = here("00_data_ingest", "04_rds_generated", "seurat_obj_cca_batch_1.loom"), mode = "r+")
batch2 <- connect(filename = here("00_data_ingest", "04_rds_generated", "seurat_obj_cca_batch_2.loom"), mode = "r+")
batch3 <- connect(filename = here("00_data_ingest", "04_rds_generated", "seurat_obj_cca_batch_3.loom"), mode = "r+")
```


```{r}
genes.use = c()
batch_list = list(batch1, batch2, batch3)

for(i in 1:3) {
    NormalizeData(object = batch_list[[i]], display.progress = TRUE, overwrite=TRUE)
    FindVariableGenes(batch_list[[i]])

    genes.use <- c(genes.use, head(x = GetVariableGenes(object = batch_list[[i]])$index, n = 1000))
}

genes.use <- intersect(genes.use, rownames(ctrl@scale.data))
genes.use <- intersect(genes.use, rownames(stim@scale.data))


obj <- RunMultiCCA(list(batch1, batch2, batch3), num.ccs = 30)
```

```{r}
data <- read.csv("zcat /mnt/raid62/Lung_cancer_10x/00_data_ingest/00_raw_data/02_raw_counts_by_patients/PLC017.csv.gz")
data <- as.data.frame(data)
rownames(data) <- data$V1
data <- data[, !colnames(data) %in% c("V1")]
```

```{r}
dpi = 300
num.pcs = 100


parser <- ArgumentParser()

# specify our desired options 
# by default ArgumentParser will add an help option 
parser$add_argument(
    "-i", 
    "--input", 
    type="character",
    help="Path to input csv.gz [default %(default)s]"
)
parser$add_argument(
    "-o", 
    "--output", 
    help="Path to output directoy"
)
parser$add_argument(
    "-l",
    "--label",
    help="",
    default="LungCancer10X"
)
parser$add_argument(
    "-m",
    "--meta",
    help=""
)
parser$add_argument(
    "--dpi",
    type="integer",
    default=300
)
parser$add_argument(
    "--pcs",
    type="integer",
    default=100
)

# get command line options, if help option encountered print help and exit,
# otherwise if options not found on command line then set defaults, 
args <- parser$parse_args()

if (is.na(args$input)) {
    stop("no input")
} else if (is.na(args$input)) {
    stop("no output")
}

input = args$input
output = args$output
label = args$label
meta = args$meta

if(dir.exists(output)) {
    dir.create(output, recursive = TRUE)
}


data <- fread(paste("zcat", input, collapse = " "))


to_sparse <- function(d_table){
  
  i_list <- lapply(d_table, function(x) which(x != 0))
  counts <- unlist(lapply(i_list, length), use.names = F)

  return(sparseMatrix(
    i = unlist(i_list, use.names = F),
    j = rep(1:ncol(d_table), counts),
    x = unlist(lapply(d_table, function(x) x[x != 0]), use.names = F),
    dims = dim(d_table),
    dimnames = list(NULL, names(d_table)))
  )
}

data1 <- as.data.frame(data)


obj <- CreateSeuratObject(raw.data = data, min.cells = 3, min.genes = 200, project = label)

mito.genes <- grep(pattern = "^MT-", x = rownames(x = obj@data), value = TRUE)
percent.mito <- Matrix::colSums(obj@raw.data[mito.genes, ])/Matrix::colSums(obj@raw.data)
obj <- AddMetaData(object = obj, metadata = percent.mito, col.name = "percent.mito")

p <- VlnPlot(object = obj, features.plot = c("nGene", "nUMI", "percent.mito"), nCol = 3, do.return = TRUE)
ggsave(filename = paste(output, "00_vlnplot.png", collapse = "/"), plot = p, width = 12, height = 6, dpi = dpi, units = "in")


png(paste(output, "01_geneplot.png", collapse = "/"), width = 12, height = 6, units = "in", res = dpi)
par(mfrow = c(1, 2))
GenePlot(object = pbmc, gene1 = "nUMI", gene2 = "percent.mito")
GenePlot(object = pbmc, gene1 = "nUMI", gene2 = "nGene")
dev.off()

obj <- NormalizeData(object = obj, normalization.method = "LogNormalize", scale.factor = 10000)
obj <- FindVariableGenes(object = obj, mean.function = ExpMean, dispersion.function = LogVMR, x.low.cutoff = 0.0125, x.high.cutoff = 3, y.cutoff = 0.5)
obj <- ScaleData(object = obj, vars.to.regress = c("nUMI", "percent.mito"))
obj <- RunPCA(object = obj, pc.genes = obj@var.genes, pcs.compute = num.pcs,do.print = FALSE)


png(paste(output, "02_pcaplot.png", collapse = "/"), width = 6, height = 6, units = "in", res = dpi)
PCAPlot(object = obj, dim.1 = 1, dim.2 = 2)
dev.off()


png(paste(output, "03_pcelbowplot.png", collapse = "/"), width = 12, height = 6, units = "in", res = dpi)
PCElbowPlot(object = obj, num.pc = num.pcs)
dev.off()

# determine how many pcs to use
pc.sd = apply(obj@dr$pca@cell.embeddings, 2, sd)
diff.pc.sd = diff(pc.sd)
sd.difference = max(pc.sd) - min(pc.sd)
num.pcs = max(which(abs(diff.pc.sd) > 0.005 * sd.difference))

# make pca heatmap
if (num.pcs > 30) {
    num.pcs.heatmap = 30
    heatmap_row = 10
} else {
    heatmap_row = ceiling(num.pcs / 3)
    num.pcs.heatmap = num.pcs
}

png(paste(output, "04_pcheatmap.png", collapse = "/"), width = 12, height = 4 * heatmap_row, units = "in", res = dpi)
PCHeatmap(mca, pc.use = 1:num.pcs.heatmap, cells.use = 500, do.balanced = TRUE)
dev.off()


obj <- FindClusters(object = obj, reduction.type = "pca", dims.use = 1:num.pcs, resolution = 0.6, print.output = 0, save.SNN = TRUE)
obj <- RunTSNE(object = obj, dims.use = 1:num.pcs, do.fast = TRUE)
obj <- RunUMAP(object = obj, dims.use = 1:num.pcs)

# make tsne and umap plot
p1 <- TSNEPlot(object = obj, do.label = TRUE, pt.size = 0.1, label.size = 10, no.legend=TRUE, do.return=TRUE, dim.use = "tsne")
p2 <- TSNEPlot(object = obj, do.label = TRUE, pt.size = 0.1, label.size = 10, no.legend=TRUE, do.return=TRUE, dim.use = "umap")

p <- ggarrange(p1, p2, ncol = 2)
ggsave(
    filename = paste(output, "05_tsne_umap.png", collapse = "/"), 
    plot = p, 
    units = "in", 
    width = 20, 
    height = 10,
    dpi = dpi
)


# make dot plot
marker <- read.csv(meta)
marker <- intersect(marker$Markers, rownames(obj@scale.data))

p <- DotPlot(object = obj, genes.plot = marker, do.return = TRUE, x.lab.rot = TRUE)
p <- p + coord_flip()
ggsave(
    filename = paste(output, "06_dotplot.png", collapse = "/"), 
    plot = p, 
    units = "in", 
    width = length(unique(obj@meta.data$res.0.6)), 
    height = 0.2 * length(marker),
    dpi = dpi
)

saveRDS(obj, paste(output, "seurat_obj.rds", collapse = "/"))
```

```{r fig.height=4, fig.width=12}
temp <- as.data.frame(colSums(obj@dr$pca@cell.embeddings))


y <- apply(obj@dr$pca@cell.embeddings, 2, sd)
x <- 1:100
model <- lm(y~I(1/x))

which(abs(model$residuals) > 0.1)

```
```{r}
x = seq(1,15)
y = c(4,5,6,5,5,6,7,8,7,7,6,6,7,8,9)
plot(x,y,type="l",ylim=c(3,10))
lo <- loess(y~x)
xl <- seq(min(x),max(x), (max(x) - min(x))/1000)
out = predict(lo,xl)
lines(xl, out, col='red', lwd=2)

infl <- c(FALSE, diff(diff(y)>0)!=0)
points(xl[infl ], out[infl], col="blue")
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

