---
title: "LungCancer10X_cca"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

load_packages <- function() {
    library("Seurat")
    library("dplyr")
    library("cowplot")
    library("repr")
    library("openxlsx")
    library("gridExtra")
    library("here")
    library("harmony")
}

suppressPackageStartupMessages(load_packages())

filenames = list()

dpi = 300

cols = 3
size = 10
dpi = 300
filenames = list()


custom_colors <- c(
    "#1B62A3",
    "#9CB8DD",
    "#EE6719",
    "#F8A966",
    "#278F36",
    "#8CC873",
    "#C70F1F",
    "#EF7E82",
    "#7D509A",
    "#B59BC7",
    "#76423A",
    "#B48880",
    "#CB60A1",
    "#F1A3C5",
    "#7E7F83",
    "#BDC3C4",
    "#ABB126",
    "#D1D57A",
    "#1DAFC3",
    "#8BCFDA"
)
```



#### 1. load seurat obj
```{r}
obj =  readRDS(here("00_data_ingest", "04_rds_generated", "seurat_obj_selected_patients_base.rds"))
```


### 2. load meta info from all samples
```{r}
meta <- read.csv(here("00_data_ingest", "03_annotation_csv", "20190407_seurat_meta.csv"))
```


### 3. Create new seurat object, which only kept normal and cancer
the cells in batch 3 and batch 2 is too few
```{r}
obj <- CreateSeuratObject(
    raw.data = obj@raw.data[, rownames(obj@meta.data[obj@meta.data$Tissue %in% c("Normal", "Tumor"), ])],
    meta.data = obj@meta.data[obj@meta.data$Tissue %in% c("Normal", "Tumor"), ],
    project = "Normal_vs_cancer"
)
```

```{r}
mito.genes <- grep(pattern = "^MT-", x = rownames(x = obj@raw.data), value = TRUE)
percent.mito <- Matrix::colSums(obj@raw.data[mito.genes, ])/Matrix::colSums(obj@raw.data)

# AddMetaData adds columns to object@meta.data, and is a great place to
# stash QC stats
obj <- AddMetaData(object = obj, metadata = percent.mito, col.name = "percent.mito")

obj <- NormalizeData(object = obj, normalization.method = "LogNormalize", scale.factor = 10000)
obj <- FindVariableGenes(object = obj, mean.function = ExpMean, dispersion.function = LogVMR, x.low.cutoff = 0.0125, x.high.cutoff = 3, y.cutoff = 0.5)
obj <- ScaleData(object = obj, vars.to.regress = c("nUMI", "percent.mito"))
obj <- RunPCA(object = obj, pc.genes = obj@var.genes, do.print = FALSE, pcs.compute = 100)
```
```{r}
obj <- RunHarmony(obj, c("Tissue"))
```

```{r}
temp <- as.data.frame(apply(obj@dr$harmony@cell.embeddings, 2, sd))
colnames(temp) <- c("sd")
temp$Harmony = 1:nrow(temp)

ggplot(temp, aes(x=Harmony, y=sd)) + geom_point() + ylab("Standard Dei of Harmony")
```

## 分群感觉分的不是很开，调整参数重新分群看效果
```{r}
n.pcs = 25
obj <- RunTSNE(object = obj, dims.use = 1:n.pcs, do.fast = TRUE, reduction.use = "harmony", reduction.name = "tsne")
# obj <- RunUMAP(object = obj, dims.use = 1:n.pcs, do.fast = TRUE, reduction.use = "harmony")
obj <- FindClusters(object = obj, reduction.type = "harmony", dims.use = 1:n.pcs, resolution = 0.6, print.output = 0, save.SNN = FALSE, force.recalc = TRUE)
```

```{r}
saveRDS(obj, here("00_data_ingest", "04_rds_generated", "seurat_obj_selected_patients_harmony_normal_vs_cancer.rds"))
```

### 4. check the CCA effect

```{r include=FALSE}
p1 <- DimPlot(
    object = obj, 
    reduction.use = "cca", 
    group.by = "Batch", 
    pt.size = 0.5, 
    do.return = TRUE
)
p2 <- VlnPlot(
    object = obj, 
    features.plot = "CC1", 
    group.by = "Batch", 
    do.return = TRUE
)

filenames[["dim_vln_plot"]] <- here(image_dir, "00_dim_vln_plot.png")
png(filename = filenames[["dim_vln_plot"]], width = 12, height = 6, res = dpi, units = "in")
plot_grid(p1, p2)
dev.off()
```
```{r}
knitr::include_graphics(filenames[["dim_vln_plot"]])
```


#### perform cluster and tSNE

```{r}
obj <- RunPCA(
    object = obj,
    pc.genes = var_genes,
    pcs.compute = 100,
    do.print = TRUE,
    pcs.print = 1,
    genes.print = 1
)
```
```{r include=FALSE}
p1 <- DimPlot(
    object = obj, 
    reduction.use = "pca", 
    group.by = "Batch", 
    pt.size = 0.5, 
    do.return = TRUE
)
p2 <- VlnPlot(
    object = obj, 
    features.plot = "PC1", 
    group.by = "Batch", 
    do.return = TRUE
)

filenames[["pca_dim_vln_plot"]] <- here(image_dir, "05_pca_dim_vln_plot.png")
png(filename = filenames[["pca_dim_vln_plot"]], width = 12, height = 6, res = dpi, units = "in")
plot_grid(p1, p2)
dev.off()
```
```{r}
knitr::include_graphics(filenames[["pca_dim_vln_plot"]])
```

### 6. decide use how many cca after
```{r include=FALSE}
filenames[["MetageneBicor"]] <- here(image_dir, "01_meta_gene_bicor_plot.png")
png(filename = filenames[["MetageneBicor"]], width = 12, height = 6, res = dpi, units = "in")
MetageneBicorPlot(
    obj, 
    grouping.var = "Batch",
    dims.eval = 1:30, 
    display.progress = FALSE
)
dev.off()
```
```{r}
knitr::include_graphics(filenames[["MetageneBicor"]])
```


```{r}
filenames[["DimHeatmap"]] <- here(image_dir, "02_cca_dim_heatmap_plot.png")
png(filename = filenames[["DimHeatmap"]], width = 12, height = 40, res = dpi, units = "in")
DimHeatmap(object = obj, reduction.type = "cca", cells.use = 500, dim.use = 1:30, do.balanced = TRUE)
dev.off()
```
```{r}
knitr::include_graphics(filenames[["DimHeatmap"]])
```


### 7. align cca subspace and calculate pca
- PCA is used for the CCA effect comparision
```{r}
n.ccs = 5

obj <- AlignSubspace(obj, reduction.type = "cca", grouping.var = "Batch", dims.align = 1:n.ccs)

# saveRDS(obj, here("00_data_ingest", "04_rds_generated", "seurat_obj_selected_patients_cca_aligned_2.rds"))
```


### 8. decide how many pcs to use
```{r include=FALSE}
filenames[["PCEbow"]] <- here(image_dir, "03_pca_bow_plot.png")
png(filename = filenames[["PCEbow"]], width = 12, height = 6, res = dpi, units = "in")
PCElbowPlot(object = obj, num.pc = 100)
dev.off()
```
```{r}
knitr::include_graphics(filenames[["PCEbow"]])
```


```{r fig.width=18, fig.height=24}
filenames[["DimHeatmap"]] <- here(image_dir, "04_pca_heatmap_plot.png")
png(filename = filenames[["DimHeatmap"]], width = 12, height = 6, res = dpi, units = "in")
PCHeatmap(
    object = obj,
    pc.use = 1:12,
    cells.use = 500,
    do.balanced = TRUE,
    label.columns = FALSE,
    do.return = TRUE
)
dev.off()
```
```{r}
knitr::include_graphics(filenames[["DimHeatmap"]])
```

### 9. Run tSNE
```{r}
n.pcs = 50

obj <- RunTSNE(obj, reduction.use = "cca.aligned", dims.use = 1:n.ccs, do.fast = T)
obj <- RunTSNE(obj, reduction.use = "pca", dim.use = 1:n.pcs, reduction.name = "tSNE", reduction.key = "tSNE_")
obj <- FindClusters(obj, genes.use = var_genes, reduction.type = "cca.aligned", resolution = 0.6, dims.use = 1:n.ccs, force.recalc=TRUE, print.output = FALSE)
```



```{r}
# 
# 整合各种已有数据
row.names(meta) <- meta$Cell
after_cca@meta.data <- cbind(after_cca@meta.data, meta[meta$Cell %in% rownames(after_cca@meta.data),])
after_cca@meta.data <- cbind(after_cca@meta.data, after_cca@dr$tSNE@cell.embeddings)
after_cca@meta.data <- cbind(after_cca@meta.data, after_cca@dr$umap@cell.embeddings)
after_cca@meta.data <- cbind(after_cca@meta.data, as.data.frame(after_cca@ident))
colnames(after_cca@meta.data)[ncol(after_cca@meta.data)] = "cluster_id"


saveRDS(after_cca, "20190226_cca_normal_tumor.rds")

after_cca <- readRDS("20190226_cca_normal_tumor.rds")
```

```{r}
p1 <- TSNEPlot(after_cca, do.return = T, pt.size = 0.5, group.by = "stim")
p2 <- TSNEPlot(after_cca, do.label = T, do.return = T, pt.size = 0.5)
plot_grid(p1, p2)
```




```{r}
after_cca@meta.data$cell_name <- after_cca@meta.data$cluster_id

current.cluster.ids <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24)
new.cluster.ids <- c(
    "CD8 T cells", # 0
    "T cells", # 1,
    "CD14+ Monocytes",  # 2
    "NK cells",  # 3
    "CD14+ Monocytes",  # 4
    "NK cells",  # 5
    "Basal cells", # 6
    "Dendritic", # 7
    "CD14+ Monocytes",  # 8
    "CD14+ Monocytes", # 9
    "Club cells", # 10
    "Club cells", # 11
    12,
    "T cells", # 13,
    "B cells", # 14 
    15, 
    16, 
    "CD14+ Monocytes", # 17
    "CD14+ Monocytes", # 18
    19, 
    "T cells", # 20, 
    21, 22, 23, 
    "Basal cells" # 24
)


after_cca@meta.data$cell_name <- plyr::mapvalues(x = after_cca@meta.data$cell_name, from = current.cluster.ids, to = new.cluster.ids)

DimPlot(
    object = obj, 
    reduction.use = "tSNE", 
    pt.size = 0.01, 
    do.return = TRUE,
    group.by = "cell_name", 
    do.label = TRUE
)

FeaturePlot(
    object = obj, 
    features.plot = c(
        "MS4A1",    # B
        "GNLY",     # NK
        "CD3E",    
        "CD14",     # CD14+ Monocytes
        "FCER1A",   # Dendritic Cells
        "FCGR3A",   # FCGR3A+ Monocytes
        "LYZ",      # CD14+ Monocytes
        "PPBP",     # Megakaryocytes
        "CD8A",     # CD8
        "NKG7",     # NK
        "KRT5",     # basal
        "SCGB1A1",   # club,
        "CCR7",      # CD8
        "C1orf186"   # mast
    ), 
    cols.use = c(
        "grey", 
        "blue"
    ), 
    reduction.use = "tSNE"
)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
