---
title: "Gene module"
author: "Zhang Yiming"
date: "2020/2/14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
root.dir = "LungCancer10x/"

full.path <- function(...) { return(paste(root.dir, ..., sep = "/")) }
```


```{r}
Stage = c("I"="#65A9A3", "II"="#4A933E", "III"="#EC7A21", "IV"="#D73F47", "Normal"="#73BDFF")
Disease = c( 
    "LUAD" = "#0084D1",  "NL(AD)"="#FECC1B",  "NL"="#778793", 
    "BSPN"="#C93311",  "NL(SC)"="#778793",  "LUSC"="#A0C807"
)
Source = c("In-house"="#F8AFA8",  "East Asians"="#FDDDA0","GTEx"="#73A089", "TCGA"="#877E45")




colors = list(
    Stage = c(
        "I"="#65A9A3", 
        "II"="#4A933E", 
        "III"="#EC7A21", 
        "IV"="#D73F47", 
        "LUAD_Normal"="#FECC1B", 
        "LUSC_Normal"="#778793"
    ),
    Disease = c(
        "AD" = "#0084D1", 
        "NL(AD)"="#FECC1B", 
        "NL(AD"="#778793", 
        "Normal"="#73BDFF", 
        "NL(SC)"="#778793", 
        "SC"="#A0C807",
        "LC"="#91822B",
        "NL(LC)"="#EDCAB0",
        "UPS"="#EBAAA4",
        "NL(UPS)"="#B43018"
    ),
    Batch = c("1"="#E9B0B7", "2"="#90D0E0", "3"="#769982"),
    Cell = c(
        "ATII"="#FF0000",
        "Basal"="#F2AD00",
        "CD4"="#F98400",
        "CD8"="#5BBCD6",
        "Cilia"="#85D4E3",
        "Club"="#F4B5BD",
        "DC"="#9C964A",
        "EC"="#CDC08C",
        "Epi"="#FAD77B",
        "Fib"="#ECCBAE",
        "Mast"="#D69C4E",
        "Mo"="#ABDDDE",
        "NK"="#000000",
        "NE"="#F3DF6C",
        "Tregs"="#CEAB07",
        "Mφ"="#046C9A"
    )
)

cluster_cols = c(
    as.character(wes_palette("Zissou1")),
    as.character(wes_palette("Royal2")),
    as.character(wes_palette("Royal1")),
    as.character(wes_palette("Darjeeling2")),
    as.character(wes_palette("Darjeeling1")),
    wes_palette("BottleRocket2"),
    wes_palette("Rushmore1")
)
```

## Load Packages

```{r cars, include=FALSE}
library(ggplot2)
library(Seurat)
library(wesanderson)
library(ComplexHeatmap)
library(circlize)
library(tsidx)
library(dplyr)
library(reshape2)
library(clusterProfiler)
library(org.Hs.eg.db)
library(stringr)

options(stringsAsFactors = F)
```

## Batch run

```{r eval=FALSE, include=FALSE}
files = list.files(
    full.path("03_each_cells/Batch"), 
    pattern = "seurat.rds", 
    full.names = T, 
    recursive = T
)


do_tsidx <- function(obj, maximum=8) {
    tryCatch({
         tsi <- tsDataSetFromMatrix(
            obj@raw.data[, rownames(obj@meta.data)], 
            obj@meta.data, 
            ~ Stage, 
            type = "expression",
            convMethod = "log2", 
            minimum=1, 
            maximum=maximum, 
            mindiff=ifelse(maximum - 1 < 3, 1, 3),
            IbMethod = "mingap", 
            tidy = FALSE, 
            breaks = NULL
        )
         
         return(tsi)
    }, error = function(e) {
        return(do_tsidx(obj, maximum - 1))
    })
}



for (f in files) {
    obj <- readRDS(f)

    obj@meta.data$Stage <- factor(
        obj@meta.data$Stage, 
        levels = intersect(c("I", "II", "III", "IV"), unique(obj@meta.data$Stage))
    )
    
    tsi <- do_tsidx(obj)
    saveRDS(tsi, paste0(dirname(f), "/tsi.rds"))
}

```

## Batch plotdensity of tsi

```{R}
files = list.files(
    full.path("03_each_cells/Batch"), 
    pattern = "tsi.rds", 
    full.names = T, 
    recursive = T
)


for (f in sort(files)) {
    cell = basename(dirname(dirname(f)))
    disease = basename(dirname(f))
    
    tsi <- readRDS(f)
    p <- ggplot(as.data.frame(tsi@tsData$Tau), aes(x=Tau)) + 
        geom_density() +
        labs(title = paste0(cell, "(", disease, ")"))
    
    print(p)
}
```


## ATII LUAD
```{r}
tsi <- readRDS(full.path("03_each_cells/Batch/ATII/LUAD/tsi.rds"))


plotHeatmap(tsi, TauMin = 0.75, show_colnames = F)
```

```{r}
for (f in files) {
    cell = basename(dirname(dirname(f)))
    disease = basename(dirname(f))
    
    tsi <- readRDS(f)
    
    meta = as.data.frame(tsi@colData)
    meta = meta[order(meta$Stage), ]
    
    meta <- meta %>%
        dplyr::group_by(Stage) %>%
        sample(100, replace = T) %>%
        unique() %>%
        as.data.frame()
    
    tau = as.data.frame(tsi@tsData$Tau)
    tau = tau[tau$Tau > 0.75, ]
    
    rank = melt(as.matrix(tsi@tsData$Rank)[rownames(tau), ])
    rank = rank %>%
        group_by(Var1) %>%
        top_n(1, wt=value) %>%
        filter(value != 0) %>%
        unique() %>%
        as.data.frame()
    rownames(rank) = rank$Var1
    
    sel_gene = intersect(rownames(tau), rownames(rank))
    
    
    tryCatch({
       　ta = HeatmapAnnotation(
            Stage = meta$Stage,
            col = list(Stage=Stage),
            show_legend = T,
            show_annotation_name = F,
            annotation_legend_param = list(
                labels_gp = gpar(fontsize = 15),
                title_gp = gpar(fontsize = 20)
                # grid_height = unit(2, "cm")
            )
        )
    
        
        
        h <- Heatmap(
            t(scale(t(tsi@rawData[sel_gene, meta$Cells]))),
            name = "Expr",
            cluster_rows = F,
            cluster_columns = F,
            show_row_names = F,
            show_column_names = F,
            row_split = rank[sel_gene, "Var2"],
            column_split = meta$Stage,
            col = colorRamp2(c(-3, 0, 3), c("blue", "white", "red")),
            column_title = paste0(cell, "(", disease, ")"),
            border = T
        )
        
        draw(h)
    }, error = function(e) {})
   
}

```

```{r}
make_stage_module_heatmap <- function(
    data,
    meta, 
    module, 
    col_fun = NULL, 
    labels_size=15, 
    title_size = 20, 
    cluster=F, 
    do.return=F, 
    show_row_names = F, 
    do.random = F,
    do.balance = T,
    order.by.patient = T,
    mark_genes = NULL,
    cluster_row = T,
    row_split = NULL,
    column_title = "",
    show_disease = F
) {
    if (is.null(col_fun)) {
        col_fun = colorRamp2(c(-5, 0, 5), c("blue", "white", "red"))
    } 
    
    meta$Disease <- as.character(meta$Disease)
    meta$Disease[meta$Disease == "LUAD"] <- "AD"
    meta$Disease[meta$Disease == "LUSC"] <- "SC"
    meta$Disease[meta$Disease == "LUAD_Normal"] <- "NL(AD)"
    meta$Disease[meta$Disease == "LUSC_Normal"] <- "NL(SC)"
    meta$Disease[meta$Disease == "Pleiomorphic Normal"] <- "NL(UPS)"
    meta$Disease[meta$Disease == "Pleiomorphic"] <- "UPS"
    meta$Disease[meta$Disease == "LULC"] <- "LC"
    meta$Disease[meta$Disease == "LULC_Normal"] <- "NL(LC)"
    
    patient_colors = cluster_cols[1:length(unique(meta$PatientID))]
    names(patient_colors) <- sort(as.character(unique(meta$PatientID)))
    module = as.character(module)

    if(do.balance) {
        minmum_cells = sort(table(meta$Stage), decreasing = F)
    
        if (minmum_cells[1] < 100) {
            minmum_cells = minmum_cells[2]
        } else {
            minmum_cells = minmum_cells[1]
        }
        
        
        meta <- meta %>%
            group_by(Stage) %>%
            sample_n(minmum_cells, replace = T) %>%
            unique() %>%
            as.data.frame()
    
        rownames(meta) <- meta$Cells
    }
    
    
    set.seed(42)
    if(order.by.patient) {
       if (show_disease) {
            meta <- meta[order(meta$Disease, meta$Stage, meta$PatientID), c("Stage", "Disease", "PatientID")]
        
            anno_legend_param = list(
                Stage=list(
                    direction = "horizontal", nrow = 1,
                    labels_gp = gpar(fontsize = labels_size),
                    title_gp = gpar(fontsize = title_size)
                ),
                PatientID=list(
                    direction = "horizontal", nrow = 4,
                    labels_gp = gpar(fontsize = labels_size),
                    title_gp = gpar(fontsize = title_size)
                ),
                Disease=list(
                    direction = "horizontal", nrow = 1,
                    labels_gp = gpar(fontsize = labels_size),
                    title_gp = gpar(fontsize = title_size)
                )
            )
       } else {
            meta <- meta[order(meta$Disease, meta$Stage, meta$PatientID), c("Stage", "PatientID")]
        
            anno_legend_param = list(
                Stage=list(
                    direction = "horizontal", nrow = 1,
                    labels_gp = gpar(fontsize = labels_size),
                    title_gp = gpar(fontsize = title_size)
                ),
                PatientID=list(
                    direction = "horizontal", nrow = 4,
                    labels_gp = gpar(fontsize = labels_size),
                    title_gp = gpar(fontsize = title_size)
                )
            )
       }
    } else {
        if (do.random) {
            meta = meta[sample(1:nrow(meta)), ]
        }
        
        
        if(!show_disease) {
             meta <- meta[order(meta$Disease, meta$Stage), c("Stage"), drop = F]
             anno_legend_param = list(
                Stage=list(
                    direction = "horizontal", nrow = 1,
                    labels_gp = gpar(fontsize = labels_size),
                    title_gp = gpar(fontsize = title_size)
                )
            )
        } else {
             meta <- meta[order(meta$Disease, meta$Stage), c("Stage", "Disease")]
             anno_legend_param = list(
                Stage=list(
                    direction = "horizontal", nrow = 1,
                    labels_gp = gpar(fontsize = labels_size),
                    title_gp = gpar(fontsize = title_size)
                ),
                Disease=list(
                    direction = "horizontal", nrow = 1,
                    labels_gp = gpar(fontsize = labels_size),
                    title_gp = gpar(fontsize = title_size)
                )
            )
        }
       
    }

    # top annotation
    ta = HeatmapAnnotation(
        df = meta,
        col = list(
            Stage =  colors[["Stage"]],
            PatientID = patient_colors,
            Disease = colors[["Disease"]]
        ),
        show_legend = T,
        show_annotation_name = T,
        annotation_legend_param = anno_legend_param,
        annotation_name_gp = gpar(fontsize = title_size - 2)
    )
    
    if(!is.null(mark_genes)) {
        mark_genes = as.character(mark_genes)
        ra = rowAnnotation(
            foo = anno_mark(
                at = which(module %in% mark_genes),
                labels = module[module %in% mark_genes],
                which = "row",
                side = "right"
            ),
            show_legend = F, show_annotation_name = F
        )
    } else {
        ra = NULL
    }
    
    module_cols = as.character(wes_palette("GrandBudapest2"))[1:length(unique(row_split))]
    names(module_cols) <- sort(unique(as.character(row_split)))

    la = rowAnnotation(
        Module = row_split,
        col = list(Module =  module_cols),
        show_legend = T,
        show_annotation_name = T,
        annotation_legend_param = list(
            Module=list(
                direction = "horizontal", nrow = 1,
                labels_gp = gpar(fontsize = labels_size),
                title_gp = gpar(fontsize = title_size)
            )
        ),
        annotation_name_rot = 0,
        annotation_name_gp = gpar(fontsize = title_size - 2)
    )
    
    h <- Heatmap(
        name = "Expr",
        data[as.character(module), rownames(meta)],
        cluster_rows = cluster_row,
        cluster_columns = F,
        show_column_names = F,
        show_row_names = show_row_names,
        col = col_fun,
        top_annotation = ta,
        left_annotation = la,
        right_annotation = ra, 
        column_split = paste(meta$Disease, meta$Stage),
        row_split = row_split,
        border = T,
        column_title = column_title,
        column_title_gp = gpar(fontsize=ifelse(column_title == "", 0, 20)),
        row_title_gp = gpar(fontsize = 0),
        heatmap_legend_param = list(
            direction = "horizontal",
            labels_gp = gpar(fontsize = labels_size),
            title_gp = gpar(fontsize = title_size)
        )
    )
    
    if(do.return) {
        return(h)
    } else {
        draw(
            h, 
            column_title_gp = gpar(fontsize = 20),
            merge_legend = TRUE, 
            heatmap_legend_side = "bottom", 
            annotation_legend_side = "bottom"
        )
    }
}

```


## make ATII new module plot
```{r}
obj <- readRDS(full.path("03_each_cells/Batch/ATII/LUAD/seurat.rds"))
tsi <- readRDS(full.path("03_each_cells/Batch/ATII/LUAD/tsi.rds"))ibrary(dplyr)
mfuzz <- read.csv(full.path("03_each_cells/Batch/ATII/LUAD/mfuzz.csv"), row.names = 1, stringsAsFactors = F)
mfuzz_dup = table(mfuzz$gene)
mfuzz_dup = names(mfuzz_dup)[mfuzz_dup > 1]
mfuzz <- mfuzz[!mfuzz$gene %in% mfuzz_dup, ]

markers <- read.csv(full.path("03_each_cells/Batch/ATII/LUAD/cluster_markers.csv"), row.names = 1, stringsAsFactors = F)
marekrs <- markers %>%
    filter(avg_logFC > 0.25 & p_val_adj < 0.05) %>%
    group_by(ident) %>%
    top_n(10, wt = avg_logFC)
```


```{r}
tau = as.data.frame(tsi@tsData$Tau)
tau = tau[tau$Tau > 0.75, ]

rank = melt(as.matrix(tsi@tsData$Rank)[rownames(tau), ])
rank = rank %>%
    group_by(Var1) %>%
    top_n(1, wt=value) %>%
    filter(value != 0) %>%
    unique() %>%
    as.data.frame()
rownames(rank) = rank$Var1draw(
    h, column_title_gp = gpar(fontsize = 20),
    merge_legend = TRUE, heatmap_legend_side = "bottom", annotation_legend_side = "bottom"
)
rank = rank[rank$Var2 == "III", ]

sel_gene = as.character(rank$Var1)
```



```{r fig.height=8, fig.width=10}
set.seed(42)
mfuzz_sel = mfuzz %>%
    group_by(Clt) %>%
    sample_n(5) %>%
    as.data.frame()

mfuzz_sel = mfuzz_sel$gene


h <- make_stage_module_heatmap(
    obj@scale.data, 
    obj@meta.data, 
    mfuzz$gene,
    order.by.patient = F,
    do.random = T,
    col_fun = colorRamp2(c(-1, 0, 1), c("purple", "black", "yellow")),
    cluster_row = F,
    row_split = mfuzz$Clt,
    mark_genes = c(as.character(marekrs$gene), as.character(mfuzz_sel), "XIST"),
    do.return = T
)


pdf(full.path("03_each_cells/Batch/ATII/LUAD/mfuzz.pdf"), width = 10, height = 8)
draw(
    h, column_title_gp = gpar(fontsize = 20),
    merge_legend = TRUE, heatmap_legend_side = "bottom", annotation_legend_side = "bottom"
)
dev.off()

draw(
    h, column_title_gp = gpar(fontsize = 20),
    merge_legend = TRUE, heatmap_legend_side = "bottom", annotation_legend_side = "bottom"
)
```


```{r}
go = NULL

for (i in unique(mfuzz$Clt)) {
    eg = bitr(mfuzz$gene[mfuzz$Clt == i], fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)
    
    temp = enrichGO(
        eg$ENTREZID,
        org.Hs.eg.db,
        ont = "BP"
    )
    
    temp = simplify(temp)
    
    temp = as.data.frame(temp)
    temp$Clt = i
    go = rbind(go,temp)
}

go$GeneRatioPercentage <- sapply(go$GeneRatio, function(x) {
    x = str_split(x, "/")[[1]]
    as.numeric(as.character(x[1])) / as.numeric(as.character(x[2])) * 100
}) 

write.csv(go, full.path("03_each_cells/Batch/ATII/LUAD/mfuzz_go.csv"))
```



### Test Stage module on bulk deconvolution
```{r}
tree <- read.csv(full.path("09_bulk/RNA_seq/MuSic/All_tree.csv"), row.names = 1, stringsAsFactors = F)

sc_set <- readRDS(full.path("09_bulk/RNA_seq/MuSic/sc_set.rds"))

est <- read.csv(full.path("09_bulk/RNA_seq/MuSic/All_est.csv"), row.names = 1, stringsAsFactors = F)
```


```{r echo=FALSE, fig.height=8, fig.width=8}
### using cutree and max number of samples as label
make_heatmap_number_merged <- function(
    data, meta, 
    scale_dot_size=0.9, 
    row_split=NULL, column_split=NULL,
    cluster_rows = T, cluster_columns = T,
    show_row_names = F, show_column_names = T,
    show_source=T
) {
    col_fun <- colorRamp2(c(0, 0.5, 1), c("blue", "white", "red"))
    
    if(show_source) {
        colors = list(Stage = Stage, Disease = Disease, Source = Source)
        
        ra <- rowAnnotation(
            Disease = meta[rownames(data), "Disease"], Stage = meta[rownames(data), "Stage"], Source = meta[rownames(data), "Source"],
            col = colors, show_annotation_name = F, show_legend = T
        )
    } else {
        colors = list(Stage = Stage, Disease = Disease)
        
        ra <- rowAnnotation(
            Disease = meta[rownames(data), "Disease"], Stage = meta[rownames(data), "Stage"],
            col = colors, show_annotation_name = F, show_legend = T
        )
    }
    
    Heatmap(
        data,
        name = "Weight",
        col = col_fun,
        right_annotation = ra,
        # bottom_annotation = ba,
        show_row_names = show_row_names,
        show_column_names = show_column_names,
        row_split = row_split,
        column_split = column_split,
        cluster_rows = cluster_rows,
        cluster_columns = cluster_columns,
        border = T
    )
}


make_piechart_of_stages_each_group <- function(tree, meta, ncol=2) {
    temp = tree
    temp$Stage = meta[rownames(temp), "Stage"]
    
    temp <- temp %>%
        group_by(Stage, Clt) %>%
        add_tally() %>%
        dplyr::select(Clt, Stage, n) %>%
        unique() %>%
        group_by(Clt) %>%
        mutate(label = paste0(Stage, "(", round(n / sum(n) * 100, digits = 2), "%)")) %>%
        as.data.frame()
    

    
    plist = list()
    for (i in sort(unique(temp$Clt))) {
        plist[[length(plist) + 1]] <- ggpie(
            temp[temp$Clt == i, ],
            "n", 
            label = "label", 
            fill = "Stage", 
            color = "white",
            palette = Stage,
            title = i,
            lab.pos = "in"
        ) + theme(legend.position = "none")
    }
    
    cowplot::plot_grid(plotlist = plist, ncol = ncol)
}
```



```{r}
best_ncluster <- function(data, method="wss") {
    # "silhouette", "wss",  "gap_stat"
    
    fviz_nbclust(data, kmeans, method = "wss") +
        geom_vline(xintercept = 4, linetype = 2)+
        labs(subtitle = "Elbow method")
}


kmean_cluster <- function(data, k=4, seed=42) {
    set.seed(seed)
    
    temp_kmeans <- kmeans(data, k)
    temp_tree <- as.data.frame(temp_kmeans$cluster)
    colnames(temp_tree) <- "Clt"
    temp_tree
}
```


## Inhouse
```{r}
bulk <- read.xlsx(full.path("09_bulk/RNA_seq/RSEM.xlsx"), rowNames = T)
gene_pos <- read.table(full.path("09_bulk/RNA_seq/gene_pos.txt"), header = F, row.names = 5)
gene_pos$V6 <- make.unique(as.character(gene_pos$V6))
rownames(bulk) <- gene_pos[rownames(bulk), "V6"]


bulk_meta <- read.xlsx(full.path("09_bulk/RNATAC50commonSample-1119.xlsx"))
bulk_meta <- bulk_meta[!is.na(bulk_meta$SampleID), ]
rownames(bulk_meta) <- bulk_meta$SampleID
bulk_meta$Stage <- str_replace(bulk_meta$Stage, "[^IV]", "")

sel_sample = rownames(tree[tree$Source == "In-house" & tree$Disease == "LUAD" & tree$Clt == 2, ])

bulk_set <- ExpressionSet(as.matrix(bulk[, sel_sample]), phenoData = AnnotatedDataFrame(bulk_meta[sel_sample, ]))

bulk_dec = music_prop(
    bulk.eset = bulk_set, 
    sc.eset = sc_set, 
    clusters = 'cell_short',
    samples = 'SampleID', 
    markers = mfuzz$gene,
    verbose = F
)
```


```{r echo=FALSE}
temp_est = bulk_dec$Est.prop.weighted

temp_tree = kmean_cluster(temp_est, k = 4)


make_heatmap_number_merged(
    temp_est,
    bulk_meta, 
    row_split = temp_tree[rownames(temp_est), "Clt"], 
    # column_split = rownames(temp_tree),
    cluster_rows = F,
    cluster_columns = T,
    show_column_names = T
)
```


## TCGA

```{R}
tcga <- read.csv(full.path("TCGA/Lung.csv"), row.names = 1, stringsAsFactors = F)


### read bulk
tcga_meta <- data.frame(
    Disease = sapply(colnames(tcga), function(x) { str_split(x, "\\.")[[1]][8] }),
    Stage = sapply(colnames(tcga), function(x) { str_split(x, "\\.")[[1]][9] })
)

rownames(tcga_meta) <- colnames(tcga)
tcga_meta <- tcga_meta[tcga_meta$Stage != "", ]
```

```{r}
sel_sample = rownames(tree[tree$Source == "TCGA" & tree$Disease == "LUAD" & tree$Clt == 1, ])
 
table(tcga_meta[sel_sample, "Stage"])
```

#### Random select samples by stage
```{r}
set.seed(42)
tcga_meta$Sample = rownames(tcga_meta)

sel_sample = tcga_meta[sel_sample, ] %>%
    group_by(Stage) %>%
    sample_n(40, replace = T) %>%
    unique() %>%
    as.data.frame()

sel_sample = sel_sample$Sample
```


## Test TCGA on all expression
```{r include=FALSE}

tcga_set <- ExpressionSet(
    as.matrix(tcga[, sel_sample]), 
    phenoData = AnnotatedDataFrame(tcga_meta[sel_sample,])
)


tcga_dec = music_prop(
    bulk.eset = tcga_set, 
    sc.eset = sc_set, 
    clusters = 'Stage',
    samples = 'SampleID', 
    markers = mfuzz$gene,
    verbose = T
)
    
```

### Check samples
```{r echo=FALSE}
temp_est = tcga_dec$Est.prop.weighted

temp_tree = kmean_cluster(temp_est, k = 4)


make_heatmap_number_merged(
    temp_est,
    tcga_meta, 
    row_split = temp_tree[rownames(temp_est), "Clt"], 
    # column_split = rownames(temp_tree),
    cluster_rows = F,
    cluster_columns = T,
    show_column_names = T
)

make_piechart_of_stages_each_group(temp_tree, tcga_meta)
```


### modify expression by ATII 

```{r include=FALSE}
tcga_set <- ExpressionSet(
    as.matrix(tcga[, sel_sample] / t(est["ATII", sel_sample])), 
    phenoData = AnnotatedDataFrame(tcga_meta[sel_sample,])
)


tcga_dec = music_prop(
    bulk.eset = tcga_set, 
    sc.eset = sc_set, 
    clusters = 'Stage',
    samples = 'SampleID', 
    markers = mfuzz$gene,
    verbose = T
)
```


### Check samples



```{r echo=FALSE}
temp_est = tcga_dec$Est.prop.weighted

temp_tree = kmean_cluster(temp_est, k = 4)


make_heatmap_number_merged(
    temp_est,
    tcga_meta, 
    row_split = temp_tree[rownames(temp_est), "Clt"], 
    # column_split = rownames(temp_tree),
    cluster_rows = F,
    cluster_columns = T,
    show_column_names = T
)

make_piechart_of_stages_each_group(temp_tree, tcga_meta)
```


### Only use ATII expression to make deconvolution
```{r include=FALSE}

atii_set = ExpressionSet(
    as.matrix(sc_set@assayData$exprs[, rownames(sc_set@phenoData@data)[sc_set@phenoData@data$cell_short == "ATII"]]), 
    phenoData = AnnotatedDataFrame(sc_set@phenoData@data[sc_set@phenoData@data$cell_short == "ATII", ])
)


tcga_set <- ExpressionSet(
    as.matrix(tcga[, sel_sample]), 
    phenoData = AnnotatedDataFrame(tcga_meta[sel_sample,])
)


tcga_dec = music_prop(
    bulk.eset = tcga_set, 
    sc.eset = atii_set, 
    clusters = 'Stage',
    samples = 'SampleID', 
    markers = mfuzz$gene,
    verbose = T
)
```


```{r echo=FALSE}
temp_est = tcga_dec$Est.prop.weighted

temp_tree = kmean_cluster(temp_est, k = 6)


make_heatmap_number_merged(
    temp_est,
    tcga_meta, 
    row_split = temp_tree[rownames(temp_est), "Clt"], 
    # column_split = rownames(temp_tree),
    cluster_rows = F,
    cluster_columns = T,
    show_column_names = T
)

make_piechart_of_stages_each_group(temp_tree, tcga_meta, ncol = 3)
```


### Only use ATII expression to make deconvolution
```{r include=FALSE}
tcga_set <- ExpressionSet(
    as.matrix(tcga[, sel_sample] / t(est["ATII", sel_sample])), 
    phenoData = AnnotatedDataFrame(tcga_meta[sel_sample,])
)


tcga_dec = music_prop(
    bulk.eset = tcga_set, 
    sc.eset = atii_set, 
    clusters = 'Stage',
    samples = 'SampleID', 
    markers = mfuzz$gene,
    verbose = T
)
```


```{r echo=FALSE}
temp_est = tcga_dec$Est.prop.weighted

temp_tree = kmean_cluster(temp_est, k = 6)


make_heatmap_number_merged(
    temp_est,
    tcga_meta, 
    row_split = temp_tree[rownames(temp_est), "Clt"], 
    # column_split = rownames(temp_tree),
    cluster_rows = F,
    cluster_columns = T,
    show_column_names = T
)

make_piechart_of_stages_each_group(temp_tree, tcga_meta, ncol=3)
```




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.