---
title: "ATII_gene_module"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
root.dir = "LungCancer10x/03_each_cells/ATII_Basal/"
knitr::opts_knit$set(root.dir = root.dir)
```


```{R include=FALSE}
options(stringsAsFactors = F)
library(ComplexHeatmap)
library(circlize)
library(stringr)
library(wesanderson)
library(dplyr)
library(reshape2)
library(Seurat)
library(doMC)
library(openxlsx)
```



```{R}
colors = list(
    Stage = c(
        "I"="#65A9A3", 
        "II"="#4A933E", 
        "III"="#EC7A21", 
        "IV"="#D73F47", 
        "LUAD_Normal"="#FECC1B", 
        "LUSC_Normal"="#778793"
    ),
    Disease = c(
        "AD" = "#0084D1", 
        "NL(AD)"="#FECC1B", 
        "NL(AD"="#778793", 
        "Normal"="#73BDFF", 
        "NL(SC)"="#778793", 
        "SC"="#A0C807",
        "LC"="#91822B",
        "NL(LC)"="#EDCAB0",
        "UPS"="#EBAAA4",
        "NL(UPS)"="#B43018"
    ),
    Batch = c("1"="#E9B0B7", "2"="#90D0E0", "3"="#769982"),
    Cell = c(
        "ATII"="#FF0000",
        "Basal"="#F2AD00",
        "CD4"="#F98400",
        "CD8"="#5BBCD6",
        "Cilia"="#85D4E3",
        "Club"="#F4B5BD",
        "DC"="#9C964A",
        "EC"="#CDC08C",
        "Epi"="#FAD77B",
        "Fib"="#ECCBAE",
        "Mast"="#D69C4E",
        "Mo"="#ABDDDE",
        "NK"="#000000",
        "NE"="#F3DF6C",
        "Tregs"="#CEAB07",
        "Mφ"="#046C9A"
    )
)

cluster_cols = c(
    as.character(wes_palette("Zissou1")),
    as.character(wes_palette("Royal2")),
    as.character(wes_palette("Royal1")),
    as.character(wes_palette("Darjeeling2")),
    as.character(wes_palette("Darjeeling1")),
    wes_palette("BottleRocket2"),
    wes_palette("Rushmore1")
)



make_stage_module_heatmap <- function(
    data,
    meta, 
    module, 
    col_fun = NULL, 
    labels_size=15, 
    title_size = 20, 
    cluster=F, 
    do.return=F, 
    show_row_names = F, 
    order.by.patient = T,
    mark_genes = NULL,
    cluster_row = T,
    row_split = NULL,
    column_title = ""
) {
    if (is.null(col_fun)) {
        col_fun = colorRamp2(c(-5, 0, 5), c("blue", "white", "red"))
    } 
    
    meta$Disease <- as.character(meta$Disease)
    meta$Disease[meta$Disease == "LUAD"] <- "AD"
    meta$Disease[meta$Disease == "LUSC"] <- "SC"
    meta$Disease[meta$Disease == "LUAD_Normal"] <- "NL(AD)"
    meta$Disease[meta$Disease == "LUSC_Normal"] <- "NL(SC)"
    meta$Disease[meta$Disease == "Pleiomorphic Normal"] <- "NL(UPS)"
    meta$Disease[meta$Disease == "Pleiomorphic"] <- "UPS"
    meta$Disease[meta$Disease == "LULC"] <- "LC"
    meta$Disease[meta$Disease == "LULC_Normal"] <- "NL(LC)"
    
    patient_colors = cluster_cols[1:length(unique(meta$PatientID))]
    names(patient_colors) <- sort(as.character(unique(meta$PatientID)))
    
    if(order.by.patient) {
        meta <- meta[order(meta$Disease, meta$Stage, meta$PatientID), c("Stage", "Disease", "PatientID")]
        
        anno_legend_param = list(
            Stage=list(
                direction = "horizontal", 
                nrow = 1,
                labels_gp = gpar(fontsize = labels_size),
                title_gp = gpar(fontsize = title_size)
            ),
            PatientID=list(
                direction = "horizontal",
                nrow = 4,
                labels_gp = gpar(fontsize = labels_size),
                title_gp = gpar(fontsize = title_size)
            ),
            Disease=list(
                direction = "horizontal",
                nrow = 1,
                labels_gp = gpar(fontsize = labels_size),
                title_gp = gpar(fontsize = title_size)
            )
        )
    } else {
        meta <- meta[order(meta$Disease, meta$Stage), c("Stage", "Disease")]
        
        anno_legend_param = list(
            Stage=list(
                direction = "horizontal", 
                nrow = 1,
                labels_gp = gpar(fontsize = labels_size),
                title_gp = gpar(fontsize = title_size)
            ),
            Disease=list(
                direction = "horizontal",
                nrow = 1,
                labels_gp = gpar(fontsize = labels_size),
                title_gp = gpar(fontsize = title_size)
            )
        )
    }

   
    # top annotation
    ta = HeatmapAnnotation(
        df = meta,
        col = list(
            Stage =  colors[["Stage"]],
            PatientID = patient_colors,
            Disease = colors[["Disease"]]
        ),
        show_legend = T,
        show_annotation_name = T,
        annotation_legend_param = anno_legend_param
    )
    
    if(!is.null(mark_genes)) {
        la = rowAnnotation(
            foo = anno_mark(
                at = which(module %in% mark_genes),
                labels = module[module %in% mark_genes],
                which = "row",
                side = "left"
            ),
             show_legend = F,
            show_annotation_name = F
        )
    } else {
        la = NULL
    }
    
    
    h <- Heatmap(
        name = "Expr",
        data[as.character(module), rownames(meta)],
        cluster_rows = cluster_row,
        cluster_columns = F,
        show_column_names = F,
        show_row_names = show_row_names,
        col = col_fun,
        top_annotation = ta,
        left_annotation = la, 
        column_split = paste(meta$Disease, meta$Stage),
        row_split = row_split,
        border = T,
        column_title = column_title,
        column_title_gp = gpar(fontsize=ifelse(column_title == "", 0, 20)),
        row_title_gp = gpar(fontsize = 0),
        heatmap_legend_param = list(
            direction = "horizontal",
            labels_gp = gpar(fontsize = labels_size),
            title_gp = gpar(fontsize = title_size)
        )
    )
    
    if(do.return) {
        return(h)
    } else {
        draw(
            h, 
            column_title_gp = gpar(fontsize = 20),
            merge_legend = TRUE, 
            heatmap_legend_side = "bottom", 
            annotation_legend_side = "bottom"
        )
    }
}



make_module_heatmap <- function(obj, res, cells.use = NULL, col_fun = NULL, order_by_patient = TRUE, gene_mark = NULL) {
    labels_size = 12
    title_size = 15
    meta = obj@meta.data
    
    if(!is.null(cells.use)) {
        meta <- meta[cells.use, ]
    }
    
    if (order_by_patient) {
        meta = meta[order(meta$Stage, meta$PatientID), ]
        
        df = data.frame(
            Stage = meta$Stage,
            PatientID = meta$PatientID
        )
        
        legend_param = list(
            Stage=list(
                direction = "horizontal", 
                nrow = 1,
                labels_gp = gpar(fontsize = labels_size),
                title_gp = gpar(fontsize = title_size)
            ),
            PatientID=list(
                direction = "horizontal",
                nrow = 4,
                labels_gp = gpar(fontsize = labels_size),
                title_gp = gpar(fontsize = title_size)
            )
        )
    } else {
        meta = meta[sample(1:nrow(meta)), ]
        meta = meta[order(meta$Stage), ]
        
        df = data.frame(
            Stage = meta$Stage
        )
        
        legend_param = list(
            Stage=list(
                direction = "horizontal", 
                nrow = 1,
                labels_gp = gpar(fontsize = labels_size),
                title_gp = gpar(fontsize = title_size)
            )
        )
    }
    
    stage_colors <- c(
        "I"="#65A9A3", 
        "II"="#4A933E",
        "III"="#EC7A21",
        "IV"="#D73F47"
    )
    
    cluster_cols = c(
        as.character(wes_palette("Zissou1")),
        as.character(wes_palette("Royal2")),
        as.character(wes_palette("Royal1")),
        as.character(wes_palette("Darjeeling2")),
        as.character(wes_palette("Darjeeling1"))
    )
    
    patient_colors = cluster_cols[1:length(unique(obj@meta.data$PatientID))]
    names(patient_colors) <- unique(obj@meta.data$PatientID)
    
    # top annotation
    ta = HeatmapAnnotation(
        df = df,
        col = list(
            Stage =  stage_colors,
            PatientID = patient_colors
        ),
        show_legend = T,
        show_annotation_name = T,
        annotation_legend_param = legend_param
    )
    
    
    module_colors = wes_palette("GrandBudapest2", length(unique(res$Clt)), type = "continuous")
    names(module_colors) <- unique(res$Clt)
    
    ra = NULL
    if (!is.null(gene_mark)) {
        gene_mark = gene_mark[gene_mark %in% res$gene]
        
        at = which(as.character(res$gene) %in% gene_mark)
        
        ra = rowAnnotation(
            gene = anno_mark(
                at = at,
                labels = as.character(res$gene[at]),
                which = "row",
                side = "right"
            )
        )
    }
    
    # left annotation
    la = rowAnnotation(
        Module = res$Clt,
        col = list(
            Module = module_colors
        ),
        show_legend = T,
        show_annotation_name = T,
        annotation_name_rot = 0,
        annotation_legend_param=list(
            Module=list(
                direction = "horizontal", 
                nrow = 1,
                labels_gp = gpar(fontsize = labels_size),
                title_gp = gpar(fontsize = title_size)
            )
        )
    )
    
    
    if (is.null(col_fun)) {
        col_fun = colorRamp2(c(-3, 0, 3), c("purple", "black", "yellow"))
    }
    
    h <- Heatmap(
        name = "Expr",
        obj@scale.data[as.character(res$gene), as.character(rownames(meta))],
        cluster_rows = F,
        cluster_columns = F,
        show_column_names = F,
        show_row_names = F,
        col = col_fun,
        top_annotation = ta,
        left_annotation = la,
        right_annotation = ra,
        border = T,
        column_title_gp = gpar(fontsize=0),
        row_title_gp = gpar(fontsize = 0),
        column_split = meta$Stage,
        row_split = res$Clt,
        heatmap_legend_param = list(
            direction = "horizontal",
            labels_gp = gpar(fontsize = labels_size),
            title_gp = gpar(fontsize = title_size)
        )
    )
    return(h)
}



find_markers <- function(obj, n.cores = 10) {
    registerDoMC(n.cores)
    res = foreach (i = unique(obj@meta.data$ident), .combine = "rbind", .errorhandling = "pass") %dopar% {
        ident.1 = rownames(obj@meta.data)[obj@meta.data$ident == i]
        ident.2 = rownames(obj@meta.data)[obj@meta.data$ident != i]
        
        if (length(ident.1) > 3 && length(ident.2) > 3) {
            temp = FindMarkers(obj, ident.1, ident.2, logfc.threshold = 0)
            temp$ident = i
            temp$gene = rownames(temp)
            return(temp)
        }
        return(NULL)
    }
    
    return (res)
}


calculate_DEGs_on_random_select <- function(meta_data, scale_data, raw_data, seeds = 1:0, n = 15, n.cores = 10, avg_logFC = 0.25, p_adj_val = 0.05, pct.1 = 0.5) {
    registerDoMC(n.cores)
    res = foreach (s = seeds, .combine = "rbind") %dopar% {
        set.seed(s)
        meta_data$Cells = rownames(meta_data)

        sel_cells = meta_data %>%
            # filter(PatientID != "PA11") %>%
            group_by(Disease, Stage) %>%
            sample_n(n, replace = F) %>%
            as.data.frame()

        temp_meta = meta_data[as.character(sel_cells$Cells),]
        temp_meta$ident = paste(temp_meta$Disease, temp_meta$Stage, sep = "-")
        temp_obj <- CreateSeuratObject(raw_data[, as.character(sel_cells$Cells)], meta.data = temp_meta)

        temp_obj@scale.data <- scale_data[, as.character(sel_cells$Cells)]
        temp_obj@data <- temp_obj@scale.data
        new_markers <- NULL
        for(i in unique(temp_meta$ident)) {
            ident.1 = rownames(temp_obj@meta.data)[temp_obj@meta.data$ident == i]
            ident.2 = rownames(temp_obj@meta.data)[temp_obj@meta.data$ident != i]

            temp = FindMarkers(temp_obj, ident.1 = ident.1, ident.2 = ident.2)
            temp$gene = rownames(temp)
            temp$ident = i
            new_markers = rbind(new_markers, temp)
        }
        new_markers = new_markers[new_markers$avg_logFC > avg_logFC & new_markers$p_val_adj < p_adj_val & new_markers$pct.1 > pct.1, c("ident", "gene", "avg_logFC")]
        return(new_markers)
    }

    res <- res %>%
        group_by(ident, gene) %>%
        mutate(avg_logFC = mean(avg_logFC)) %>%
        group_by(ident, gene) %>%
        add_tally() %>%
        mutate(freq = n / length(seeds)) %>%
        dplyr::select(ident, gene, freq, avg_logFC) %>%
        group_by(gene) %>%
        add_tally() %>%
        as.data.frame()

    set.seed(1)
    return(res)
}


get_seurat_obj <- function(expr, meta, cell) {
    obj <- CreateSeuratObject(
        expr[, rownames(meta)[as.character(meta$cell_short) == cell]],
        meta.data = meta
    )
    
    mito.genes <- grep(pattern = "^MT-", x = rownames(x = obj@raw.data), value = TRUE)
    percent.mito <- Matrix::colSums(obj@raw.data[mito.genes, ])/Matrix::colSums(obj@raw.data)
    
    # AddMetaData adds columns to object@meta.data, and is a great place to
    # stash QC stats
    obj <- AddMetaData(object = obj, metadata = percent.mito, col.name = "percent.mito")
     
    obj <- NormalizeData(object = obj, normalization.method = "LogNormalize", scale.factor = 10000)
    obj <- FindVariableGenes(object = obj, mean.function = ExpMean, dispersion.function = LogVMR, x.low.cutoff = 0.0125, x.high.cutoff = 3, y.cutoff = 0.5)
    obj <- ScaleData(object = obj, vars.to.regress = c("nUMI", "percent.mito"))
    
    obj
}
```


### Format 
```{r}
meta <- read.csv(paste0(root.dir, "../../02_rds/meta_after_singleR.csv"), row.names = 1, stringsAsFactors = F)

nm_meta = read.xlsx(paste0(root.dir, "../../02_rds/NM_meta.xlsx"))
rownames(nm_meta) <- nm_meta$SampleID

meta$Disease[meta$SampleID %in% nm_meta$SampleID] = nm_meta[meta$SampleID[meta$SampleID %in% nm_meta$SampleID], "Disease"]
meta$Stage[meta$SampleID %in% nm_meta$SampleID] = nm_meta[meta$SampleID[meta$SampleID %in% nm_meta$SampleID], "Stage"]

expr <- readRDS(paste0(root.dir, "../../02_rds/all_cell_expr.rds"))
```

```{r}
atii <- get_seurat_obj(expr, meta, "ATII")
basal <- get_seurat_obj(expr, meta, "Basal")
```

```{R}
saveRDS(atii, paste0(root.dir, "all_atii_cells.rds"))
saveRDS(basal, paste0(root.dir, "all_basal_cells.rds"))
```

## Calculate gene moudle of ATII
```{r}
temp_meta = atii@meta.data
temp_meta$ident = paste(temp_meta$Disease, temp_meta$Stage, sep = "-")
res_markers <- calculate_DEGs_on_random_select(
    temp_meta, 
    atii@scale.data,
    atii@raw.data, 
    seeds = 1:200, 
    pct.1 = 0.25, 
    p_adj_val = 0.05, 
    avg_logFC = 0.25, 
    n.cores = 20
)

saveRDS(res_markers, paste0(root.dir, "atii_all_markers1.rds"))
```


```{r eval=FALSE, fig.height=8, fig.width=10, include=FALSE}
atii <- readRDS(paste0(root.dir, "all_atii_cells.rds"))
set.seed(42)
meta_data = atii@meta.data
meta_data$Cells = rownames(meta_data)

sel_cells = meta_data[meta_data$PatientID != "PA11", ] %>%
    group_by(Disease, Stage) %>%
    sample_n(15, replace = F) %>%
    as.data.frame()
```


```{r fig.height=8, fig.width=10}
res_markers <- readRDS(paste0(root.dir, "atii_all_markers.rds"))

atii@meta.data$ident = paste(atii@meta.data$Disease, atii@meta.data$Stage, sep = "-")
# res_markers <- find_markers(atii, n.cores = 20)

cluster_markers <- read.csv(paste0(root.dir, "ATII_all_cells_cluster_markers.csv"), row.names = 1, stringsAsFactors = F)

temp <- cluster_markers %>%
    filter(p_val_adj < 0.05) %>%
    group_by(ident) %>%
    top_n(10, wt=avg_logFC)

temp_markers = unique(res_markers[, 1:2])
temp_markers = temp_markers[order(temp_markers[,1], temp_markers[,2]), ]
temp_markers = temp_markers[str_detect(temp_markers$ident, "^LUAD-"), ]
temp_markers = temp_markers[temp_markers$gene %in% names(table(temp_markers$gene))[table(temp_markers$gene) == 1], ]

mfuzz <- read.csv(paste0(root.dir, "ATII_mfuzz.csv"), row.names = 1, stringsAsFactors = F)

filtered_markers = NULL
for (i in unique(temp_markers$ident)) {
    j = paste0("M.", str_split(i, "-")[[1]][2])
    
    temp = temp_markers[temp_markers$ident == i, ]
    
    if (j %in% unique(mfuzz$Clt)) {
        temp = temp[temp$gene %in% mfuzz$gene[mfuzz$Clt == j], ]
    }
    
    filtered_markers <- rbind(filtered_markers, temp)
}


filtered_markers <- rbind(
    filtered_markers, 
    temp_markers[temp_markers$ident == "LUAD-II", ]
)


pdf(paste0(root.dir, "ATII_markers.pdf"), width = 10, height = 8)
make_stage_module_heatmap(
    as.matrix(atii@scale.data[, as.character(sel_cells$Cells)]), 
    atii@meta.data[as.character(sel_cells$Cells), ], 
    filtered_markers$gene,
    order.by.patient = F,
    col_fun = colorRamp2(c(-1, 0, 1), c("purple", "black", "yellow")),
    mark_genes = mfuzz$gene,
    cluster_row = F,
    row_split = filtered_markers$ident
)

dev.off()
```


### Test on sample
```{r fig.height=8, fig.width=10}
test_on_sampled_cells <- function(data, meta, markers, seeds = 1:100) {
    
    for(i in seeds) {
        set.seed(i)
        meta_data$Cells = rownames(meta_data)
        
        sel_cells = meta_data %>%
            filter(PatientID != "PA11") %>%
            group_by(Disease, Stage) %>%
            sample_n(15, replace = F) %>%
            as.data.frame()
        
        make_stage_module_heatmap(
            as.matrix(data[, as.character(sel_cells$Cells)]), 
            meta[as.character(sel_cells$Cells), ], 
            markers$gene,
            order.by.patient = F,
            col_fun = colorRamp2(c(-4, 0, 4), c("blue", "white", "red")),
            # mark_genes = new_markers$gene[new_markers$avg_logFC > 0.25 & new_markers$p_val_adj < 0.05 & new_markers$ident == "LUAD-I"],
            cluster_row = F,
            row_split = markers$ident,
            column_title = paste("Seed:", i)
        )
    }
}

test_on_sampled_cells(atii@scale.data, meta_data, res_markers, seeds = 2:20)
```



## Similirity between gene expression of all ATII and all Basal
```{r}
atii <- readRDS(paste0(root.dir, "all_atii_cells.rds"))
basal <- readRDS(paste0(root.dir, "all_basal_cells.rds"))


expr <- cbind(atii@raw.data, basal@raw.data)
meta <- rbind(
    atii@meta.data[intersect(colnames(atii@meta.data), colnames(basal@meta.data))], 
    basal@meta.data[intersect(colnames(atii@meta.data), colnames(basal@meta.data))]
)

saveRDS(expr, paste0(root.dir, "all_cells_expr.rds"))
saveRDS(meta, paste0(root.dir, "all_cells_meta.rds"))
```


## SCTransform (Seurat3)
```{R}
library(Seurat)

expr <- readRDS(paste0(root.dir, "all_cells_expr.rds"))
meta <- readRDS(paste0(root.dir, "all_cells_meta.rds"))

obj <- CreateSeuratObject(counts = expr, project = "pbmc3k", min.cells = 3, min.features = 200, meta.data = meta)

obj[["percent.mt"]] <- PercentageFeatureSet(obj, pattern = "^MT-")
obj <- SCTransform(obj, vars.to.regress = "percent.mt", verbose = FALSE)

saveRDS(obj@assays$SCT@scale.data, paste0(root.dir, "all_cells_sctransform.rds"))
```


### format seurat obj data into 
```{r}
format_expr <- function(data, meta, group.by = c("PatientID", "Disease", "Stage", "Batch")) {
    
    expr <- as.matrix(data)
    colnames(expr) <- sapply(as.character(colnames(expr)), function(x) {
        paste(meta[x, group.by], collapse = "-")
    })
    expr <- melt(expr)
    expr
}
```


```{r}
make_corr_heatmap <- function(data) {
    colors = list(
        Stage = c(
            "I"="#65A9A3", 
            "II"="#4A933E", 
            "III"="#EC7A21", 
            "IV"="#D73F47", 
            "LUAD_Normal"="#FECC1B", 
            "LUSC_Normal"="#778793"
        ),
        Disease = c(
            "AD" = "#0084D1", 
            "NL(AD)"="#FECC1B", 
            "NL(AD"="#778793", 
            "Normal"="#73BDFF", 
            "NL(SC)"="#778793", 
            "SC"="#A0C807",
            "LC"="#91822B",
            "NL(LC)"="#EDCAB0",
            "UPS"="#EBAAA4",
            "NL(UPS)"="#B43018"
        ),
        Batch = c("1"="#E9B0B7", "2"="#90D0E0", "3"="#769982"),
        Cell = c(
            "ATII"="#FF0000",
            "Basal"="#F2AD00",
            "CD4"="#F98400",
            "CD8"="#5BBCD6",
            "Cilia"="#85D4E3",
            "Club"="#F4B5BD",
            "DC"="#9C964A",
            "EC"="#CDC08C",
            "Epi"="#FAD77B",
            "Fib"="#ECCBAE",
            "Mast"="#D69C4E",
            "Mo"="#ABDDDE",
            "NK"="#000000",
            "NE"="#F3DF6C",
            "Tregs"="#CEAB07",
            "Mφ"="#046C9A"
        )
    )
    
    meta = data.frame(
        Cell = sapply(colnames(data), function(x) { str_split(x, "-")[[1]][1] }),
        Disease = sapply(colnames(data), function(x) { 
            x = str_split(x, "-")[[1]][3] 
            disease = c(
                "LUAD"="AD",
                "LUSC"="SC",
                "LUAD_Normal"="NL(AD)",
                "LUSC_Normal"="NL(SC)",
                "LULC"="LC",
                "LULC_Normal"="NL(LC)",
                "Pleiomorphic"="UPS",
                "Pleiomorphic Normal"="NL(UPS)"
            )
        
            as.character(disease[x])
        }),
        Stage = sapply(colnames(data), function(x) { str_split(x, "-")[[1]][4] }),
        Batch = sapply(colnames(data), function(x) { str_split(x, "-")[[1]][5] })
    )
    
    ba <- HeatmapAnnotation(
        df = meta,
        col = colors
    )
    
    
    ra <- rowAnnotation(
        df = meta,
        col = colors,
        show_annotation_name = F,
        show_legend = F
    )
    
    Heatmap(
        data,
        name = "Corr",
        col = colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
        show_column_names = F,
        show_row_names = F,
        bottom_annotation = ba,
        right_annotation = ra,
        clustering_method_rows = "ward.D",
        clustering_method_columns = "ward.D"
    )
}
```



```{r fig.height=10, fig.width=12}
expr <- readRDS(paste0(root.dir, "all_cells_sctransform.rds"))
```

```{r fig.height=4, fig.width=6}
expr1 <- format_expr(expr, meta, group.by = c("cell_short", "PatientID", "Disease", "Stage", "Batch"))

expr1 <- dcast(expr1, Var1~Var2, value.var = "value", fun.aggregate = mean)

rownames(expr1) <- expr1$Var1
expr1 <- expr1[, colnames(expr1) != "Var1"]

expr_corr <- cor(expr1)

pdf(paste0(root.dir, "ATII_Basal_gene_corr_pearson_all_genes.pdf"), width = 6, height = 4)
make_corr_heatmap(expr_corr)
dev.off()
```

```{r}
pdf(paste0(root.dir, "ATII_Basal_gene_corr_spearman_all_genes.pdf"), width = 6, height = 4)
expr_corr <- cor(expr1, method = "spearman")
make_corr_heatmap(expr_corr)
dev.off()
```



```{r fig.height=10, fig.width=12}
sds = apply(expr, 1, sd)
expr1 <- format_expr(expr[names(sds)[sds > summary(sds)[3]], ], meta, group.by = c("cell_short", "PatientID", "Disease", "Stage", "Batch"))

expr1 <- dcast(expr1, Var1~Var2, value.var = "value", fun.aggregate = mean)
rownames(expr1) <- expr1$Var1
expr1 <- expr1[, colnames(expr1) != "Var1"]

pdf(paste0(root.dir, "ATII_Basal_gene_corr_pearson.pdf"), width = 6, height = 4)
expr_corr <- cor(expr1)
make_corr_heatmap(expr_corr)
dev.off()
```


```{r fig.height=10, fig.width=12}
pdf(paste0(root.dir, "ATII_Basal_gene_corr_spearman.pdf"), width = 6, height = 4)
expr_corr <- cor(expr1, method = "spearman")
make_corr_heatmap(expr_corr)
dev.off()
```


```{r}
# read expression data
atii <- readRDS(paste0(root.dir, "ATII.rds"))
basal <- readRDS(paste0(root.dir, "Basal.rds"))
```



## Plot mfuzz markers
```{r fig.height=6, fig.width=8}

all_markers <- readRDS(paste0(root.dir, "atii_all_markers.rds"))
mfuzz <- read.csv(paste0(root.dir, "ATII_mfuzz.csv"), row.names = 1, stringsAsFactors = F)
cluster_markers <- read.csv(paste0(root.dir, "ATII_all_cells_cluster_markers.csv"), row.names = 1, stringsAsFactors = F)

temp <- cluster_markers %>%
    filter(p_val_adj < 0.05) %>%
    group_by(ident) %>%
    top_n(10, wt=avg_logFC)

gene_mark = c(
    "CYP2B7P", "MUC1",
    "CKS1B", "H2AFZ", 
    "LYZ", "SRGN",
    "FABP5", "SFTPC", "SFTPD",
    mfuzz$gene[mfuzz$gene %in% temp$gene]
)

set.seed(1)
temp <- mfuzz %>%
    group_by(Clt) %>%
    top_n(5)


gene_mark = unique(c(gene_mark, temp$gene))


pdf(paste0(root.dir, "ATII_mfuzz.pdf"), height = 6, width = 8)
ComplexHeatmap::draw(
    make_module_heatmap(
        atii, 
        mfuzz, 
        col_fun = colorRamp2(c(-2, 0, 2), c("purple", "black", "yellow")),
        order_by_patient = F,
        gene_mark = gene_mark
    ), 
    column_title = "",
    column_title_gp = gpar(fontsize = 20),
    merge_legend = TRUE, 
    heatmap_legend_side = "bottom", 
    annotation_legend_side = "bottom"
)
dev.off()
```


```{r fig.height=6, fig.width=8}
cluster_markers <- read.csv(paste0(root.dir, "Basal_all_cells_cluster_markers.csv"), row.names = 1, stringsAsFactors = F)
markers <- read.csv(paste0(root.dir, "Basal_mfuzz.csv"), row.names = 1, stringsAsFactors = F)

temp <- cluster_markers %>%
    filter(p_val_adj < 0.05) %>%
    group_by(ident) %>%
    top_n(10, wt=avg_logFC)

gene_mark = c(
    "CYP2B7P", "MUC1",
    "CKS1B", "H2AFZ", 
    "LYZ", "SRGN",
    "FABP5", "SFTPC", "SFTPD",
    markers$gene[markers$gene %in% temp$gene]
)

set.seed(1)
temp <- markers %>%
    group_by(Clt) %>%
    top_n(5)

gene_mark = unique(c(gene_mark, temp$gene))


pdf(paste0(root.dir, "Basal_mfuzz.pdf"), height = 6, width = 8)
ComplexHeatmap::draw(
    make_module_heatmap(
        basal, 
        markers[order(as.character(markers$Clt)), ], 
        col_fun = colorRamp2(c(-2, 0, 2), c("purple", "black", "yellow")),
        order_by_patient = F,
        gene_mark = gene_mark
    ), 
    column_title = "",
    column_title_gp = gpar(fontsize = 20),
    merge_legend = TRUE, 
    heatmap_legend_side = "bottom", 
    annotation_legend_side = "bottom"
)
dev.off()
```

```{r fig.height=8, fig.width=10}
make_stage_module_heatmap(
    as.matrix(sct_data[, rownames(meta_data)[meta_data$PatientID != "PA11"]]), 
    meta_data[meta_data$PatientID != "PA11", ], 
    unique(new_markers$gene[new_markers$avg_logFC > 0.25 & new_markers$p_val_adj < 0.05]),
    order.by.patient = F,
    col_fun = colorRamp2(c(-3, 0, 1), c("blue", "white", "red")),
    # mark_genes = new_markers$gene[new_markers$avg_logFC > 0.25 & new_markers$p_val_adj < 0.05 & new_markers$ident == "LUAD-I"],
    cluster_row = F,
    row_split = new_markers$ident[new_markers$avg_logFC > 0.25 & new_markers$p_val_adj < 0.05]
)
```


## Cluster markers
```{r}
atii <- readRDS(paste0(root.dir, "total_atii.rds"))
basal <- readRDS(paste0(root.dir, "total_basal.rds"))
```


```{r}
calculate_cluster_perc <- function(obj, seed.range = 1:50) {
    meta = obj@meta.data

    num = table(meta$Stage)
    
    temp <- meta %>%
        group_by(ident, Stage) %>%
        add_tally() %>%
        dplyr::select(ident, Stage, n) %>%
        unique() %>%
        as.data.frame()
    
    temp$freq <- apply(temp, 1, function(row) {
        as.numeric(row[3]) / num[[row[2]]]
    })
    
    temp
}
```


```{r}
atii_perc <- calculate_cluster_perc(atii)

p <- ggplot(atii_perc, aes(x=factor(ident), y = freq, color = Stage)) + 
    geom_point() +
    geom_line(aes(x=ident, y = freq, color = Stage)) +
    scale_color_manual(values = c(
        "I"="#65A9A3", 
        "II"="#4A933E",
        "III"="#EC7A21",
        "IV"="#D73F47"
    )) +
    labs(x = "", y = "", color = "") +
    theme_bw() +
    theme(
        axis.text = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.position = c(0.9, 0.8)
    )
p
ggsave(
    filename = paste0(root.dir, "ATII_all_stage_violin.pdf"),
    plot = p,
    width = 3,
    height = 3
)
```

```{r}
atii_markers <- find_markers(obj, 20)

basal@meta.data$ident <- as.numeric(basal@meta.data$res.0.1) + 1
basal_markers <- find_markers(basal, 20)
```

```{r}
write.csv(atii_markers, paste0(root.dir, "ATII_all_cells_cluster_markers.csv"))
write.csv(basal_markers, paste0(root.dir, "Basal_all_cells_cluster_markers.csv"))
```

---

### Compare different markers
```{r}
all_markers <- readRDS(paste0(root.dir, "atii_all_markers.rds"))
mfuzz <- read.csv(paste0(root.dir, "ATII_mfuzz.csv"), row.names = 1, stringsAsFactors = F)
cluster_markers <- read.csv(paste0(root.dir, "ATII_all_cells_cluster_markers.csv"), row.names = 1, stringsAsFactors = F)
```


```{r fig.height=12, fig.width=8}
obj <- readRDS(paste0(root.dir, "total_atii.rds"))
obj@meta.data$ident = as.numeric(obj@meta.data$res.0.3) + 1

temp <- cluster_markers %>%
    filter(p_val_adj < 0.05) %>%
    group_by(ident) %>%
    top_n(10, wt=avg_logFC)


cells_order = rownames(obj@meta.data)[order(as.numeric(obj@meta.data$ident))]

ccls <- cluster_cols[1:length(unique(obj@meta.data$ident))]
names(ccls) <- unique(obj@meta.data$ident)

ta = HeatmapAnnotation(
    Cluster = obj@meta.data$ident[order(as.numeric(obj@meta.data$ident))],
    col = list(
        Cluster = ccls
    ),
    show_annotation_name = F,
    show_legend = F
)


# pdf(paste0(root.dir, "ATII_cluster_markers_heatmap.pdf"), width = 6, height = 8)
Heatmap(
    name = "Expr",
    obj@scale.data[temp$gene, cells_order],
    cluster_rows = F,
    cluster_columns = F,
    show_row_names = T,
    show_column_names = F,
    top_annotation = ta,
    column_split = obj@meta.data$ident[order(as.numeric(obj@meta.data$ident))],
    row_split = temp$ident,
    heatmap_legend_param = list(
        labels_gp = gpar(fontsize = 15),
        title_gp = gpar(fontsize = 15)
    ),
    col = colorRamp2(c(-2, 0, 2), c("purple", "black", "yellow"))
)
# dev.off()
```


```{r}
sort(temp$gene[temp$gene %in% all_markers$gene])
```

```{r}
sort(temp$gene[temp$gene %in% mfuzz$gene])
```

```{r fig.height=8, fig.width=12}
lt = list()

# format all markers
for (i in unique(all_markers$ident[str_detect(all_markers$ident, "^LUAD-")])) {
    lt[[i]] <- as.character(all_markers$gene[all_markers$ident == i])
}

# format mfuzz
for (i in unique(mfuzz$Clt)) {
    lt[[i]] <- as.character(mfuzz$gene[mfuzz$Clt == i])
}

# format cluster markers
for (i in c(1, 2, 3, 4)) {
    lt[[paste("Cluster", i)]] <- as.character(cluster_markers$gene[cluster_markers$p_val_adj < 0.05 & cluster_markers$avg_logFC > 0.25 & cluster_markers$ident == i])
}


# "I"="#65A9A3", 
# "II"="#4A933E", 
# "III"="#EC7A21", 
# "IV"="#D73F47",


library(UpSetR)
upset(
    fromList(lt), 
    order.by = "freq", 
    sets = names(lt),
    queries = list(
        # Cluster 1 and Stage I
        list(query = intersects,params = list("Cluster 1", "LUAD-I"),color = "#65A9A3",active = T,query.name = "C1 I"),
        list(query = intersects,params = list("Cluster 1", "M.I"),color = "#65A9A3",active = T,query.name = "C1 MI"),
        list(query = intersects,params = list("Cluster 1", "M.I", "LUAD-I"),color = "#65A9A3",active = T,query.name = "C1 MI I"),

        # Cluster 2 and Stage III
        list(query = intersects,params = list("Cluster 2", "LUAD-III"),color = "#4A933E",active = T,query.name = "C2 II"),
        list(query = intersects, params = list("Cluster 2", "M.II"),color = "#4A933E",active = T,query.name = "C2 MII"),

        # Cluster 3 and Stage III, IV
        list(query = intersects,params = list("Cluster 3", "LUAD-III"),color = "#EC7A21", active = T,query.name = "C3 III"),
        list(query = intersects,params = list("Cluster 3", "M.IV"),color = "#D73F47",active = T,query.name = "C3 IV"),

        # Cluster 5 and Stage III
        list(query = intersects,params = list("Cluster 5", "Cluster 3"),color = "#EC7A21",active = T,query.name = "C3 C5"),
        list(query = intersects,params = list("Cluster 5", "Cluster 3", "LUAD-III"),color = "#EC7A21",active = T,query.name = "C3 C5 III"),
        list(query = intersects,params = list("Cluster 5", "LUAD-III"),color = "#EC7A21",active = T,query.name = "C5 III")
    )
)
```

## Seletect ATII markers
```{r}
stageI = Reduce(intersect, list(
    as.character(cluster_markers$gene[cluster_markers$p_val_adj < 0.05 & cluster_markers$avg_logFC > 0.25 & cluster_markers$ident == 1]),
     mfuzz$gene[mfuzz$Clt == "M.I"]
))

stageII = Reduce(intersect, list(
    as.character(cluster_markers$gene[cluster_markers$p_val_adj < 0.05 & cluster_markers$avg_logFC > 0.25 & cluster_markers$ident == 2]),
    mfuzz$gene[mfuzz$Clt == "M.II"]
))

stageIII = Reduce(intersect, list(
    as.character(cluster_markers$gene[cluster_markers$p_val_adj < 0.05 & cluster_markers$avg_logFC > 0.25 & cluster_markers$ident == 3]),
    as.character(cluster_markers$gene[cluster_markers$p_val_adj < 0.05 & cluster_markers$avg_logFC > 0.25 & cluster_markers$ident == 5]),
    all_markers$gene[all_markers$ident == "LUAD-III"]
))

stageIII = Reduce(intersect, list(
    as.character(cluster_markers$gene[cluster_markers$p_val_adj < 0.05 & cluster_markers$avg_logFC > 0.25 & cluster_markers$ident == 3]),
    as.character(cluster_markers$gene[cluster_markers$p_val_adj < 0.05 & cluster_markers$avg_logFC > 0.25 & cluster_markers$ident == 5]),
    all_markers$gene[all_markers$ident == "LUAD-III"]
))

stageIV = c(
    intersect(
        as.character(cluster_markers$gene[cluster_markers$p_val_adj < 0.05 & cluster_markers$avg_logFC > 0.25 & cluster_markers$ident == 3]), 
        all_markers$gene[all_markers$ident == "LUAD-IV"]
    ),
    intersect(
        as.character(cluster_markers$gene[cluster_markers$p_val_adj < 0.05 & cluster_markers$avg_logFC > 0.25 & cluster_markers$ident == 3]),
        mfuzz$gene[mfuzz$Clt == "M.IV"]
    )
)

markers <- data.frame(
    ident = c(rep("I", length(stageI)), rep("II", length(stageII)), rep("III", length(stageIII)), rep("IV", length(stageIV))),
    gene = c(stageI, stageII, stageIII, stageIV)
)


make_stage_module_heatmap(
    as.matrix(atii@scale.data[, as.character(sel_cells$Cells)]), 
    atii@meta.data[as.character(sel_cells$Cells), ], 
    markers$gene,
    order.by.patient = F,
    col_fun = colorRamp2(c(-1, 0, 1), c("purple", "black", "yellow")),
    # mark_genes = c("SCGB3A2"),
    cluster_row = F,
    row_split = markers$ident
)
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.