---
title: "Basal"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
root.dir = "LungCancer10x/"

full.path <- function(...) { return(paste(root.dir, ..., sep = "/")) }
```


```{r eval=FALSE, include=FALSE}
library(Seurat)
library(here)
library(scatterpie)
library(dplyr)
library(ggthemes)
library(ggrastr)
library(ComplexHeatmap)
library(circlize)
library(wesanderson)
library(openxlsx)
library(stringr)
# library(Mfuzz)
library(reticulate)
library(ggradar)
library(reshape2)
library(ggrepel)
library(ggpubr)
library(ggthemes)
library(tmod)

use_python("program/pyenv/shims/python")
```



```{R}
make_feature_plot <- function(
    obj,
    gene, 
    coord=NULL, 
    pt.size = 0.1, 
    alpha = 0.5, 
    sort = T,
    reduction.use = "umap"
) {
    library(ggplot2)
    library(ggrastr)
    data = as.data.frame(obj@meta.data)
    data$value = obj@data[which(rownames(obj@data) == gene), rownames(data)]
    
    if(is.null(coord)) {
        coord = obj@dr[[reduction.use]]@cell.embeddings
    }
    
    data$x = coord[rownames(data), 1]
    data$y = coord[rownames(data), 2]
    
    if(sort) {
        data = data[order(data$value, decreasing = F), ]
    }
    
    p <- ggplot(data = data, aes(x=x, y=y, color = value)) +
        geom_point_rast(size = pt.size, alpha = alpha) +
        scale_color_gradient(low = "lightgrey", high = "blue") +
        theme_bw() +
        labs(title = gene, x="UMAP1", y="UMAP2") +
        theme(
            plot.title = element_text(hjust = 0.5, face="bold", size = 20), 
            legend.position = "none",
            axis.text = element_text(size = 12),
            axis.title = element_text(size = 15)
        )
    p
}
```



```{R}
# obj <- readRDS(full.path("03_each_cells/Batch/Basal/LUSC/seurat.rds"))
obj <- readRDS(full.path("11_CNV/each_cells/Basal/LUSC/seurat_obj.rds"))
urd <- readRDS(full.path("11_CNV/each_cells/Basal/LUSC/URD.rds"))
```


```{r}
genes = c(
    "KRT17", "KRT5", "MIR205HG", # "DLK2", "POSTN", "ISLR",
     "SNCA", "PCDH7",
    "ADH7", "KRT7", # "SCGB3A2",
    "HES1", "SERPINB3", "CLCA2"
)
```


KRT17	General
KRT5	General
MIR205HG	General
DLK2	General
POSTN	Bas_px
ISLR	Bas_px
SNCA	Bas_px
PCDH7	Bas_px
ADH7	Bas_px
ALOX15	Bas_px
MKI67	Bas_p
TOP2A	Bas_p
PBK	Bas_p
GTSE1	Bas_p
KRT7	Bas_d
SCGB3A2	Bas_d
HES1	Bas_d
SERPINB3	Bas_d
CLCA2	Bas_d

```{R fig.height=4, fig.width=4}
for (i in genes) {
    p <- make_feature_plot(obj, i)
    print(p)
}
```


```{r}
library(reticulate)
tryCatch({
    kee = import("kneed")
    n.pcs = kee$KneeLocator(
        1:ncol(obj@dr$pca@cell.embeddings),
        apply(obj@dr$pca@cell.embeddings, 2, sd),
        curve='convex',
        direction='decreasing'
    )
    n.pcs = n.pcs$knee
}, error = function(e) {
    print(e)
    n.pcs = 10
})

n.pcs = 14

for (i in seq(0.04, 0.09, by=0.01)) {
    obj <- FindClusters(
        obj, resolution = i, 
        force.recalc = T, reduction.type = "harmony", 
        print.output = F, dims.use = 1:n.pcs
    )
}

# saveRDS(obj, here("seurat.rds"))
```




```{r fig.height=4, fig.width=5}
for (i in c(0.03, 0.05, seq(0.1, 0.5, by = 0.1))) {
    p <- DimPlot(obj, reduction.use = "umap", group.by = paste0("res.", i), do.return=T, pt.size = 0.1)
    p = p + labs(title = i)
    print(p)
}

```


```{r fig.height=4, fig.width=6}
DimPlot(obj, reduction.use = "umap", group.by = "Disease", do.return=T, pt.size = 0.1)
```


```{R}
markers = read.csv(full.path("11_CNV/each_cells/Basal/LUSC/cluster_markers.csv"), row.names = 1, stringsAsFactors = F)

markers <- markers %>%
    group_by(ident) %>%
    top_n(2, wt = avg_logFC) %>%
    as.data.frame()
```



```{R fig.height=8, fig.width=5}
obj@meta.data$ident1 = as.numeric(obj@meta.data$res.0.2) + 1


generic = c("KRT17", "KRT5", "MIR205HG")
bas_px = c("SNCA", "PCDH7", "ADH7")
bas_p = c("MKI67", "TOP2A", "PBK", "GTSE1")
bas_d = c("KRT7", "SERPINB3")

# genes = c(
#     , # "DLK2", "POSTN", "ISLR",
#      
#     , ", # "SCGB3A2",
#     "HES1", , "CLCA2"
# )

mark_genes = markers$gene[markers$ident %in% 1:3]
mark_genes = c(mark_genes, markers$gene[markers$ident == 4], "KPNA2", bas_p) # "ALOX15",  
mark_genes = c(mark_genes, markers$gene[markers$ident == 5], c("AQP3"))
mark_genes = c(mark_genes, markers$gene[markers$ident > 5], c("SPHK1", "PPT1", "KPNA2"))
mark_genes = c(
    mark_genes, c("PRKDC", "ATM", "ATR", "TP53", "MDM2")
)

cols = as.character(wes_palette("Darjeeling1"))

genes = unique(c(mark_genes, generic, bas_px, bas_d))
a = rep("black", length(genes))

a[genes %in% generic] = cols[2]
a[genes %in% bas_px] = cols[4]
a[genes %in% bas_p] = cols[1]
a[genes %in% bas_d] = cols[5]

p <- DotPlot(
    obj, 
    genes.plot = genes, 
    do.return = T, x.lab.rot = T, 
    group.by = "ident1"
) + coord_flip() +
    theme_bw(base_family = "Arial Unicode MS") +
    theme(
        legend.position = "right",
        panel.grid = element_blank(),
        axis.text = element_text(size = 12),
        axis.text.y = element_text(colour = rev(a)),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 15)
    ) +
    labs(x = "", y = "")

p

ggsave(
    filename = full.path("11_CNV/each_cells/Basal/LUSC/cluster_stem_dotplot.pdf"),
    plot = p, width = 5, height = 8, device = cairo_pdf
)
```



```{r}
meta = readRDS(full.path("11_CNV/meta.rds"))
obj@meta.data$ident1 = as.numeric(obj@meta.data$res.0.2) + 1
obj@meta.data$Malignant = meta[rownames(obj@meta.data), "Malignant_tree"]

temp = obj@meta.data %>%
    group_by(ident1, Malignant) %>%
    add_tally() %>%
    dplyr::select(ident1, Malignant, n) %>%
    unique() %>%
    group_by(ident1) %>%
    mutate(p = n / sum(n)) %>%
    as.data.frame()


ggplot(temp, aes(x=ident1, y=p, fill=Malignant)) +
    geom_bar(stat = "identity")
```


```{r}
temp = obj@meta.data %>%
    group_by(ident1, Malignant) %>%
    add_tally() %>%
    dplyr::select(ident1, Malignant, n) %>%
    unique() %>%
    group_by(Malignant) %>%
    mutate(p = n / sum(n)) %>%
    as.data.frame()

ggplot(temp, aes(x=Malignant, y=p, fill=as.character(ident1))) +
    geom_bar(stat = "identity") +
    labs(x="Malignant subtypes", fill="Basal clusters", y="") +ﬁ
    theme_bw() +
    theme(
        axis.text = element_text(size = 15),
        legend.text = element_text(size = 15),
        axis.title = element_text(size = 15)
    )
```


```{r fig.height=4, fig.width=4}
temp = as.data.frame(urd@dm@eigenvectors)
obj@meta.data$ident1 = as.numeric(obj@meta.data$res.0.2) + 1

temp$ident = obj@meta.data[rownames(temp), "ident1"]
temp = temp[!is.na(temp$ident), ]


# pdf(full.path("11_CNV/each_cells/Basal/LUSC/DM_clt.pdf"), width = 6, height = 4)
ggplot(temp, aes(x=DC1, y=DC2, color=as.character(ident))) +
    geom_point_rast(size = 0.1) +
    theme_bw() +
    theme(
        axis.text = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        axis.title = element_text(size = 15),
        panel.grid = element_blank(),
        legend.position = c(0.1, 0.3)
    )+
    guides(color = guide_legend(override.aes = list(size = 5)))
# dev.off()
```


```{r}
set.seed(42)
temp_meta = obj@meta.data
temp_meta$Cells = rownames(temp_meta)
temp_meta <- temp_meta %>%
    group_by(res.0.03) %>%
    sample_n(min(table(obj@meta.data$res.0.2))) %>%
    as.data.frame()


col_fun = colorRamp2(c(-2, 0, 2), c("purple", "black", "yellow"))

Heatmap(
    obj@scale.data[genes, temp_meta$Cells],
    name = "Expr",
    col = col_fun,
    show_row_names = T,
    show_column_names = F,
    cluster_columns = F,
    cluster_rows = F,
    column_split = temp_meta$res.0.2
)
```

```{r fig.height=4, fig.width=4}
library(ggrastr)
stemness = c("MKI67", "TOP2A", "PBK", "GTSE1")

mtx = apply(obj@raw.data, 2, function(col) { col * 10^6 / sum(col) })

mtx <- mtx[stemness, ]
mtx <- na.omit(mtx)

mtx <- melt(as.matrix(mtx))
mtx$Clt <- obj@meta.data[mtx$Var2, "res.0.2"]
mtx <- na.omit(mtx)


mtx <- mtx %>%
    group_by(Var2) %>%
    mutate(value = mean(value)) %>%
    dplyr::select(Var2, value, Clt) %>%
    unique() %>%
    as.data.frame()

for (i in colnames(obj@dr$umap@cell.embeddings)) {
    mtx[, i] <- obj@dr$umap@cell.embeddings[mtx$Var2, i]
}

mtx = mtx[order(mtx$value), ]


p <- ggplot(mtx, aes(x=UMAP1, y=UMAP2, color=log2(value + 1))) +
    geom_point_rast(size = 0.1, alpha = 0.5) +
    scale_color_gradient2(low = "white", high = "blue", mid = "grey") +
    theme_bw(base_family = "Arial Unicode MS") +
    theme(
        panel.grid = element_blank(),
        legend.position = c(0.25, 0.25),
        legend.background = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 10)
    ) +
    guides(color = guide_colorbar(
        title.position = "bottom"
    ))

ggsave(
    filename = full.path("11_CNV/each_cells/Basal/LUSC/stem_umap.pdf"),
    plot = p, width = 4, height = 4, device = cairo_pdf
)
```

```{r}
gg_color_hue <- function(n) {
    hues = seq(15, 375, length = n + 1)
    hcl(h = hues, l = 65, c = 100)[1:n]
}


stage_colors = c("I"="#65A9A3", "II"="#4A933E", "III"="#EC7A21", "IV"="#D73F47", "LUAD_Normal"="#FECC1B", "LUSC_Normal"="#778793")
disease_colors = c("LUAD" = "#0084D1", "LUAD_Normal"="#FECC1B", "Normal"="#73BDFF", "LUSC_Normal"="#778793", "LUSC"="#A0C807", "LELC"="#6A0019", "CPI"="#C5000B")
tissue_colors = c("Bronichi"="#FB290F", "Lymph node"="#488F17", "Tumor"="#FC8210")


cluster_cols = c(
    as.character(wes_palette("Zissou1")),
    as.character(wes_palette("Royal2")),
    as.character(wes_palette("Royal1")),
    as.character(wes_palette("Darjeeling2")),
    as.character(wes_palette("Darjeeling1"))
)

patient_colors = cluster_cols[1:length(unique(obj@meta.data$PatientID))]
names(patient_colors) <- unique(obj@meta.data$PatientID)

colors = c(stage_colors, disease_colors, tissue_colors, patient_colors)


labels_size = 12
title_size = 15
```




```{R}
make_scatterpie_plot <- function(
    object, 
    color="Disease", 
    alpha=0.8, 
    colors=colors, 
    pt.size = 0.1, 
    guide_ncol=1,
    text_size = 8,
    reduction.use = "tsne",
    r=0.05,
    legend.position="none",
    legend.direction="horizontal",
    legend.breaks=waiver(),
    custom_pos = NULL,
    legend.text = 15
) {
    meta = object@meta.data

    coord = object@dr[[reduction.use]]@cell.embeddings
    colnames(coord) = c("x", "y")

    meta = merge(meta, coord, by = 0)

    scatter = eval(
        parse(
            text=paste0(
                "meta %>% group_by(ident, ", color, ") %>% add_tally() %>% group_by(ident) %>% mutate(x1 = median(x), y1 = median(y), perc = n / sum(n)) %>% dplyr::select(ident, x1, y1, perc, ", color, ") %>% unique()"
            )
        )
    )

    colnames(scatter)[colnames(scatter) == color] = "group"

    scatter = dcast(scatter, ident + x1 + y1 ~ group, value.var="perc")

    scatter[is.na(scatter)] = 0
    scatter$r = min(max(meta$x) - min(meta$x), max(meta$y) - min(meta$y)) * r

    clt_color = gg_color_hue(length(unique(meta$ident)))
    names(clt_color) = 1:length(unique(meta$ident))
    
    if (!is.null(custom_pos)) {
        scatter$x1 = custom_pos[scatter$ident, "x1"]
        scatter$y1 = custom_pos[scatter$ident, "y1"]
    }

    p <- ggplot() + 
        geom_point_rast(
            aes(x=x, y=y, colour=factor(ident)), 
            data=meta, 
            alpha=alpha, 
            size=pt.size
        ) +
        geom_scatterpie(
            aes(
                x=x1,
                y=y1, 
                group=ident,
                r=r
            ),
            data=scatter,
            cols=as.character(unique(meta[, color])),
            color=NA
        ) +
        coord_equal() +
        geom_text(
            aes(x=x1, y=y1 - r - 0.5, label=ident),
            data=scatter,
            size = text_size
        ) +
        theme_bw() +
        theme(
            legend.position = legend.position,
            axis.text = element_text(size = 15),
            axis.title = element_text(size = 20),
            legend.text = element_text(size = legend.text),
            # legend.background = element_blank()
            legend.direction=legend.direction,
            legend.spacing.x = unit(0.1, 'cm')
        ) +
        scale_color_manual(values=clt_color, breaks = c()) + 
        scale_fill_manual(values=colors, breaks = legend.breaks) +
        guides(fill = guide_legend(ncol = guide_ncol))

    if (reduction.use == "tsne") {
        p <- p + labs(x="tSNE_1", y="tSNE_2", color="", fill = "")
    } else if (reduction.use == "umap") {
        p <- p + labs(x="UMAP1", y="UMAP2", color="", fill="")
    }

    return(p)
}

```



```{r fig.height=4, fig.width=12}
# Function to make combined bar plots
# 1. percentage plot by Stage
# 2. percentage plot by Disease
# 3. percentage plot by PatientID
# 4. barplot of number of cells
# 5. boxplots of number of transcripts
# Note: custom_colors needs to exists in namespace, cause this function using customed color palette
# :param object: Seurat object 2.3+
# :param group.by: how to make groups
# :param group.type: if int is set, sort the plot order by numeric, else by str
# :param rel_height: this function separate legend and main plot into two grid (up and down), so using this to adjust the relative position of legend and main plot
# :param with_disease: whethter plot the disease percentage
# :param ncol: the legend ncol of patient plot
# :return ggplot2 object
make_combined_barplot <- function(
    object, 
    group.by = "res.0.8", 
    group.type = "int", 
    rel_height = c(0.2, 1),
    with_disease=TRUE,
    ncol=4
) {
    library(scales)
    object@meta.data$PatientID = as.character(object@meta.data$PatientID)
    ## make percentage bar plot with stage
    meta <- object@meta.data
    meta$group <- as.character(meta[, group.by])
    # meta$newTissue <- as.character(meta$Tissue)
    # meta$newTissue[meta$newTissue == "Tumor" & meta$Disease != ""] <- as.character(meta$Disease[meta$newTissue == "Tumor" & meta$Disease != ""])

    plot_order = c(unique(meta$group), "Total")
    
    if(group.type == "int") {
        plot_order = as.character(c(sort(as.numeric(unique(meta$group))), "Total"))
    }
    
    ### make percentage plot of stage
    meta$Stage = as.character(meta$Stage)
    temp <- as.data.frame(
        meta %>% 
            dplyr::select(group, Stage) %>% 
            dplyr::group_by(group, Stage) %>% 
            dplyr::add_tally() %>% 
            unique() %>% 
            dplyr::group_by(group) %>% 
            dplyr::mutate(freq = n / sum(n)) %>% unique()
    )
    
    temp1 <- as.data.frame(table(meta$Stage))
    colnames(temp1) <- c("Stage", "n")
    
    temp1$group = "Total"
    temp1$freq = temp1$n / sum(as.numeric(temp1$n))
    
    temp1 <- temp1[, colnames(temp)]
    
    temp <- rbind(temp, temp1)
    
    temp$group <- factor(temp$group, levels = plot_order)
    
    p1 <- ggplot(data = temp, aes(x=group, y=freq, fill=Stage)) + 
        geom_bar(stat = "identity", width = .75) + 
        coord_flip() + 
        theme_bw() +
        theme(
            axis.text = element_text(size = 15),
            axis.title = element_text(size = 15),
            legend.text = element_text(size = 15),
            legend.title = element_blank(), 
            legend.justification = "center"
        ) +
        labs(y="Fraction of cells", x = "") +
        # scale_x_discrete(breaks=plot_order) + 
        guides(fill = guide_legend(ncol=4)) + 
        scale_fill_manual(values=stage_colors)
    
    p1_legend <- get_legend(p1)
    
    p1 <- p1 + theme(legend.position = "none")
    print("p1")
    
    ### make percentage of tissue
    temp <- as.data.frame(
        meta %>% 
            dplyr::select(group, Disease) %>% 
            dplyr::group_by(group, Disease) %>% 
            dplyr::add_tally() %>% 
            unique() %>% 
            dplyr::group_by(group) %>% 
            dplyr::mutate(freq = n / sum(n)) %>% 
            unique()
    )
    
    temp1 <- as.data.frame(table(meta$Disease))
    
    colnames(temp1) <- c("Disease", "n")
    temp1$group = "Total"
    temp1$freq = temp1$n / sum(as.numeric(temp1$n))
    
    temp1 <- temp1[, colnames(temp)]
    temp <- rbind(temp, temp1)
    
    temp$group <- factor(temp$group, levels = plot_order)
    
    ## combine colors of tissue and disease
    p2 <- ggplot(data = temp, aes(x=group, y=freq, fill=Disease)) + 
        geom_bar(stat = "identity", width = .75) + 
        coord_flip() + 
        theme_bw() +
        theme(
            legend.position = "right", 
            legend.title = element_blank(), 
            legend.justification = "center",
            axis.text = element_text(size = 15),
            axis.title = element_text(size = 15),
            legend.text = element_text(size = 15)
        ) +
        labs(y="Fraction of cells", x = "") +
        # scale_x_discrete(breaks=plot_order) + 
        guides(fill = guide_legend(ncol=1)) + 
        scale_fill_manual(values=colors)
    
    p2_legend <- get_legend(p2)
    
    p2 <- p2 + theme(legend.position = "none")
    print("p2")
    
    ### make percentage barplot of patients
    temp <- as.data.frame(
        meta %>% 
            dplyr::select(group, PatientID) %>% 
            dplyr::group_by(group, PatientID) %>% 
            dplyr::add_tally() %>% 
            unique() %>% 
            dplyr::group_by(group) %>% 
            dplyr::mutate(freq = n / sum(n)) %>% 
            unique()
    )
    
    temp1 <- as.data.frame(table(object@meta.data$PatientID))
    colnames(temp1) <- c("PatientID", "n")
    temp1$group = "Total"
    temp1$freq = temp1$n / sum(temp1$n)
    
    temp <- rbind(temp, temp1)
    temp$group <- factor(temp$group, levels = plot_order)
    
    p3 <- ggplot(data = temp, aes(x=group, y=freq, fill=PatientID)) + 
        geom_bar(stat = "identity", width = .75) + 
        coord_flip() + 
        theme_bw() +
        theme(
            legend.position = "right", 
            legend.title = element_blank(), 
            legend.justification = "center",
            legend.spacing.x = unit(0.1, 'cm'),
            axis.text = element_text(size = 15),
            axis.title = element_text(size = 15),
            legend.text = element_text(size = 10)
        ) +
        labs(y="Fraction of cells", x = "") +
        # scale_x_discrete(breaks=plot_order) + 
        guides(fill = guide_legend(ncol=ncol)) + 
        scale_fill_manual(values=colors)
    p3_legend <- get_legend(p3)
    
    p3 <- p3 + theme(legend.position = "none")
    print("p3")
    
    ### make bar plot of number of cells
    temp <- as.data.frame(
        meta %>% 
            dplyr::select(group) %>% 
            dplyr::group_by(group) %>% 
            dplyr::add_tally() %>% 
            unique()
    )
    
    temp1 <- data.frame(group="Total", n=nrow(meta))
    temp <- rbind(temp, temp1)
    temp$group <- factor(temp$group, levels = plot_order)
    p4 <- ggplot(temp, aes(x=group, y=n, fill=stata_pal("s1color")(1))) + 
        geom_bar(stat = "identity") +
        coord_flip() + 
        theme_bw() +
        theme(
            legend.position = "none",
            axis.text = element_text(size = 15),
            axis.title = element_text(size = 15),
            legend.text = element_text(size = 10)
        ) +
        labs(y="Number of cells", x = "") +
        # scale_x_discrete(breaks=plot_order) +
        scale_y_log10(
            breaks = trans_breaks("log10", function(x) 10^x),
            labels = trans_format("log10", math_format(10^.x))
        )
    print("p4")
    
    ### make boxplot of transcript number
    meta1 = meta
    meta1$group <- "Total"
    meta <- rbind(meta, meta1)
    meta$group <- factor(meta$group, levels = plot_order)

    p5 <- ggplot(meta, aes(x=group, group=group, y=nUMI)) + 
        geom_boxplot(fill = hc_pal()(1)) + 
        coord_flip() + 
        # scale_x_discrete(breaks=plot_order[1:(length(plot_order) - 1)]) +
        labs(y="Number of transcripts", x = "") + 
        scale_y_log10(
            breaks = trans_breaks("log10", function(x) 10^x),
            labels = trans_format("log10", math_format(10^.x))
        ) +
        theme_bw() + 
        theme(
            axis.text = element_text(size = 15),
            axis.title = element_text(size = 15),
            legend.text = element_text(size = 10)
        )
    print("p5")
    
    plots = list(p1_legend, p2_legend, p3_legend, ggplot() + theme_void(), ggplot() + theme_void() , p1, p2, p3, p4, p5)
    
    if (!with_disease) {
        plots = list(p1_legend, p3_legend, ggplot() + theme_void(), ggplot() + theme_void(), p1, p3, p4, p5)
    }
    
    p <- plot_grid(plotlist=plots, nrow = 2, rel_heights = rel_height)
    return(p)
}

```


```{R}
library(TCGAbiolinks)

lusc_cli <- readRDS("LungCancer10x/TCGA/LUSC_clinical.rds")
tcga <- read.csv(
    "LungCancer10x/TCGA/Lung.csv", 
    row.names = 1, 
    stringsAsFactors = F
)

tcga <- tcga[, str_detect(colnames(tcga), "LUSC")]

for (i in unique(temp$ident)) {
    genes = as.data.frame(temp)[temp$ident == i, "gene"]
    
    expr <- median(apply(tcga[intersect(genes, rownames(tcga)), ], 1, mean))

    samples <- colnames(tcga)[apply(tcga[intersect(genes, rownames(tcga)), ], 2, function(x) {
        mean(x) > expr
    })]
    
    samples <- sapply(samples, function(x) {
        temp = str_split(x, "\\.")[[1]]
        paste(temp[1], temp[2], temp[3], sep = "-")
    })
    
    lusc_cli$ident = ifelse(lusc_cli$submitter_id %in% samples, "Up", "Down")
    
    TCGAanalyze_survival(
        lusc_cli,
        "ident",
        main = paste("Cluster", i),
        height = 6, 
        width=8,
        risk.table = F,
        dpi = 600,
        conf.int = F,
        cumevents = F,
        cumcensor = F,
        ggtheme = theme_minimal() + theme(
            text = element_text(size = 15),
            axis.text = element_text(size = 15),
            axis.title = element_text(size = 18),
            legend.text = element_text(size = 15),
            legend.title = element_text(size = 18),
            title = element_text(size = 20)
        ),
        filename = paste("Basal_", i, ".pdf", sep = "")
    )
}



```


### Stage module heatmap

```{r}
make_module_heatmap <- function(obj, res, mark_genes) {
    
    meta = obj@meta.data[order(obj@meta.data$Stage, obj@meta.data$PatientID), ]
    
    meta$Disease <- as.character(meta$Disease)
    meta$Disease[meta$Disease == "LUAD"] <- "AD"
    meta$Disease[meta$Disease == "LUSC"] <- "SC"
    meta$Disease[meta$Disease == "LUAD_Normal"] <- "NL(AD)"
    meta$Disease[meta$Disease == "LUSC_Normal"] <- "NL(SC)"
    meta$Disease[meta$Disease == "Pleiomorphic Normal"] <- "NL(UPS)"
    meta$Disease[meta$Disease == "Pleiomorphic"] <- "UPS"
    meta$Disease[meta$Disease == "LULC"] <- "LC"
    meta$Disease[meta$Disease == "LULC_Normal"] <- "NL(LC)"
    set.seed(42)
    meta = meta[sample(1:nrow(meta)), ]
    meta <- meta[order(meta$Stage), c("Stage"), drop = F] # meta$Disease, , "Disease"
    
    
    meta$Cells = rownames(meta)
    num_cells = sort(table(meta$Stage), decreasing = F)
    num_cells = ifelse(num_cells[1] < 20, num_cells[2], num_cells[1])
    meta <- meta %>% group_by(Stage) %>% sample_n(num_cells, replace = T) %>% unique() %>% as.data.frame()
        
        rownames(meta) <- meta$Cells
    
    stage_colors <- c(
        "I"="#65A9A3", 
        "II"="#4A933E",
        "III"="#EC7A21",
        "IV"="#D73F47"
    )
    
    cluster_cols = c(
        as.character(wes_palette("Zissou1")),
        as.character(wes_palette("Royal2")),
        as.character(wes_palette("Royal1")),
        as.character(wes_palette("Darjeeling2")),
        as.character(wes_palette("Darjeeling1"))
    )
    
    patient_colors = cluster_cols[1:length(unique(obj@meta.data$PatientID))]
    names(patient_colors) <- unique(obj@meta.data$PatientID)

    # top annotation
    ta = HeatmapAnnotation(
        Stage = meta$Stage,
        # PatientID = meta$PatientID,
        col = list(
            Stage =  stage_colors
            # PatientID = patient_colors
        ),
        show_legend = T,
        show_annotation_name = T,
        annotation_legend_param = list(
            Stage=list(
                direction = "horizontal", 
                nrow = 1,
                labels_gp = gpar(fontsize = labels_size),
                title_gp = gpar(fontsize = title_size)
            )
            # PatientID=list(
            #     direction = "horizontal",
            #     nrow = 4,
            #     labels_gp = gpar(fontsize = labels_size),
            #     title_gp = gpar(fontsize = title_size)
            # )
        )
    )
    
    
    module_colors = wes_palette("GrandBudapest2", length(unique(res$Clt)), type = "continuous")
    names(module_colors) <- unique(res$Clt)
    
    # left annotation
    if(!is.null(mark_genes)) {
        la = rowAnnotation(
            foo = anno_mark(
                at = which(res$gene %in% mark_genes),
                labels = res$gene[res$gene %in% mark_genes],
                which = "row",
                side = "left"
            ),
            Module = res$Clt,
            col = list(
                Module = module_colors
            ),
            show_legend = F,
            show_annotation_name = F,
            annotation_name_rot = 0,
            annotation_legend_param=list(
                Module=list(
                    direction = "horizontal", 
                    nrow = 1,
                    labels_gp = gpar(fontsize = labels_size),
                    title_gp = gpar(fontsize = title_size)
                )
            )
        )
    } else {
        la = rowAnnotation(
            Module = res$Clt,
            col = list(
                Module = module_colors
            ),
            show_legend = T,
            show_annotation_name = T,
            annotation_name_rot = 0,
            annotation_legend_param=list(
                Module=list(
                    direction = "horizontal", 
                    nrow = 1,
                    labels_gp = gpar(fontsize = labels_size),
                    title_gp = gpar(fontsize = title_size)
                )
            )
        )
    }
    
    h <- Heatmap(
        name = "Expr",
        obj@scale.data[as.character(res$gene), as.character(rownames(meta))],
        cluster_rows = F,
        cluster_columns = F,
        show_column_names = F,
        show_row_names = F,
        col = colorRamp2(c(-2, 0, 2), c("purple", "black", "yellow")),
        top_annotation = ta,
        left_annotation = la,
        border = T,
        column_title_gp = gpar(fontsize=0),
        row_title_gp = gpar(fontsize = 0),
        column_split = meta$Stage,
        row_split = res$Clt,
        heatmap_legend_param = list(
            direction = "horizontal",
            labels_gp = gpar(fontsize = labels_size),
            title_gp = gpar(fontsize = title_size)
        ),
        use_raster = T
    )
    return(h)
}
```


### Perform Suerat pipeline 

```{r}
new_obj@meta.data$ident1 <- as.numeric(new_obj@meta.data$res.0.1) + 1

p <- DimPlot(
    new_obj, 
    reduction.use = "umap", 
    pt.size = 1, 
    do.label = T, 
    no.legend = T,
    do.return = T,
    label.size = 10,
    group.by = "ident1"
)

p <- p + 
    theme_bw() +
    theme(
        text = element_text(size = 10),
        legend.position = "none",
        plot.title = element_text(size = 30, face="bold", hjust = 0.5),
        axis.title = element_text(size = 25),
        axis.text = element_text(size = 20),
        aspect.ratio = 1
    ) +
    labs(title = "Basal (LUSC)")

p

ggsave(
    filename = paste0(root.dir, "Basal_umap_cluster.pdf"),
    plot = p,
    width = 6,
    height = 6
)
```

```{r}
p <- make_scatterpie_plot(
    new_obj, 
    color = "Stage", 
    colors = colors, 
    reduction.use = "umap",
    legend.position = c(0.8, 0.1)
) + theme(aspect.ratio = 1)


ggsave(
    filename = paste0(root.dir, "Basal_umap_stage.pdf"),
    plot = p,
    width = 6,
    height = 6
)


p <- make_scatterpie_plot(
    new_obj, 
    color = "PatientID", 
    colors = colors, 
    reduction.use = "umap",
    legend.position = c(0.8, 0.15),
    guide_ncol=2
) + theme(aspect.ratio = 1)

ggsave(
    filename = paste0(root.dir, "Basal_umap_patient.pdf"),
    plot = p,
    width = 6,
    height = 6
)
```

```{r}
pdf(paste0(root.dir, "Basal_combination_bar_plot.pdf"), width = 12, height = 6, onefile = F)
make_combined_barplot(new_obj, group.by = "ident", with_disease = F)
dev.off()
```


### Markers
```{r}
find_markers <- function(object, group.by = "Stage", n.cores = 4, min.pct = 0.1) {
    
    temp_group <- as.character(sort(unique(object@meta.data[, group.by])))
    
    library(doMC)
    registerDoMC(n.cores)
    res = foreach(i = temp_group, .combine = "rbind") %dopar% {

        groups = rownames(object@meta.data[object@meta.data[, group.by] == i, ])
        
        if(length(groups) >= 3) {
            temp <- FindMarkers(
                object = object, 
                ident.1 = groups,
                logfc.threshold = 0,
                min.pct = min.pct
            )
            
            temp$ident = i
            temp$gene = rownames(temp)
            return(temp)
        }
        return(NULL)
    }    
    return(res)
}

registerDoMC(5)
obj@meta.data$Clt = as.numeric(as.character(obj@meta.data$res.0.2)) + 1
markers <- find_markers(obj, group.by = "Clt", n.cores = 5)
write.csv(markers, paste0(root.dir, "/cluster_markers.csv"))
```


### cluter annotation

```{r}

do_tmod <- function(data, output_xlsx, msig, fig = NULL, force=FALSE) {
    if (file.exists(output_xlsx) && !force) {
        cerno = read.xlsx(output_xlsx, sheet = 1)
        utest = read.xlsx(output_xlsx, sheet = 2)
        ztest = read.xlsx(output_xlsx, sheet = 3)
    } else {
        sel <- msig$MODULES$Category == "C5"
        
        data = data[order(abs(data$avg_logFC), decreasing = T), ]
        cerno = NULL
        utest = NULL
        ztest = NULL
        
        if ("ident" %in% colnames(data)) {
            for(i in unique(data$ident)) {
                print(i)
                fig_data = list()
                
                temp_data = data[data$ident == i, ]
                temp = tmodCERNOtest(temp_data$gene, mset = msig[sel])
                fig_data$CERNO = temp
                
                if(nrow(temp) > 0) {
                    temp$ident = i
                    temp$Methods = "CERNO"
                    cerno = rbind(cerno, temp)
                }
                
                temp = tmodUtest(temp_data$gene, mset = msig[sel])
                fig_data$Utest = temp
                if(nrow(temp) > 0) {
                    temp$ident = i
                    temp$Methods = "Utest"
                    utest = rbind(utest, temp)
                }
                
                temp = tmodZtest(temp_data$gene, mset = msig[sel])
                fig_data$Ztest = temp
                if(nrow(temp) > 0) {
                    temp$ident = i
                    temp$Methods = "Ztest"
                    ztest = rbind(ztest, temp)
                }
            }
        } else {
            cerno = tmodCERNOtest(data$gene, mset = msig[sel])
            if (!is.null(cerno) && nrow(cerno) > 0) {
                cerno$Methods = "CERNO"
            }
            
            utest = tmodUtest(data$gene, mset = msig[sel])
            if (!is.null(utest) && nrow(utest) > 0) {
                utest$Methods = "Utest"
            }
            
            ztest = tmodZtest(data$gene, mset = msig[sel])
            if (!is.null(ztest) && nrow(ztest) > 0) {
                ztest$Methods = "Ztest"
            }
        }
        
        
        if(!is.null(cerno) || !is.null(utest) || !is.null(ztest)) {
            wb = createWorkbook()
            addWorksheet(wb, "CERNO")
            writeData(wb, 1, cerno)
            
            addWorksheet(wb, "Utest")
            writeData(wb, 2, utest)
            
            addWorksheet(wb, "Ztest")
            writeData(wb, 3, ztest)
            
            saveWorkbook(wb, output_xlsx, overwrite = T)
        }
    }
    
    if(
        "ident" %in% c(colnames(cerno), colnames(utest), colnames(ztest)) && 
        !is.null(fig) && 
        (!is.null(cerno) || !is.null(utest) || !is.null(ztest)) 
    ) {
        for(i in unique(c(cerno[, "ident"], utest[, "ident"], ztest[, "ident"]))) {
            p <- make_tmod_plots(
                cerno[cerno$ident == i, ], 
                utest[utest$ident == i, ], 
                ztest[ztest$ident == i, ], 
                topn = 12, 
                scales = "free"
            )
            
            ggsave(
                filename = paste(fig, paste0(i, "_tmod.pdf"), sep = "/"),
                plot = p,
                width = 20,
                height = 12,
                units = "in"
            )
        }
    }
    
    return(list(
        "CERNO"=cerno,
        "Utest"=utest,
        "Ztest"=ztest
    ))
}


msig <- tmodImportMSigDB(here("../../../project_scripts/12_final_plots/msigdb_v6.2.xml"))
```


```{r}
markers <- find_markers(new_obj, group.by = "ident", n.cores = 4)
write.xlsx(markers, paste0(root.dir, "Basal_cluster_markers.xlsx"), overwrite=T)
```



```{r fig.height=10, fig.width=8}
obj <- readRDS(full.path("11_CNV/each_cells/Basal/seurat_obj_LUSC.rds"))

markers <- read.csv(
    full.path("11_CNV/each_cells/Basal/cluster_markers_LUSC.csv"), 
    row.names = 1, stringsAsFactors = F
)


set.seed(42)
markers <- markers %>%
    filter(p_val_adj < 0.05 & avg_logFC > 0.5) %>%
    group_by(ident) %>%
    arrange(desc(avg_logFC)) %>%
    as.data.frame()
markers$x = 1:nrow(markers)
    

temp <- markers %>%
    filter(p_val_adj < 0.05 & avg_logFC > 0.25) %>%
    group_by(ident) %>%
    top_n(10, wt=avg_logFC) %>%
    as.data.frame()

obj@meta.data$ident = obj@meta.data$res.0.2

obj@meta.data$Cells = rownames(obj@meta.data)
cells_order <- obj@meta.data %>%
    group_by(ident) %>%
    arrange(ident) %>%
    sample_n(min(table(obj@meta.data$ident))) %>%
    as.data.frame()

# cells_order <- cells_order[sample(1:nrow(cells_order)), ]
# cells_order = cells_order[order(as.numeric(cells_order$ident)), ]

ccls <- as.character(wes_palette("GrandBudapest2", length(unique(cells_order$ident)), type = "continuous"))
names(ccls) <- unique(cells_order$ident)

ta = HeatmapAnnotation(
    Cluster = cells_order$ident,
    col = list(
        Cluster = ccls
    ),
    show_annotation_name = F,
    show_legend = F
)

ha = rowAnnotation(
    foo = anno_mark(
        at = temp$x, 
        labels = temp$gene,
        side = "left"
    )
)


# pdf(paste0(root.dir, "/cluster_markers.pdf"), width = 5, height = 6)
Heatmap(
    obj@scale.data[markers$gene, cells_order$Cells],
    name = "Expr",
    cluster_rows = F,
    cluster_columns = F,
    show_row_names = F,
    show_column_names = F,
    show_heatmap_legend = F,
    top_annotation = ta,
    left_annotation = ha,
    column_split = cells_order$ident,
    row_split = markers$ident,
    row_title = NULL,
    heatmap_legend_param = list(
        labels_gp = gpar(fontsize = 15),
        title_gp = gpar(fontsize = 15)
    ),
    col = colorRamp2(c(-2, 0, 2), c("purple", "black", "yellow"))
)
# dev.off()
```

```{r}
do_tmod(
    read.xlsx(paste0(root.dir, "Basal_cluster_markers.xlsx"), rowNames = T), 
    here("clusters/tmod.xlsx"), 
    msig=msig, 
    fig = NULL, 
    force = TRUE
)
```

```{R}
cerno <- read.xlsx(here("clusters/tmod.xlsx"), sheet = 1)

cerno <- cerno %>%
    filter(adj.P.Val < 0.05) %>%
    group_by(ident) %>%
    top_n(5, wt = AUC)

cerno <- reshape2::dcast(cerno, Title~ident, value.var = "AUC", fun.aggregate = mean, fill = 0)

rownames(cerno) <- cerno$Title
cerno <- cerno[, colnames(cerno) != "Title"]
```


```{R fig.height=12, fig.width=16}

col_fun <- colorRamp2(c(0.9, 1), c("white", "red"))

h <- Heatmap(
    as.matrix(cerno),
    name = "AUC",
    col = col_fun,
    column_names_rot = 0,
    column_names_centered = T,
    column_names_gp = gpar(fontsize = 20),
    row_names_gp = gpar(fontsize = 15),
    heatmap_legend_param = list(
        direction = "horizontal",
        labels_gp = gpar(fontsize = labels_size),
        title_gp = gpar(fontsize = title_size),
        title_position = "lefttop"
    ),
    border = T,
    width = unit(12, "cm"),
    rect_gp = gpar(col = "grey75", lwd = 1)
)

pdf(here("clusters/heatmap_cerno.pdf"), width = 16, height = 12)
draw(
    h, 
    heatmap_legend_side = "bottom"
)
dev.off()
```



```{r}
markers <- find_markers(new_obj, group.by = "Stage", n.cores = 4)
write.xlsx(markers, paste0(root.dir, "Basal_stage_markers.xlsx"), overwrite=T, rowNames = T)
```

```{R}
obj <- readRDS(paste0(root.dir, "Basal.rds"))
markers <- read.xlsx(paste0(root.dir, "Basal_stage_markers.xlsx"), rowNames = T)
nms = unique(markers$gene)
ig_genes = c(grep("^IGJ", nms, v=T), 
             grep("^IGH",nms,v=T),
             grep("^IGK", nms, v=T), 
             grep("^IGL", nms, v=T))

bad_genes = unique(c(grep("^MT-", nms, v=T), grep("^MTMR", nms, v=T), grep("^MTND", nms, v=T),"NEAT1","TMSB4X", "TMSB10", ig_genes))



markers <- markers %>% 
    filter(avg_logFC > 0.5 & p_val_adj < 0.05 & !gene %in% bad_genes)


expr = ExpressionSet(
    as.matrix(obj@scale.data[unique(markers$gene),])
)
    
m = mestimate(expr)

temp = Dmin(expr, m=m, crange=seq(2,10,1), repeats=3, visu=FALSE)
print(temp)
if (sum(!is.na(temp)) == 0) {
    c = length(unique(obj@meta.data$Stage))
} else {
    new_temp = temp[!is.na(temp)]
    
    # using kneed to selected best cluster
    kee = import("kneed")

    c = kee$KneeLocator(
        1:length(new_temp),
        new_temp, 
        curve='convex', 
        direction='decreasing'
    )
    c = c$knee
    
    c = which(temp == new_temp[c])
}

cl <- mfuzz(expr, c=c, m=m)


# extract results
res = as.data.frame(cl$cluster)
colnames(res) <- "Clt"
res$gene = rownames(res)

res = res[order(res$Clt), ]

res$Clt <- sapply(res$Clt, function(x) {
    temp = c(
        "1"="M.I",
        "2"="M.III",
        "3"="M.II"
    )
    
    return(temp[as.character(x)])
})

write.csv(res, paste0(root.dir, "Basal_mfuzz.csv"))
```

```{r height = 6, width = 6}
res = read.csv(paste0(root.dir, "11_CNV/each_cells/Basal/LUSC/mfuzz.csv"), row.names = 1, stringsAsFactors = F)

set.seed(42)
temp = res$gene[sample(1:nrow(res))]

labels_size = 12
title_size = 15

pdf(paste0(root.dir, "11_CNV/each_cells/Basal/LUSC/Basal_mfuzz.pdf"), height = 6, width = 8)
draw(
    make_module_heatmap(obj, res, mark_genes = c(head(temp, n = 10), "PPT1", "AQP3", "SPHK1", "PPT1", "KPNA2", "KLF6")), 
    column_title = "",
    column_title_gp = gpar(fontsize = 20),
    merge_legend = TRUE, 
    heatmap_legend_side = "bottom", 
    annotation_legend_side = "bottom"
)
dev.off()
```


### Make a cluster plot of all LUSC Basal cells
```{r}
meta <- read.csv(paste0(root.dir, "../../02_rds/meta_after_singleR.csv"), row.names = 1, stringsAsFactors = F)

nm_meta = read.xlsx(paste0(root.dir, "../../02_rds/NM_meta.xlsx"))
rownames(nm_meta) <- nm_meta$SampleID

meta$Disease[meta$SampleID %in% nm_meta$SampleID] = nm_meta[meta$SampleID[meta$SampleID %in% nm_meta$SampleID], "Disease"]
meta$Stage[meta$SampleID %in% nm_meta$SampleID] = nm_meta[meta$SampleID[meta$SampleID %in% nm_meta$SampleID], "Stage"]

meta <- meta[as.character(meta$cell_short) == "Basal", ]
meta <- meta[as.character(meta$Disease) == "LUSC", ]

expr <- readRDS(paste0(root.dir, "../../02_rds/all_cell_expr.rds"))
```

```{r}
all_obj <- CreateSeuratObject(
    expr[, rownames(meta)],
    meta.data = meta
)


mito.genes <- grep(pattern = "^MT-", x = rownames(x = all_obj@raw.data), value = TRUE)
percent.mito <- Matrix::colSums(all_obj@raw.data[mito.genes, ])/Matrix::colSums(all_obj@raw.data)

# AddMetaData adds columns to object@meta.data, and is a great place to
# stash QC stats
all_obj <- AddMetaData(object =all_obj, metadata = percent.mito, col.name = "percent.mito")


### 3. do PCA and harmony

pcs = min(100, min(nrow(all_obj@raw.data), ncol(all_obj@raw.data)) - 1)

all_obj <- NormalizeData(object = all_obj, normalization.method = "LogNormalize", scale.factor = 10000)
all_obj <- FindVariableGenes(object = all_obj, mean.function = ExpMean, dispersion.function = LogVMR, x.low.cutoff = 0.0125, x.high.cutoff = 3, y.cutoff = 0.5)
all_obj <- ScaleData(object = all_obj, vars.to.regress = c("nUMI", "percent.mito"))
all_obj <- RunPCA(object = all_obj, pc.genes = all_obj@var.genes, do.print = FALSE, pcs.compute = pcs)

all_obj <- RunHarmony(all_obj, c("batch", "PatientID"))


#### 4. make dotplot of SD of PCAs and Harmonys

temp <- as.data.frame(apply(all_obj@dr$harmony@cell.embeddings, 2, sd))
colnames(temp) <- c("sd")
temp$Harmony = 1:nrow(temp)

kee = import("kneed")
n.pcs = kee$KneeLocator(
    1:ncol(all_obj@dr$pca@cell.embeddings),
    apply(all_obj@dr$pca@cell.embeddings, 2, sd), 
    curve='convex', 
    direction='decreasing'
)
n.pcs = n.pcs$knee


#### 5. tSNE and UMAP
all_obj <- RunTSNE(
    object = all_obj, 
    dims.use = 1:n.pcs, 
    do.fast = TRUE, 
    reduction.use = "harmony", 
    reduction.name = "tsne",
    perplexity=min(30, floor(pcs / 3) + 1)
)
all_obj <- RunUMAP(
    object = all_obj, 
    dims.use = 1:n.pcs, 
    do.fast = TRUE, 
    reduction.use = "harmony",
    n_neighbors = min(30, pcs)
)
all_obj <- FindClusters(object = all_obj, reduction.type = "harmony", dims.use = 1:10, resolution = 0.02, print.output = 0, save.SNN = FALSE, force.recalc = TRUE)

saveRDS(all_obj, paste0(root.dir, "total_basal.rds"))
```


```{r}
all_obj <- readRDS(paste0(root.dir, "total_basal.rds"))
```


```{r}
all_obj@meta.data$ident1 <- as.numeric(all_obj@meta.data$res.0.03) + 1
all_obj@meta.data$ident <- all_obj@meta.data$ident1

p <- DimPlot(
    all_obj, 
    reduction.use = "umap", 
    pt.size = 0.01, 
    do.label = T, 
    no.legend = T,
    do.return = T,
    label.size = 10,
    group.by = "ident1"
)

p <- p + 
    theme_bw() +
    theme(
        text = element_text(size = 10),
        legend.position = "none",
        plot.title = element_text(size = 30, face="bold", hjust = 0.5),
        axis.title = element_text(size = 25),
        axis.text = element_text(size = 20),
        aspect.ratio = 1
    ) +
    labs(title = "Basal (LUSC)")

p

ggsave(
    filename = paste0(root.dir, "Basal_all_umap_cluster.pdf"),
    plot = p,
    width = 6,
    height = 6
)
```


```{r fig.height=6, fig.width=6}
p <- make_scatterpie_plot(
    all_obj, 
    color = "Stage", 
    colors = colors, 
    reduction.use = "umap",
    legend.position = c(0.2, 0.08),
    guide_ncol = 3
) + theme(aspect.ratio = 1)

p

ggsave(
    filename = paste0(root.dir, "Basal_all_umap_stage.pdf"),
    plot = p,
    width = 6,
    height = 6
)


cluster_cols = c(
    as.character(wes_palette("Zissou1")),
    as.character(wes_palette("Royal2")),
    as.character(wes_palette("Royal1")),
    as.character(wes_palette("Darjeeling2")),
    as.character(wes_palette("Darjeeling1"))
)
patient_colors = cluster_cols[1:length(unique(as.character(all_obj@meta.data$PatientID)))]
names(patient_colors) <- as.character(unique(all_obj@meta.data$PatientID))
colors = c(stage_colors, disease_colors, tissue_colors, patient_colors)


p <- make_scatterpie_plot(
    all_obj, 
    color = "PatientID", 
    colors = colors, 
    reduction.use = "umap",
    legend.position = c(0.1, 0.85),
    guide_ncol=3,
    legend.text = 12
) + theme(aspect.ratio = 1)

p

ggsave(
    filename = paste0(root.dir, "Basal_all_umap_patient.pdf"),
    plot = p,
    width = 6,
    height = 6
)
```


```{r fig.height=6, fig.width=12}
pdf(paste0(root.dir, "Basal_all_combination_bar_plot.pdf"), width = 12, height = 6, onefile = F)
make_combined_barplot(all_obj, group.by = "ident", with_disease = F)
dev.off()
```


```{r fig.height=3, fig.width=4}
calculate_cluster_perc <- function(obj, seed.range = 1:50) {
    meta = obj@meta.data

    num = table(meta$Stage)

    sf = num / min(num)
    
    # sf = table(meta$ident)
    # sf = sf / min(sf)
    
    temp <- meta %>%
        group_by(ident, Stage) %>%
        add_tally() %>%
        dplyr::select(ident, Stage, n) %>%
        unique() %>%
        as.data.frame()
    
    temp$freq <- apply(temp, 1, function(row) {
        as.numeric(row[3]) / sf[[row[2]]] # / num[[row[2]]]
    })
    
    temp <- temp %>% group_by(ident) %>%
        mutate(freq = freq / sum(freq)) %>%
        as.data.frame()
    
    temp
}

atii_perc <- calculate_cluster_perc(all_obj)

p <- ggplot(atii_perc, aes(x=factor(ident), y = freq, color = Stage)) + 
    geom_point() +
    geom_line(aes(x=ident, y = freq, color = Stage)) +
    scale_color_manual(values = c(
        "I"="#65A9A3", 
        "II"="#4A933E",
        "III"="#EC7A21",
        "IV"="#D73F47"
    )) +
    labs(x = "", y = "", color = "") +
    theme_bw() +
    theme(
        axis.text = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.position = c(0.1, 0.25),
        legend.background = element_blank()
    )
p


ggsave(
    filename = paste0(root.dir, "Basal_all_stage_perc.pdf"),
    plot = p,
    width = 4,
    height = 3
)
```


```{r}
markers <- find_markers(all_obj, group.by = "ident", n.cores = 5, min.pct = 0.1)

write.csv(markers, paste0(root.dir, "Basal_all_cells_cluster_markers.csv"))
```

```{R fig.height=10, fig.width=8}
markers <- read.csv()
markers <- markers[order(markers$avg_logFC, decreasing = T), ]

temp <- markers %>%
    filter(p_val_adj < 0.05) %>%
    group_by(ident) %>%
    top_n(10, wt=avg_logFC)


cells_order = rownames(all_obj@meta.data)[order(as.character(all_obj@meta.data$ident))]

ccls <- cluster_cols[1:length(unique(all_obj@meta.data$ident))]
names(ccls) <- unique(all_obj@meta.data$ident)

ta = HeatmapAnnotation(
    Cluster = as.character(all_obj@meta.data$ident)[order(as.character(all_obj@meta.data$ident))],
    col = list(
        Cluster = ccls
    ),
    show_annotation_name = F,
    show_legend = F
)


# pdf(paste0(root.dir, "ATII_cluster_markers_heatmap.pdf"), width = 6, height = 8)
Heatmap(
    name = "Expr",
    all_obj@scale.data[temp$gene, cells_order],
    cluster_rows = F,
    cluster_columns = F,
    show_row_names = T,
    show_column_names = F,
    top_annotation = ta,
    column_split = as.character(all_obj@meta.data$ident[order(as.character(all_obj@meta.data$ident))]),
    row_split = as.character(temp$ident),
    heatmap_legend_param = list(
        labels_gp = gpar(fontsize = 15),
        title_gp = gpar(fontsize = 15)
    ),
    col = colorRamp2(c(-2, 0, 2), c("purple", "black", "yellow"))
)
# dev.off()
```

## make violin plot
```{r fig.height=30, fig.width=16}
VlnPlot(all_obj, features.plot = sort(temp$gene), group.by = "ident1", point.size.use = 0)
```


```{r}
mfuzz <- read.csv(paste0(root.dir, "Basal_mfuzz.csv"), row.names = 1, stringsAsFactors = F)

sort(temp$gene[temp$gene %in% mfuzz$gene])
```

```{r}
sort(intersect(as.character(temp$gene), as.character(scores$gene_name[scores$V1 > 0.5])))
```


### make final cluste vln plot
```{r fig.height=10, fig.width=5}
target_genes = c(
    "AKR1B10", "AKR1C3", "KRT15",
    "PCNA",
    "LCN2", "SLPI",
    "IGLC2", "CD74",
    "CALD1", "IGFBP7", "LGALS1"
    
)

temp_expr <- melt(as.matrix(all_obj@scale.data[target_genes, ]))
temp_expr$ident <- all_obj@meta.data[temp_expr$Var2, "ident"]

library(ggpubr)
p <- ggviolin(
    data = temp_expr,
    x = "ident",
    y = "value",
    fill = "ident",
    add = "boxplot", 
    add.params = list(fill = "white")
) +
    facet_grid(Var1~., scales = "free") +
    labs(x = "", y = "") +
    theme_bw() +
    theme(
        legend.position = "none",
        axis.text.y = element_text(size = 12),
        axis.text.x = element_text(size = 15),
        strip.text = element_text(size = 12)
    )

p

ggsave(
    filename = paste0(root.dir, "Basal_cluster_markers_vln.pdf"),
    plot = p,
    width = 5,
    height = 10
)
```

### Stage module radar plot
```{R}
format_radar_expr_mat <- function(
  obj, 
  module,
  legend.position="bottom",
  plot.title="",
  legend.title = "",
  axis.label.size = 8,
  batch=F,
  Stage = NULL
) {
    meta = as.data.frame(obj@meta.data)
    
    if (!is.null(Stage)) {
        meta = meta[meta$Stage == Stage, ]
    }
    
    if (batch) {
        meta = meta[meta$Batch == 3, ]
        meta$Group = meta$PatientID
    } else {
        meta = meta[meta$Batch != 3, ]
        meta$Group = meta$PatientID # paste0(meta$PatientID, "(", meta$Stage, ")")
    }
    

    cells = meta[, "Group", drop = F]
    cells$Cells = rownames(cells)
    
    mat = obj@scale.data[as.character(module$gene), ]
    
    mat = melt(as.matrix(mat))
    mat = merge(cells, mat, by.x = "Cells", by.y = "Var2")
    
    module$Clt = as.character(module$Clt)
    # module$Clt = sapply(module$Clt, function(x) { paste0("M.", x) })
    
    module = merge(module, mat, by.x = "gene", by.y = "Var1", all.x = T)
    
    
    module = module %>% group_by(Clt, Group, gene) %>%
        mutate(m = mean(value)) %>%
        dplyr::select(Clt, Group, m) %>%
        unique() %>%
        group_by(Clt, Group) %>%
        mutate(value = mean(m)) %>%
        dplyr::select(Clt, Group, value) %>%
        unique()
    
    # return(module)
    module = as.data.frame(module)
    
    # return(module)
    #print(module)
    # 
    # module
    p <- ggradar(
        dcast(module, Group~Clt),
        centre.y = floor(min(module$value)),
        values.radar = round(seq(from=floor(min(module$value)), to=max(module$value), length.out = 3), 2),
        grid.max = ceiling(max(module$value)),
        grid.label.size = 8,
        legend.position = legend.position,
        legend.title = legend.title,
        plot.title = plot.title,
        axis.label.size = axis.label.size,
        group.line.width = 0.5,
        group.point.size = 4,
        axis.label.offset = 1.15,
        base.size = 10
    )
    p = p + theme(
        legend.direction="horizontal",
        legend.spacing.x = unit(0.2, 'cm'),
        legend.position = c(0.5, 0.01),
        legend.background = element_blank(),
        legend.text = element_text(size = 20),
        plot.title = element_text(size = 25, hjust = 0.5)
    ) +
        guides(
            color = guide_legend(override.aes = list(ncol = 3))
        )
    return(p)
}

```

```{R fig.height=6, fig.width=6}

for (i in unique(obj@meta.data$Stage)) {
    p = format_radar_expr_mat(
        obj, 
        res,
        legend.position="bottom",
        # plot.title=basename(dirname(dirname(i))),
        plot.title = paste0("Basal ", "(", i, ")"),
        axis.label.size = 10,
        Stage=i
    )
    
    ggsave(
        filename = here(paste0("mfuzz/mfuzz_radar_", i, ".pdf")),
        plot = p,
        width = 6,
        height = 6
    )
}


```

### Corrlation
```{R fig.height=8, fig.width=8}
stage_cor <- cor(t(obj@scale.data[res$gene, ]))

module_colors = wes_palette("GrandBudapest2", length(unique(res$Clt)), type = "continuous")
names(module_colors) <- unique(res$Clt)


ba <- HeatmapAnnotation(
    Module=res$Clt,
    col = list(
        Module=module_colors
    )
)


ra <- rowAnnotation(
    Module=res$Clt,
    col = list(
        Module=module_colors
    ),
    show_legend = F,
    show_annotation_name = F
)


pdf("mfuzz/module_gene_cor.pdf", width = 8, height = 6)
p <- Heatmap(
    stage_cor[res$gene, res$gene],
    name = "Corr",
    show_row_names = F,
    show_column_names = F,
    right_annotation = ra,
    bottom_annotation = ba
)
draw(p, merge_legend=T)
dev.off()
# ggcorrplot::ggcorrplot(stage_cor)
```


```{r}

stage_cor <- cor(obj@scale.data[res$gene, ])

module_colors = wes_palette("GrandBudapest2", length(unique(res$Clt)), type = "continuous")
names(module_colors) <- unique(res$Clt)

meta <- obj@meta.data
meta <- meta[order(meta$Stage, meta$PatientID), ]

ba <- HeatmapAnnotation(
    Stage=meta$Stage,
    Patient=meta$PatientID,
    col = list(
        Stage=stage_colors,
        Patient=patient_colors
    )
)


ra <- rowAnnotation(
    Stage=meta$Stage,
    Patient=meta$PatientID,
    col = list(
        Stage=stage_colors,
        Patient=patient_colors
    ),
    show_legend = F,
    show_annotation_name = F
)


h <- Heatmap(
    stage_cor[rownames(meta), rownames(meta)],
    name = "Corr",
    show_row_names = F,
    show_column_names = F,
    right_annotation = ra,
    bottom_annotation = ba,
    cluster_rows = F,
    cluster_columns = F
)

pdf("mfuzz/module_cell_cor.pdf", width = 8, height = 6)
draw(h, merge_legend=T)
dev.off()
```

## corr on pseudobulk

```{r}
temp <- melt(as.matrix(obj@scale.data))

temp$PatientID <- obj@meta.data[temp$Var2, "PatientID"]

temp <- temp %>%
    group_by(Var1, PatientID) %>%
    mutate(value = mean(value)) %>%
    dplyr::select(Var1, value, PatientID) %>%
    unique()

temp <- dcast(temp, Var1 ~ PatientID, fun.aggregate = mean, value.var = "value")

rownames(temp) <- temp$Var1
temp <- temp[, colnames(temp) != "Var1"]


temp_cor <- cor(temp)

temp_meta <- unique(obj@meta.data[, c("PatientID", "Stage")])
rownames(temp_meta) <- temp_meta$PatientID

ba <- HeatmapAnnotation(
    Stage=temp_meta[colnames(temp_cor), "Stage"],
    col = list(
        Stage=stage_colors
    )
)

ra <- rowAnnotation(
    Stage=temp_meta[rownames(temp_cor), "Stage"],
    col = list(
        Stage=stage_colors
    )
)

Heatmap(
    temp_cor,
    right_annotation = ra,
    bottom_annotation = ba,
    clustering_method_rows = "average",
    clustering_method_columns = "average"
)
```


```{r}
temp <- melt(as.matrix(obj@scale.data[res$gene, ]))

temp$PatientID <- obj@meta.data[temp$Var2, "PatientID"]

temp <- temp %>%
    group_by(Var1, PatientID) %>%
    mutate(value = mean(value)) %>%
    dplyr::select(Var1, value, PatientID) %>%
    unique()

temp <- dcast(temp, Var1 ~ PatientID, fun.aggregate = mean, value.var = "value")

rownames(temp) <- temp$Var1
temp <- temp[, colnames(temp) != "Var1"]


temp_cor <- cor(temp)

temp_meta <- unique(obj@meta.data[, c("PatientID", "Stage")])
rownames(temp_meta) <- temp_meta$PatientID

ba <- HeatmapAnnotation(
    Stage=temp_meta[colnames(temp_cor), "Stage"],
    col = list(
        Stage=stage_colors
    )
)

ra <- rowAnnotation(
    Stage=temp_meta[rownames(temp_cor), "Stage"],
    col = list(
        Stage=stage_colors
    ),
    show_legend = F,
    show_annotation_name = F
)

h <- Heatmap(
    temp_cor,
    right_annotation = ra,
    bottom_annotation = ba
)

draw(h, merge_legend = T)
```

## format meta info

```{R}
expr <- readRDS(here("../../../02_rds/seurat_obj_nm.rds"))
```


```{r}
meta <- read.xlsx(here("../../../02_rds/NM_meta.xlsx"))

rownames(meta) <- meta$SampleID
meta$PatientID <- paste0("DL", meta$PatientID)

expr@meta.data$PatientID <- as.character(expr@meta.data$PatientID)
expr@meta.data$Disease <- as.character(expr@meta.data$Disease)
expr@meta.data$PatientID[expr@meta.data$Batch == 3] <- meta[as.character(expr@meta.data$SampleID[expr@meta.data$Batch == 3]), "PatientID"]
expr@meta.data$Disease[expr@meta.data$Batch == 3] <- meta[as.character(expr@meta.data$SampleID[expr@meta.data$Batch == 3]), "Disease"]
expr@meta.data$Stage[expr@meta.data$Batch == 3] <- meta[as.character(expr@meta.data$SampleID[expr@meta.data$Batch == 3]), "Stage"]
```


```{r fig.height=6, fig.width=6}
temp_meta <- expr@meta.data[expr@meta.data$cell_name == "Basal" & expr@meta.data$Disease %in% c("LUSC"), ]

temp_obj <- CreateSeuratObject(
    expr@raw.data[, rownames(temp_meta)],
    meta.data = temp_meta
)

temp_obj@scale.data = expr@scale.data[, rownames(temp_meta)]

temp_obj@meta.data$PatientID <- paste(temp_obj@meta.data$PatientID, temp_obj@meta.data$Stage, sep = "-")

p <- format_radar_expr_mat(
    temp_obj, 
    res,
    legend.position="bottom",
    # plot.title=basename(dirname(dirname(i))),
    plot.title = "Basal (DL2018)",
    axis.label.size = 10,
    batch = T
)

p

ggsave(
    filename = here(paste0("mfuzz/mfuzz_radar_DL.pdf")),
    plot = p,
    width = 6,
    height = 6
)
```

### Radar plots on bulk data
```{r}
bulk_meta <- read.xlsx(here("../../../09_bulk/RNATAC50commonSample-1119.xlsx"))
bulk_meta <- bulk_meta[!is.na(bulk_meta$SampleID), ]
rownames(bulk_meta) <- bulk_meta$SampleID
```

### MuSic
```{r}
bulk <- read.xlsx(here("../../../09_bulk/RNA_seq/RSEM.xlsx"), rowNames = T)
bulk <- bulk[, str_detect(colnames(bulk), "LUSC")]

estimate <- readRDS(here("../../../09_bulk/RNA_seq/MuSic/bulk.rds"))
estimate <- estimate$Est.prop.weighted
estimate <- as.matrix(t(estimate))

for (i in colnames(bulk)) {
    if (estimate["Basal", i] > 0) {
        bulk[, i] <- bulk[, i] / estimate["Basal", i]
    } else {
        bulk[, i] <- 0
    }
    
}

bulk <- scale(bulk)
bulk <- melt(as.matrix(bulk))
bulk <- na.omit(bulk)
```

```{r}
gene_pos <- read.table(here("../../../09_bulk/RNA_seq/gene_pos.txt"))
rownames(gene_pos) <- gene_pos$V5

bulk$Var1 <- gene_pos[bulk$Var1, "V6"]

bulk <- merge(bulk, res, by.x = "Var1", by.y = "gene")
bulk$Stage <- bulk_meta[bulk$Var2, "Stage"]
bulk$Stage <- str_replace_all(bulk$Stage, "[^IV]", "")
bulk <- bulk[bulk$Stage != "", ]
```

```{r}
temp <- dcast(bulk, Var1~Var2, fun.aggregate = mean, value.var = "value")
rownames(temp) <- temp$Var1
temp <- temp[, colnames(temp) != "Var1"]

rownames(res) <- res$gene

temp1 = as.matrix(temp[rowSums(temp) > 0, ])
col_fun <- colorRamp2(c(-1, 0, 1), c("blue", "white", "red"))

# pdf(here("mfuzz/mfuzz_bulk_MuSic.pdf"), width = 7, height = 15) 
Heatmap(
    temp1,
    name = "Expr",
    cluster_columns = T,
    cluster_rows = T,
    show_row_names = F,
    show_column_names = F,
    row_split = res[rownames(temp1), "Clt"],
    column_split = str_replace_all(bulk_meta[colnames(temp1), "Stage"], "[^IV]", ""),
    col = col_fun,
    border = T
)
# dev.off()
```

### Radar plots of MuSiC results
```{r}
temp <- bulk %>%
    group_by(Clt, Stage, Var2, Var1) %>%
    mutate(value = mean(value)) %>%
    unique() %>%
    group_by(Clt, Stage, Var2) %>%
    mutate(value = mean(value)) %>%
    dplyr::select(Var2, value, Clt, Stage) %>%
    unique()
```


```{r}
for(i in unique(temp$Stage)) {
    temp1 <- dcast(temp[temp$Stage == i, ], Var2~Clt, fun.aggregate = mean, value.var = "value")

    p <- ggradar(
        temp1,
        centre.y = -0.1,
        values.radar = round(seq(
            from=-0.1,
            to=0.1,
            length.out = 3
        ), 2),
        grid.max = 0.1,
        grid.label.size = 8,
        legend.position = "bottom",
        legend.title = "",
        plot.title = i,
        axis.label.size = 5,
        group.line.width = 0.5,
        group.point.size = 4,
        axis.label.offset = 1.15,
        base.size = 10
    )
    p = p + theme(
        legend.direction="horizontal",
        legend.spacing.x = unit(0.2, 'cm'),
        legend.position = c(0.5, 0.01),
        legend.background = element_blank(),
        legend.text = element_text(size = 20),
        plot.title = element_text(size = 25, hjust = 0.5)
    ) +
    guides(
        color = guide_legend(override.aes = list(ncol = 3))
    )

    ggsave(
        filename = here(paste0("mfuzz/mfuzz_radar_MuSiC_", i, ".pdf")),
        plot = p,
        width = 6,
        height = 6
    )
}
```



```{r}
temp <- bulk %>%
    group_by(Var2, Clt, Stage, Var1) %>%
    mutate(value = mean(value)) %>%
    group_by(Var2, Clt, Stage) %>%
    mutate(value = mean(value)) %>%
    dplyr::select(Var2, value, Clt, Stage) %>%
    unique()

temp <- reshape2::dcast(temp, Var2~Stage + Clt, fun.aggregate = mean, value.var = "value" )
```


### make violin plot
```{r}
target_genes <- c(
    "PDCD1", "CD28", "CTLA4",
    "ICOS", "BTLA", "LAG3",
    "TNFRSF9", "TNFRSF4",
    "CD27", "CD40LG", "HAVCR2",
    "GZMA", "GZMB", "GZMM", 
    "GZMH", "GZMK", "SIGLEC15"
)

```


```{r}
format_expr_by_mean_for_heatmap <- function(object, group.by, genes.use=NULL) {
    if (is.null(genes.use)) {
        genes.use = rownames(object@scale.data)
    }
    
    expr <- reshape2::melt(as.matrix(object@scale.data[genes.use, ]))
    expr$group = object@meta.data[expr$Var2, group.by]
    
    expr <- expr %>% 
        group_by(Var1, group) %>%
        mutate(value = mean(value)) %>%
        dplyr::select(Var1, group, value) %>%
        unique()
    
    expr <- reshape2::dcast(expr, Var1~group, fun.aggregate = mean, value.var = "value")
    rownames(expr) <- expr$Var1

    return(expr[, colnames(expr) != "Var1"])
}



# gene_drug_lusc <- gene_drug[!gene_drug %in% c("KDR", "EFCAB13")]

expr <- format_expr_by_mean_for_heatmap(all_obj, group.by = "ident", genes.use = target_genes)
# colnames(expr) <- sapply(colnames(expr), function(x) {
#     renames <- c(
#         "LUAD"="LUAD (n=21)",
#         "LUSC"="LUSC (n=2428)",
#         "LUAD_Normal"="Norm. LUAD (n=5)",
#         "LUSC_Normal"="Norm. LUSC (n=1) "
#     )
#     
#     renames[[x]]
# })
```


```{r fig.height=12, fig.width=15}
pdf(here("gene_drug_vln/heatmap_disease_normal.pdf"), width = 15, height = 15)
Heatmap(
    as.matrix(expr),
    col = colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
    heatmap_legend_param = list(
        title = "Mean expr.",
        labels_gp = gpar(fontsize = labels_size),
        title_gp = gpar(fontsize = title_size)
    ),
    column_names_rot = 0,
    column_names_centered = T,
    column_names_gp = gpar(fontsize = 20),
    row_names_gp = gpar(fontsize = 15)
)
dev.off()
```

```{r fig.height=12, fig.width=8}

expr_clt <- format_expr_by_mean_for_heatmap(all_obj, group.by = "ident", genes.use = target_genes)

pdf(paste0(root.dir, "Basal_gene_drug_heatmap_clusters.pdf"), width = 8, height = 6)
Heatmap(
    as.matrix(expr_clt),
    col = colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
    heatmap_legend_param = list(
        title = "Mean expr.",
        # legend_height = unit(3, "cm"),
        # legend_width = unit(10, "cm"),
        labels_gp = gpar(fontsize = 20),
        title_gp = gpar(fontsize = 25),
        legend_height = unit(10, "cm"),
        grid_width = unit(1, "cm")
    ),
    column_names_rot = 0,
    column_names_centered = T,
    column_names_gp = gpar(fontsize = 20),
    row_names_gp = gpar(fontsize = 15)
)
dev.off()
```

```{r fig.height=20, fig.width=16}
library(ggpubr)

expr <- melt(as.matrix(obj@raw.data[target_genes, ]))
expr$ident <- obj@meta.data[expr$Var2, "res.0.8"]

ggviolin(
    data = expr,
    x = "ident",
    y = "value"
) + facet_wrap(.~Var1, scales = "free_y") +
    scale_y_log10()
```





##### First Round
```{r}
read_protein <- function(path) {
    
    wb <- loadWorkbook(path)
    sheets <- wb$sheet_names
    
    res = NULL
    data <- read.xlsx(wb, sheet = which(sheets == "SNR532"))
    data <- data[, !colnames(data) %in% c("Block", "Column", "Row")]
    # data[, 3:ncol(data)] <- scale(data[, 3:ncol(data)])
    data <- reshape2::melt(data, id.vars = c("Name", "ID"))
    data$source = "SNR532"
    res = rbind(data, res)
    
    data <- read.xlsx(wb, sheet = which(sheets == "SNR635"))
    data <- data[, !colnames(data) %in% c("Block", "Column", "Row")]
    # data[, 3:ncol(data)] <- scale(data[, 3:ncol(data)])
    data <- reshape2::melt(data, id.vars = c("Name", "ID"))
    data$source = "SNR635"
    
    res = rbind(res, data)
    
    res$variable = as.character(res$variable)
    res$variable[str_detect(res$variable, "^LA")] <- "LUAD"
    res$variable[str_detect(res$variable, "^LS")] <- "LUSC"
    res$variable[str_detect(res$variable, "^NC")] <- "Normal"
    res <- res[res$variable %in% c("LUAD", "LUSC", "Normal"), ]
    
    return(res)
}


data1 <- read_protein(here("../../../10_protein/1_round.xlsx"))
data2 <- read_protein(here("../../../10_protein/2_round.xlsx"))
```



### protein heatmap
```{r fig.height=12, fig.width=8}
temp = data1[data1$Name %in% target_genes & data1$source == "SNR532", ]

temp = reshape2::dcast(temp, Name~variable, value.var = "value", fun.aggregate = mean)
rownames(temp) <- temp$Name
temp <- temp[, colnames(temp) != "Name"]

# temp <- temp[!rownames(temp) %in% c("CD40LG", "NCAM1", "KDR", "FLT3", "EFCAB13"), ]

col_fun <- colorRamp2(c(-2, 0, 2), c("blue", "white", "red"))

pdf(here("gene_drug_vln/heatmap_SNR532_IgG.pdf"), width = 7, height = 6) 
Heatmap(
    t(scale(t(temp))),
    # col = col_fun,
    column_names_rot = 0,
    column_names_gp = gpar(fontsize = 20),
    column_names_centered = T,
    row_names_gp = gpar(fontsize = 15),
    heatmap_legend_param = list(
        title = "Expr",
        labels_gp = gpar(fontsize = 20),
        title_gp = gpar(fontsize = 25),
        legend_height = unit(10, "cm"),
        grid_width = unit(1, "cm")
    ),
    column_title = "IgG",
    column_title_gp = gpar(fontsize = 30)
)
dev.off()
```


```{r fig.height=12, fig.width=8}
temp = data1[data1$Name %in% target_genes & data1$source == "SNR635", ]

temp = reshape2::dcast(temp, Name~variable, value.var = "value", fun.aggregate = mean)
rownames(temp) <- temp$Name
temp <- temp[, colnames(temp) != "Name"]

# temp <- temp[!rownames(temp) %in% c("ENO2"), ]

temp <- t(scale(t(temp)))
# temp <- melt(as.matrix(temp))


col_fun <- colorRamp2(c(-1, 0, 1), c("blue", "white", "red"))


pdf(here("gene_drug_vln/heatmap_SNR635_IgM.pdf"), width = 7, height = 6) 
Heatmap(
    temp,
    col = col_fun,
    name = "Expr",
    column_names_rot = 0,
    column_names_gp = gpar(fontsize = 20),
    column_names_centered = T,
    row_names_gp = gpar(fontsize = 15),
    # heatmap_legend_param = list(
    #     title = "log10(expr)",
    #     labels_gp = gpar(fontsize = 20),
    #     title_gp = gpar(fontsize = 25),
    #     legend_height = unit(10, "cm"),
    #     grid_width = unit(1, "cm")
    # ),
    column_title = "IgM",
    column_title_gp = gpar(fontsize = 30)
)
dev.off()
```


```{r}
library(bigSCale)
obj <- readRDS("03_each_cells/LUSC/Basal/seurat.rds")
cluster <- read.xlsx("03_each_cells/LUSC/Basal/markers_by_cluster.xlsx", rowNames=T)

# cluster <- cluster[cluster$avg_logFC > 0.25 & cluster$p_val_adj < 0.05, ]


msig <- tmodImportMSigDB("project_scripts/12_final_plots/msigdb_v6.2.xml")
registerDoMC(10)
network <- foreach(i = as.character(unique(cluster$ident)), .errorhandling = "pass") %dopar% {
    print(i)
    cells = rownames(obj@meta.data)[as.character(obj@meta.data$res.0.8) == i]
    
    res = list()
    res[["ident"]] = i
    
    res[["graph"]] = compute.network(
        expr.data = obj@raw.data[, cells],
        gene.names = rownames(obj@raw.data),
        quantile.p = 0.995
    )
    
    temp <- res[["graph"]]$centrality
    temp <- temp[order(temp$PAGErank, decreasing = T), ]
    res[["cerno"]] = tmodCERNOtest(rownames(temp), mset = msig)
    return(res)
}

saveRDS(network, "03_each_cells/LUSC/Basal/clusters")
```


```{r}

cerno = NULL
for(i in network) {
    temp <- i[["cerno"]]
    temp$ident = i[["ident"]]
    cerno = rbind(cerno, temp)
}


cerno <- cerno[cerno$adj.P.Val < 0.05 & cerno$AUC > 0.99, ]
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.