---
title: "cell_score"
author: "Zhang Yiming"
date: "2020/3/10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

root.dir = "LungCancer10x/"
full.path <- function(...) {
    paste(root.dir, ..., sep = "/")
}
```


```{r include=FALSE}
library(Seurat)
library(stringr)
library(doMC)
library(openxlsx)
library(ggplot2)
library(clusterProfiler)
library(org.Hs.eg.db)
library(ggrastr)
library(reshape2)
library(dplyr)
library(DESeq2)
library(ggpubr)
```


##  Scatter plot
```{r}
combat_edata <- readRDS(full.path("09_bulk/RNA_seq/DEGs/data/All_expr_combat.rds"))
bulk_meta <- readRDS(full.path("09_bulk/RNA_seq/DEGs/data/All_est_meta.rds"))
tree = read.csv(full.path("09_bulk/RNA_seq/MuSic/All_tree.csv"), row.names = 1, stringsAsFactors = F)
```


```{r}
markers <- read.csv(
    full.path("09_bulk/RNA_seq/MuSic/merged_markers.csv"), 
    row.names = 1, stringsAsFactors = F
)
expr <- readRDS(full.path("02_rds/all_cell_expr.rds"))
meta <- readRDS(full.path("11_CNV/meta.rds"))
meta$cell_short = as.character(meta$cell_short)
meta$cell_short[meta$cell_short == "ATII"] = "AT2"
meta$cell_short[meta$cell_short == "Epi"] = "AT1"
markers$Cells = as.character(markers$Cells)
markers$Cells[markers$Cells == "ATII"] = "AT2"
markers$Cells[markers$Cells == "Epi"] = "AT1"

immu_cells = c(
    "Tregs", "Mo", "Mφ","CD8", "Mast",
    "NK", "CD4", "DC", "B", "Gran"
)


cell_colors = ggthemes::tableau_color_pal(palette = "Tableau 20")(length(unique(meta$cell_short)) + 1)
set.seed(42)
names(cell_colors) <- sort(as.character(unique(meta$cell_short)))
cell_colors[["NE"]] = cell_colors[[length(cell_colors)]]


axis_size = 18
axis_title_size = 20
title_size = 25
legend.position="right"
legend_text_size = 15
legend_title_size = 20

my_theme <- ggplot2::theme(
    aspect.ratio=1,
    plot.title = element_text(hjust = 0.5, face="bold", size = title_size),
    legend.position = legend.position,
    panel.background = element_blank(),
    axis.text = element_text(size = axis_size),
    axis.text.x = element_text(size = axis_size),
    axis.title = element_text(size = axis_title_size),
    legend.text = element_text(size = legend_text_size),
    legend.title = element_text(size = legend_title_size)
)
```


## Test LUAD samples
```{r}
luad <- expr[, rownames(meta)[str_detect(meta$Disease, "LUAD") & !meta$cell_short %in% immu_cells]]
# temp_expr <- scale(t(luad[markers$Markers[markers$Cells %in% c("ATII", "NE", "Fib", "Epi")], ]))
# temp_expr <- temp_expr - min(temp_expr)

luad = apply(luad, 2, function(col) { col * 10^6 / sum(col) })
 
luad <- melt(luad[markers$Markers[markers$Cells %in% c("AT2", "NE", "Fib", "AT1")], ])

# colnames(luad) <- c("Var2", "Var1", "value")
luad$Var1 <- as.character(luad$Var1)
luad$Var2 <- as.character(luad$Var2)

luad <- merge(
    luad, markers[markers$Cells %in% c("AT2", "NE", "Fib", "AT1"), ],
    by.x = "Var1", by.y = "Markers"
)
```


#### Calculate SC score
```{r}
luad <- luad %>%
    group_by(Var2, Cells) %>%
    mutate(value = mean(value)) %>%
    dplyr::select(Var2, value, Cells) %>%
    unique() %>%
    as.data.frame()
```


#### Calculate D
```{R}
luad <- dcast(luad, Var2~Cells, value.var = "value", fun.aggregate = mean, fill = 0)


luad$y <- apply(as.data.frame(luad), 1, function(row) {
    row = row[!names(row) %in% c("Var2", "Cell", "Disease", "x", "y")]
    row1 = as.numeric(row)
    names(row1) <- names(row)
    row = row1
    
    if (sum(row) != 0 && names(row)[which(row == max(row))] %in% c("NE", "Epi")) {
        val = row["AT1"] - row["NE"]
    } else {
        val = row["AT2"] - row["Fib"]
    }
    
    weight = 1
    if (val < 0) {
        weight = -1
    }

    return(weight * log2(abs(val) + 1))
   
})

luad$x <- apply(as.data.frame(luad), 1, function(row) {
    row = row[!names(row) %in% c("Var2", "Cell", "Disease", "x", "y")]
    row1 = as.numeric(row)
    names(row1) <- names(row)
    row = row1
    
    if (sum(row) != 0 && names(row)[which(row == max(row))] %in% c("AT2", "AT1")) {
        val = row["AT2"] - row["AT1"]
    } else {
        val = row["Fib"] - row["NE"]
    }
    
    weight = 1
    if (val < 0) {
        weight = -1
    }
    
    return(weight * log2(abs(val) + 1))
    
})
```


```{r fig.height=4, fig.width=5}
luad$Cell = meta[luad$Var2, "cell_short"]
luad$Disease = meta[luad$Var2, "Disease"]
luad$Disease[str_detect(luad$Disease, "Normal")] = "NL(AD)"


annotations <- data.frame(
    xpos = c(-Inf,-Inf,Inf,Inf),
    ypos =  c(-Inf, Inf,-Inf,Inf),
    annotateText = c("NE", "AT1", "Fib", "AT2"),
    hjustvar = c(0,0,1,1) ,
    vjustvar = c(0,1,0,1)
)

p <- ggplot() +
    geom_point_rast(
        data = as.data.frame(luad), 
        aes(x=x, y=y, color=Cell, shape=Disease)
    ) +
    geom_text(
        data=annotations,
        aes(
            x=xpos, y=ypos,
            hjust=hjustvar, vjust=vjustvar,
            label=annotateText
        ),
        size = 10
    ) +
    labs(x = "", y = "") +
    theme_bw(base_family = "Arial Unicode MS") +
    my_theme +
    theme(
        legend.title = element_text(size = 15),
        legend.text = element_text(size = 12),
        legend.spacing.y = unit(0.05, 'cm'),
        panel.grid = element_blank()
    ) +
    guides(
        color = guide_legend(override.aes = list(size = 3.5)),
        shape = guide_legend(override.aes = list(size = 3.5))
    ) +
    scale_color_manual(values = cell_colors)

p

ggsave(
    filename = full.path("14_cycling/cell_score/LUAD.pdf"),
    plot = p, device = cairo_pdf,
    height = 4,
    width = 5
)
```





## Test LUSC samples
```{r}
lusc <- expr[, rownames(meta)[str_detect(meta$Disease, "LUSC") & !meta$cell_short %in% immu_cells]]
# temp_expr <- scale(t(lusc[markers$Markers[markers$Cells %in% c("ATII", "NE", "Fib", "Basal")], ]))
# temp_expr <- temp_expr - min(temp_expr)

lusc = apply(lusc, 2, function(col) { col * 10^6 / sum(col) })

lusc <- melt(lusc[markers$Markers[markers$Cells %in% c("AT2", "AT1", "Fib", "Basal")], ])
# colnames(lusc) <- c("Var2","Var1", "value")
lusc$Var1 <- as.character(lusc$Var1)
lusc$Var2 <- as.character(lusc$Var2)
```


```{r}
lusc <- merge(
    lusc, markers[markers$Cells %in% c("AT2", "AT1", "Fib", "Basal"), ],
    by.x = "Var1", by.y = "Markers"
)

lusc <- lusc %>%
    group_by(Var2, Cells) %>%
    mutate(value = mean(value)) %>%
    dplyr::select(Var2, value, Cells) %>%
    unique() %>%
    as.data.frame()
```


#### Calculate D
```{R}
lusc <- dcast(lusc, Var2~Cells, value.var = "value", fun.aggregate = mean, fill = 0)

lusc$y = 0
lusc$x = 0

for (i in 1:nrow(lusc)) {
    row = lusc[i, !colnames(lusc) %in% c("Var2", "Cell", "Disease")]
    row1 = as.numeric(row)
    names(row1) <- names(row)
    row = row1
    
    if (sum(row) != 0 && names(row)[which(row == max(row))] %in% c("AT2", "AT1")) {
        val = as.numeric(row["AT2"] - row["AT1"])
    } else {
        val = as.numeric(row["Basal"] - row["Fib"])
    }
    
    weight = 1
    if (val < 0) {
        weight = -1
    }
    lusc[i, "y"] = weight * log2(abs(val) + 1)
    
    
    if (sum(row) != 0 && names(row)[which(row == max(row))] %in% c("Basal", "AT2")) {
        val = row["Basal"] - row["AT2"]
    } else {
        val = row["Fib"] - row["AT1"]
    }
    
    weight = 1
    if (val < 0) {
        weight = -1
    }
    
    lusc[i, "x"] = weight * log2(abs(val) + 1)
}


```


```{r fig.height=4, fig.width=5}
lusc$Cell = meta[lusc$Var2, "cell_short"]
lusc$Disease = meta[lusc$Var2, "Disease"]
lusc$Disease[str_detect(lusc$Disease, "Normal")] = "NL(SC)"


annotations <- data.frame(
    xpos = c(-Inf,-Inf,Inf,Inf),
    ypos =  c(-Inf, Inf,-Inf,Inf),
    annotateText = c("AT1", "AT2", "Fib", "Basal"),
    hjustvar = c(0,0,1,1) ,
    vjustvar = c(0,1,0,1)
)

p <- ggplot() +
    geom_point_rast(
        data = as.data.frame(lusc), 
        aes(x=x, y=y, color=Cell, shape=Disease)
    ) +
    # geom_density2d(
    #     data = as.data.frame(lusc), 
    #     aes(x=x, y=y, color=Cell, shape=Disease),
    #     geom="raster"
    # ) +
    geom_text(
        data=annotations,
        aes(
            x=xpos, y=ypos,
            hjust=hjustvar, vjust=vjustvar,
            label=annotateText
        ),
        size = 10
    ) +
    labs(x = "", y = "") +
    theme_bw(base_family = "Arial Unicode MS") +
    my_theme +
    theme(
        legend.title = element_text(size = 15),
        legend.text = element_text(size = 12),
        legend.spacing.y = unit(0.05, 'cm'),
        panel.grid = element_blank()
    ) +
    guides(
        color = guide_legend(override.aes = list(size = 3.5)),
        shape = guide_legend(override.aes = list(size = 3.5))
    ) +
    scale_color_manual(values = cell_colors)

p


ggsave(
    filename = full.path("14_cycling/cell_score/LUSC.pdf"),
    plot = p, device = cairo_pdf,
    height = 4,
    width = 5
)
```



----
## MSC score


```{r}
umap <- read.csv(full.path("02_rds/umap.csv"), row.names = 1, stringsAsFactors = F)
```


### Make feature plot to check MSC markers
```{R}
genes_from_SR <- c(
    "NESH", "CXCR7", "GPNMB", "MYBPH", "SVEP1",
    "ARHGAP6", "TSPEAR", "PIK3CG", "ABL2", "NCAM1",
    "ZNF145", "FKBP5", "DPT", "OMD", "ID4", "CRYAB",
    "SORT1", "CTNNB1", "ZEB", "LPL", "FABP4", "PDK4",
    "ACDC", "CCN3", "NOV", "CCN4", "WISP1", "CCN5", "WISP2",
    "ADAMTS-5", "CXCL12", "FLT3L", "IL-3", "TPO", "KITL"
)

intersect(genes_from_SR, rownames(expr))
```

```{r}
genes_from_PeerJ = c()
temp <- read.xlsx(full.path("14_cycling/TableS6-DEG.xlsx"), sheet = 1)
temp = bitr(temp$ENSEMBL.ID, fromType = "ENSEMBL", toType = "SYMBOL", OrgDb = org.Hs.eg.db)
genes_from_PeerJ = c(genes_from_PeerJ, temp$SYMBOL)

temp <- read.xlsx(full.path("14_cycling/TableS6-DEG.xlsx"), sheet = 2)
temp = bitr(temp$ENSEMBL.ID, fromType = "ENSEMBL", toType = "SYMBOL", OrgDb = org.Hs.eg.db)

genes_from_PeerJ = intersect(genes_from_PeerJ, temp$SYMBOL)

intersect(genes_from_PeerJ, rownames(expr))
```

```{R}
intersect(genes_from_PeerJ, genes_from_SR)
```


## Make feature plot to check the genes
```{R}
make_umap_features <- function(
    coord,
    expr,
    gene,
    pt.size = 0.1, 
    alpha = 0.5,
    text_size = 10,
    axis_size = 18,
    axis_title_size = 20,
    title_size = 25,
    group_by = "",
    label = "cell_name",
    legend.position = "none"
) {
    coord = as.data.frame(coord)
    coord$Expr <- as.numeric(as.character(expr[gene, rownames(coord)]))
    coord <- coord[order(coord$Expr, decreasing = F), ]
    ggplot(coord, aes(x=UMAP_1, y=UMAP_2, color = Expr)) +
        geom_point_rast(size = pt.size) +
        scale_color_gradient(low = "grey", high = "blue") +
        theme_bw() +
        labs(
            title = gene
        ) +
        theme(
            plot.title = element_text(hjust = 0.5, face="bold", size = title_size),
            legend.position = legend.position,
            axis.text = element_text(size = axis_size),
            axis.title = element_text(size = axis_title_size)
        )
}
```


#### Scale expression of target genes
```{R}
temp_expr <- 
temp_expr <- t(scale(t(temp_expr)))
```



```{r fig.height=5, fig.width=5}
registerDoMC(10)
temp = foreach (i = unique(c(genes_from_PeerJ, genes_from_SR)), .errorhandling = "pass") %dopar% {
    p <- make_umap_features(as.data.frame(umap), as.data.frame(temp_expr), i)
    
    ggsave(
        filename = full.path("14_cycling/cell_score/features", paste0(i, ".pdf")),
        plot = p,
        width = 5,
        height = 5
    )
}
```


### Calculate MSC score
```{R}
temp_expr = apply(expr[unique(c(genes_from_PeerJ, genes_from_SR)), ], 1, function(col) {
    col * 10^6 / sum(col)
})


temp_expr <- melt(as.matrix(temp_expr))
temp_expr <- na.omit(temp_expr)

temp_expr <- temp_expr %>%
    group_by(Var1) %>%
    mutate(value = mean(value)) %>%
    dplyr::select(Var1, value) %>%
    unique() %>%
    as.data.frame()

temp_expr$UMAP1 = coord[temp_expr$Var1, "UMAP_1"]
temp_expr$UMAP2 = coord[temp_expr$Var1, "UMAP_2"]
```


```{R fig.height=5, fig.width=6}
ggplot(temp_expr, aes(x=UMAP1, y=UMAP2, color = log10(value + 1))) +
    geom_point_rast(size = 0.1) +
    scale_color_gradient(low = "grey", high = "blue") +
    theme_bw() +
    theme(
        plot.title = element_text(hjust = 0.5, face="bold", size = 25),
        legend.position = "right",
        axis.text = element_text(size = 18),
        axis.title = element_text(size = 20)
    )
```

```{r}
wb = createWorkbook()
addWorksheet(wb, "LUAD")
writeData(wb, 1, luad)
addWorksheet(wb, "LUSC")
writeData(wb, 2, lusc)
addWorksheet(wb, "MSC")
writeData(wb, 3, temp_expr)
saveWorkbook(wb, file = full.path("14_cycling/cell_score/scores.xlsx"))
```


### plot with MSC score
```{R fig.height=5, fig.width=6}
rownames(temp_expr) <- temp_expr$Var1

luad$MSC = temp_expr[luad$Var2, "value"]


ggplot(luad, aes(x=x, y=y, color = log10(MSC + 1))) +
    geom_point_rast(size = 0.1) +
    scale_color_gradient(low = "grey", high = "blue") +
    theme_bw() +
    theme(
        plot.title = element_text(hjust = 0.5, face="bold", size = 25),
        legend.position = "right",
        axis.text = element_text(size = 18),
        axis.title = element_text(size = 20)
    )
```

```{r}
lusc$MSC = temp_expr[lusc$Var2, "value"]


ggplot(lusc, aes(x=x, y=y, color = log10(MSC + 1))) +
    geom_point_rast(size = 0.1) +
    scale_color_gradient(low = "grey", high = "blue") +
    theme_bw() +
    theme(
        plot.title = element_text(hjust = 0.5, face="bold", size = 25),
        legend.position = "right",
        axis.text = element_text(size = 18),
        axis.title = element_text(size = 20)
    )
```

### violin plots of MSC score
```{r fig.height=12, fig.width=15}
temp_expr$Cell = meta[temp_expr$Var1, "cell_short"]
temp_expr$Disease = meta[temp_expr$Var1, "Disease"]
    
ggplot(temp_expr, aes(x=Disease, y=log2(value + 1), fill = Disease)) +
    geom_violin() +
    facet_wrap(.~Cell, scales = "free") +
    theme(axis.text.x = element_blank())
```


#### ATII
```{R}
obj <- readRDS(full.path("03_each_cells/Batch/ATII/LUAD/seurat.rds"))

obj@meta.data$Clt = obj@meta.data$res.0.1
obj@meta.data$MSC = temp_expr[rownames(obj@meta.data), "value"]


temp <- obj@meta.data %>%
    group_by(Clt) %>%
    mutate(v = mean(MSC)) %>%
    dplyr::select(Clt, v) %>%
    unique() %>% as.data.frame()

temp = temp[order(temp$v, decreasing = T), ]

my_comparisons = list()
for(i in 1:(nrow(temp) - 1)) {
    for (j in (i + 1): nrow(temp)) {
        my_comparisons[[length(my_comparisons) + 1]] = c(temp$Clt[i], temp$Clt[j])
    }
}

obj@meta.data$Clt = factor(obj@meta.data$Clt, levels = temp$Clt)

ggviolin(
    data = obj@meta.data,
    x="Clt", y = "MSC",
    fill="Clt", add = "boxplot", 
    add.params = list(fill = "white")
) +
    scale_y_log10() +
    stat_compare_means(comparisons = my_comparisons) + # Add pairwise comparisons p-value
    stat_compare_means(label.y = 5) 
```


#### Basal
```{R fig.height=8, fig.width=12}
obj <- readRDS(full.path("03_each_cells/Batch/Basal/LUSC/seurat.rds"))

obj@meta.data$Clt = as.character(as.numeric(as.character(obj@meta.data$res.0.03)) + 1)
obj@meta.data$MSC = temp_expr[rownames(obj@meta.data), "value"]


temp <- obj@meta.data %>%
    group_by(Clt) %>%
    mutate(v = mean(MSC)) %>%
    dplyr::select(Clt, v) %>%
    unique() %>% as.data.frame()

temp = temp[order(temp$v, decreasing = T), ]

obj@meta.data$Clt = factor(obj@meta.data$Clt, levels = temp$Clt)

my_comparisons = list()
for(i in 1:(nrow(temp) - 1)) {
    for (j in (i + 1): nrow(temp)) {
        my_comparisons[[length(my_comparisons) + 1]] = c(temp$Clt[i], temp$Clt[j])
    }
}

ggviolin(
    data = obj@meta.data,
    x="Clt", y = "MSC",
    fill="Clt", add = "boxplot", 
    add.params = list(fill = "white")
) +
    scale_y_log10() +
    stat_compare_means(comparisons = my_comparisons) + # Add pairwise comparisons p-value
    stat_compare_means(label.y = 5) 
```


#### NE
```{R fig.height=6, fig.width=8}
obj <- readRDS(full.path("03_each_cells/Batch/NE/LUAD/seurat.rds"))
obj@meta.data$Clt = as.character(as.numeric(as.character(obj@meta.data$res.0.1)) + 1)
obj@meta.data$MSC = temp_expr[rownames(obj@meta.data), "value"]

temp <- obj@meta.data %>%
    group_by(Clt) %>%
    mutate(v = mean(MSC)) %>%
    dplyr::select(Clt, v) %>%
    unique() %>% as.data.frame()

temp = temp[order(temp$v, decreasing = T), ]

obj@meta.data$Clt = factor(obj@meta.data$Clt, levels = temp$Clt)

my_comparisons = list()
for(i in 1:(nrow(temp) - 1)) {
    for (j in (i + 1): nrow(temp)) {
        my_comparisons[[length(my_comparisons) + 1]] = c(temp$Clt[i], temp$Clt[j])
    }
}

ggviolin(
    data = obj@meta.data,
    x="Clt", y = "MSC",
    fill="Clt", add = "boxplot", 
    add.params = list(fill = "white")
) +
    scale_y_log10() +
    stat_compare_means(comparisons = my_comparisons) + # Add pairwise comparisons p-value
    stat_compare_means(label.y = 7) 
```


#### Epi
```{R fig.height=6, fig.width=8}
obj <- readRDS(full.path("03_each_cells/Batch/Epi/LUAD/seurat.rds"))
obj@meta.data$Clt = as.character(as.numeric(as.character(obj@meta.data$res.0.1)) + 1)
obj@meta.data$MSC = temp_expr[rownames(obj@meta.data), "value"]

my_comparisons = list()
for(i in 1:(nrow(temp) - 1)) {
    for (j in (i + 1): nrow(temp)) {
        my_comparisons[[length(my_comparisons) + 1]] = c(temp$Clt[i], temp$Clt[j])
    }
}

ggviolin(
    data = obj@meta.data,
    x="Clt", y = "MSC",
    fill="Clt", add = "boxplot", 
    add.params = list(fill = "white")
) +
    scale_y_log10() +
    stat_compare_means(comparisons = my_comparisons) + # Add pairwise comparisons p-value
    stat_compare_means(label.y = 5) 
```


#### Fib
```{R}
obj <- readRDS(full.path("03_each_cells/Batch/Fib/LUAD/seurat.rds"))
obj@meta.data$Clt = as.character(as.numeric(as.character(obj@meta.data$res.0.1)) + 1)
obj@meta.data$MSC = temp_expr[rownames(obj@meta.data), "value"]

temp <- obj@meta.data %>%
    group_by(Clt) %>%
    mutate(v = mean(MSC)) %>%
    dplyr::select(Clt, v) %>%
    unique() %>% as.data.frame()

temp = temp[order(temp$v, decreasing = T), ]

obj@meta.data$Clt = factor(obj@meta.data$Clt, levels = temp$Clt)

my_comparisons = list()
for(i in 1:(nrow(temp) - 1)) {
    for (j in (i + 1): nrow(temp)) {
        my_comparisons[[length(my_comparisons) + 1]] = c(temp$Clt[i], temp$Clt[j])
    }
}

ggviolin(
    data = obj@meta.data,
    x="Clt", y = "MSC",
    fill="Clt", add = "boxplot", 
    add.params = list(fill = "white")
) +
    scale_y_log10() +
    stat_compare_means(comparisons = my_comparisons) + # Add pairwise comparisons p-value
    stat_compare_means(label.y = 4) 
```


#### Fib
```{R fig.height=6, fig.width=8}
obj <- readRDS(full.path("03_each_cells/Batch/Fib/LUSC/seurat.rds"))
obj@meta.data$Clt = as.character(as.numeric(as.character(obj@meta.data$res.0.1)) + 1)
obj@meta.data$MSC = temp_expr[rownames(obj@meta.data), "value"]

temp <- obj@meta.data %>%
    group_by(Clt) %>%
    mutate(v = mean(MSC)) %>%
    dplyr::select(Clt, v) %>%
    unique() %>% as.data.frame()

temp = temp[order(temp$v, decreasing = T), ]

obj@meta.data$Clt = factor(obj@meta.data$Clt, levels = temp$Clt)

my_comparisons = list()
for(i in 1:(nrow(temp) - 1)) {
    for (j in (i + 1): nrow(temp)) {
        my_comparisons[[length(my_comparisons) + 1]] = c(temp$Clt[i], temp$Clt[j])
    }
}

ggviolin(
    data = obj@meta.data,
    x="Clt", y = "MSC",
    fill="Clt", add = "boxplot", 
    add.params = list(fill = "white")
) +
    scale_y_log10() +
    stat_compare_means(comparisons = my_comparisons) + # Add pairwise comparisons p-value
    stat_compare_means(label.y = 7) 
```


### Score on bulk sample
```{R}
merged_est <- readRDS(full.path("09_bulk/RNA_seq/MuSic/All_est.rds"))
meta <- readRDS(full.path("09_bulk/RNA_seq/MuSic/All_est_meta.rds"))

final_tree <- read.csv(full.path("09_bulk/RNA_seq/MuSic/All_tree.csv"), row.names = 1, stringsAsFactors = F)
```


#### LUAD
```{r}
luad <- as.data.frame(t(merged_est[, rownames(final_tree)[final_tree$Disease == "LUAD"]]))

luad$y <- apply(as.data.frame(luad), 1, function(row) {
    row = row[!names(row) %in% c("Var2", "Cell", "Disease", "x", "y")]
    row1 = as.numeric(row)
    names(row1) <- names(row)
    row = row1
    
    if (sum(row) != 0 && names(row)[which(row == max(row))] %in% c("NE", "Epi")) {
        val = row["Epi"] - row["NE"]
    } else {
        val = row["ATII"] - row["Fib"]
    }
    
    weight = 1
    if (val < 0) {
        weight = -1
    }

    return(weight * log2(abs(val) + 1))
   
})

luad$x <- apply(as.data.frame(luad), 1, function(row) {
    row = row[!names(row) %in% c("Var2", "Cell", "Disease", "x", "y")]
    row1 = as.numeric(row)
    names(row1) <- names(row)
    row = row1
    
    if (sum(row) != 0 && names(row)[which(row == max(row))] %in% c("ATII", "Epi")) {
        val = row["ATII"] - row["Epi"]
    } else {
        val = row["Fib"] - row["NE"]
    }
    
    weight = 1
    if (val < 0) {
        weight = -1
    }
    
    return(weight * log2(abs(val) + 1))
    
})
```


```{r fig.height=6, fig.width=6}
luad$Clt = as.character(sapply(final_tree[row.names(luad), "Clt"], function(x) {
    groups = c("NE-high", "Epi-high", "ATII-high", "Fib-high")
    
    groups[x]
}))


annotations <- data.frame(
    xpos = c(-Inf,-Inf,Inf,Inf),
    ypos =  c(-Inf, Inf,-Inf,Inf),
    annotateText = c("NE", "Epi", "Fib", "ATII"),
    hjustvar = c(0,0,1,1) ,
    vjustvar = c(0,1,0,1)
)

p <- ggplot(data = as.data.frame(luad)) +
    geom_point_rast(aes(x=x, y=y, color=Clt)) +
    stat_density2d(aes(x=x, y=y, color=Clt), data = as.data.frame(luad)) +
    geom_text(
        data=annotations,
        aes(
            x=xpos, y=ypos,
            hjust=hjustvar, vjust=vjustvar,
            label=annotateText
        ),
        size = 10
    ) +
    labs(x = "", y = "", color = "") +
    theme_bw() +
    my_theme +
    theme(
        legend.position = c(0.5, 0.12),
        legend.background = element_blank()
    ) +
    guides(col = guide_legend(ncol = 2, override.aes = list(size = 5)))

print(p)


ggsave(
    filename = full.path("14_cycling/cell_score/LUAD_bulk.pdf"),
    plot = p,
    height = 6,
    width = 6
)
```

#### LUSC
```{r}
lusc <- as.data.frame(t(merged_est[, rownames(final_tree)[final_tree$Disease == "LUSC"]]))

lusc$y = 0
lusc$x = 0

for (i in 1:nrow(lusc)) {
    row = lusc[i, !colnames(lusc) %in% c("Var2", "Cell", "Disease")]
    row1 = as.numeric(row)
    names(row1) <- names(row)
    row = row1
    
    if (sum(row) != 0 && names(row)[which(row == max(row))] %in% c("ATII", "Epi")) {
        val = as.numeric(row["ATII"] - row["Epi"])
    } else {
        val = as.numeric(row["Basal"] - row["Fib"])
    }
    
    weight = 1
    if (val < 0) {
        weight = -1
    }
    lusc[i, "y"] = weight * log2(abs(val) + 1)
    
    
    if (sum(row) != 0 && names(row)[which(row == max(row))] %in% c("Basal", "ATII")) {
        val = row["Basal"] - row["ATII"]
    } else {
        val = row["Fib"] - row["Epi"]
    }
    
    weight = 1
    if (val < 0) {
        weight = -1
    }
    
    lusc[i, "x"] = weight * log2(abs(val) + 1)
}
```


```{r fig.height=6, fig.width=6}
lusc$Clt = as.character(sapply(final_tree[row.names(lusc), "Clt"], function(x) {
    groups = c("Fib-high", "ATII-high", "Basal-high", "Basal-Fib hybrid", "Epi-Fib-ATII hybrid")
    
    groups[x]
}))


annotations <- data.frame(
    xpos = c(-Inf,-Inf,Inf,Inf),
    ypos =  c(-Inf, Inf,-Inf,Inf),
    annotateText = c("Epi", "ATII", "Fib", "Basal"),
    hjustvar = c(0,0,1,1) ,
    vjustvar = c(0,1,0,1)
)

p <- ggplot() +
    geom_point_rast(
        data = lusc, 
        aes(x=x, y=y, color=Clt)
    ) +
    geom_density2d(
        data = lusc, 
        aes(x=x, y=y, color=Clt)
    ) +
    geom_text(
        aes(
            x=xpos, y=ypos,
            hjust=hjustvar, vjust=vjustvar,
            label=annotateText
        ),
        data=annotations,
        size = 10
    ) +
    labs(x = "", y = "", color = "") +
    theme_bw() +
    my_theme +
    theme(
        legend.position = c(0.5, 0.1),
        legend.background = element_blank(),
        legend.spacing.x = unit(0, 'cm')
    ) +
    guides(col = guide_legend(ncol = 2, override.aes = list(size = 5)))


print(p)


ggsave(
    filename = full.path("14_cycling/cell_score/LUSC_bulk.pdf"),
    plot = p,
    height = 6,
    width = 6
)
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.