---
title: "Cell cycle"
author: "Zhang Yiming"
date: "2020/2/29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

root.dir = "LungCancer10x/"

full.path <- function(...) { return(paste(root.dir, ..., sep = "/")) }
```


## Load packages

```{r cars, include=FALSE}
library(dplyr)
library(stringr)
library(ComplexHeatmap)
library(circlize)
library(reshape2)
library(wesanderson)
library(ggplot2)

options(stringsAsFactor = F)
```

# Cell cycle using Seurat3

```{r include=FALSE}
library(Seurat)
library(reshape2)
library(dplyr)
library(ComplexHeatmap)
library(circlize)
```


```{r}
expr <- readRDS(full.path("02_rds/all_cell_expr.rds"))
meta <- readRDS(full.path("02_rds/meta.rds"))
```


```{R fig.height=10, fig.width=10}
temp = meta %>%
    group_by(cell_short, Disease, Phase) %>%
    add_tally() %>%
    dplyr::select(cell_short, Disease, Phase, n) %>%
    unique() %>%
    group_by(cell_short, Disease) %>%
    mutate(p = n / sum(n) * 100) %>%
    as.data.frame()

p <- ggplot(temp[temp$Disease %in% c("LUAD", "LUSC", "LUAD_Normal", "LUSC_Normal"), ], 
       aes(x=cell_short, y=p, fill=Phase, label = round(p, 2))) +
        geom_bar(stat = "identity", position = position_dodge()) +
        geom_text(angle = 90, position = position_dodge(width = 1)) +
        facet_grid(Disease~.) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))

ggsave(filename = full.path("14_cycling/cycling_percentage_barplot.pdf"),
       plot=p,
       width = 10,
       height = 10
       )
```


## Cell cycle
```{r eval=FALSE, include=FALSE}
# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can
# segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

# Create our Seurat object and complete the initalization steps
marrow <- CreateSeuratObject(counts = expr[, rownames(meta)], meta.data = meta)

# store mitochondrial percentage in object meta data
marrow <- PercentageFeatureSet(marrow, pattern = "^MT-", col.name = "percent.mt")

# run sctransform
marrow <- SCTransform(marrow, vars.to.regress = c("percent.mt", "Source"), verbose = FALSE)


marrow <- RunPCA(marrow, features = VariableFeatures(marrow), ndims.print = 0, nfeatures.print = 0)


marrow <- CellCycleScoring(marrow, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

# view cell cycle scores and phase assignments
head(marrow[[]])

temp = marrow[[]]

saveRDS(temp, full.path("09_bulk/RNA_seq/DEGs/Res/cell_cycle.rds"))
```



### Calcuate G1/G2 mean score
```{R  eval=FALSE, include=FALSE}

cell_cycle = data.frame(
    Gene=c(cell_cycle[["G1"]], cell_cycle[["G2"]]),
    Phase=c(rep("G1", length(cell_cycle[["G1"]])), rep("G2", length(cell_cycle[["G2"]])))
)

row.names(cell_cycle) <- make.unique(as.character(cell_cycle$Gene))

expr <- melt(as.matrix(
    expr[unique(as.character(cell_cycle$Gene)), ]
))

expr <- na.omit(expr)


expr$Phase = cell_cycle[expr$Var1, "Phase"]


expr <- expr %>%
    group_by(Var2, Phase) %>%
    mutate(value = mean(value)) %>%
    dplyr::select(Var2, value, Phase) %>%
    unique() %>%
    as.data.frame()

expr <- dcast(expr, Var2~Phase, fun.aggregate = mean, value.var = "value", fill=0)

for (i in c("S.Score", "G2M.Score", "Phase")) {
    expr[, i] <- meta[expr$Var2, i]
}
```


```{r  eval=FALSE, include=FALSE}
plist = list()
plist[[1]] = ggplot(expr, aes(x=log2(G1 + 1), y=G2M.Score)) +
    geom_point(size = 0.5) +
    geom_density2d() +
    annotate("text", x=4, y=1.5, label = cor(log2(expr$G1 + 1), y=expr$G2M.Score))


plist[[2]] = ggplot(expr, aes(x=log2(G2 + 1), y=G2M.Score)) +
    geom_point(size = 0.5) +
    geom_density2d() +
    annotate("text", x=4, y=1.5, label = cor(log2(expr$G2 + 1), y=expr$G2M.Score))

plist[[3]] = ggplot(expr, aes(x=log2(G1 + 1), y=S.Score)) +
    geom_point(size = 0.5) +
    geom_density2d() +
    annotate("text", x=4, y=1.5, label = cor(log2(expr$G1 + 1), y=expr$S.Score))

plist[[4]] = ggplot(expr, aes(x=log2(G2 + 1), y=S.Score)) +
    geom_point(size = 0.5) +
    geom_density2d() +
    annotate("text", x=4, y=1.5, label = cor(log2(expr$G2 + 1), y=expr$S.Score))


cowplot::plot_grid(plotlist = plist, ncol = 2)
```


```{r fig.height=4, fig.width=6, eval=FALSE, include=FALSE}
ggplot(expr, aes(x=S.Score, y=G2M.Score, color=Phase)) +
    geom_point()
```


```{R eval=FALSE, include=FALSE}
ggplot(expr, aes(x=log2(G1 + 1), y=log2(G2 + 1), color=Phase)) +
    geom_point()
```


### Cell score
```{R}
markers <- read.csv(full.path("09_bulk/RNA_seq/MuSic/merged_markers.csv"), row.names = 1, stringsAsFactors = F)
rownames(markers) <- make.unique(markers$Markers)


make_cell_score <- function(expr, meta, markers, seed = 42) {
    expr <- expr[as.character(unique(markers$Markers)), ]
    temp_expr = melt(as.matrix(expr))
    temp_expr$Cell = markers[temp_expr$Var1, "Cells"]
    temp_expr$cell_short = meta[temp_expr$Var2, "cell_short"]
    temp_expr$Disease = meta[temp_expr$Var2, "Disease"]
    
    set.seed(seed)
    temp_cell_score <- temp_expr %>%
        group_by(Cell, Var2) %>%
        mutate(value = mean(value)) %>%
        dplyr::select(Var2, value, Cell) %>%
        unique() %>%
        as.data.frame()
    
    temp_cell_score <- dcast(
        temp_cell_score, Cell~Var2, 
        fun.aggregate = mean, value.var = "value",
        fill = 0
    )
    
    temp_cell_score <- as.data.frame(temp_cell_score)
    temp_cell_score <- temp_cell_score[!is.na(temp_cell_score$Cell), ]
    rownames(temp_cell_score) <- as.character(temp_cell_score$Cell)
    temp_cell_score <- temp_cell_score[, colnames(temp_cell_score) != "Cell"]
    # temp_cell_score_scale = scale(temp_cell_score)
    
    temp_cell_score <- temp_cell_score[rownames(temp_cell_score) != "T", ]
}

```

```{r}
meta <- readRDS(full.path("02_rds/meta.rds"))
```

```{r}
col_fun = colorRamp2(c(-1, 0, 1), c("blue", "white", "red"))


immu = c(
    "B", "CD4", "CD8", "DC",
    "Gran", "Mast", "Mo", "NK",
    "Tregs", "Mφ"
)


### sort cells
meta$isImmu <- meta$cell_short %in% immu
meta <- meta[order(meta$cell_short, meta$isImmu), ]

cells <- unique(meta$cell_short)

# cell_color = wes_palette("Moonrise3", length(cell_types), type = "continuous")
cell_color = c(
    wes_palette("Darjeeling1"), 
    wes_palette("Moonrise3"),
    wes_palette("Darjeeling2"),
    wes_palette("Moonrise1")
)[1:length(cells)]

names(cell_color) = cells

cell_cycle_color = as.character(wes_palette("Royal1"))
names(cell_cycle_color) <- c("Non-Cycling", "G2M", "S", "Cycling")


set.seed(42)
for (i in unique(meta$Disease)) {
    temp_meta = meta[meta$Disease == i, ]
    
    temp_meta <- temp_meta %>%
        filter(!isImmu) %>%
        group_by(cell_short) %>%
        sample_n(500, replace = T) %>%
        unique() %>%
        as.data.frame()
    
    rownames(temp_meta) <- temp_meta$Cells
    
    temp_meta = temp_meta[order(temp_meta$cell_short), ]
    temp_meta$Phase = as.character(temp_meta$Phase)
    temp_meta$Phase[temp_meta$Phase == "G1"] = "Non-Cycling"
    temp_meta$Phase = factor(temp_meta$Phase, levels=c("Non-Cycling", "S", "G2M"))
    
    ta = HeatmapAnnotation(
        Cycle = temp_meta$Phase,
        Cell = temp_meta$cell_short,
        col = list(
            Cell=cell_color,
            Cycle=cell_cycle_color
        ),
        show_legend = T,
        show_annotation_name = T
    )
    
    h <- Heatmap(
        as.matrix(temp_cell_score[!rownames(temp_cell_score) %in% immu, rownames(temp_meta)]),
        name = "Score",
        col = col_fun,
        border = T,
        cluster_rows = F,
        cluster_columns = F,
        show_column_names = F,
        column_split = temp_meta$Phase,
        column_title = i,
        top_annotation = ta
    )
    
    draw(h)
}
```



```{r}

col_fun = colorRamp2(c(-1, 0, 1), c("blue", "white", "red"))


immu = c(
    "B", "CD4", "CD8", "DC",
    "Gran", "Mast", "Mo", "NK",
    "Tregs", "Mφ"
)


### sort cells
meta$isImmu <- meta$cell_short %in% immu
meta <- meta[order(meta$cell_short, meta$isImmu), ]

cells <- unique(meta$cell_short)

# cell_color = wes_palette("Moonrise3", length(cell_types), type = "continuous")
cell_color = c(
    wes_palette("Darjeeling1"), 
    wes_palette("Moonrise3"),
    wes_palette("Darjeeling2"),
    wes_palette("Moonrise1")
)[1:length(cells)]

names(cell_color) = cells

cell_cycle_color = as.character(wes_palette("Royal1"))
names(cell_cycle_color) <- c("Non-Cycling", "G2M", "S", "Cycling")

make_cell_score_seurat <- function(obj, meta, markers, scale.data = F, use.sum = F, seed=42) {
    expr = obj@raw.data
    
    if (scale.data) {
        expr = obj@scale.data
    }
    
    expr <- expr[as.character(unique(markers$gene)), ]
    temp_expr = melt(as.matrix(expr))
    temp_expr$Clt = markers[temp_expr$Var1, "ident"]
    
    set.seed(seed) 
    temp_cell_score <- temp_expr %>%
        group_by(Clt, Var2) %>%
        mutate(value1 = sum(value)) %>%
        group_by(Clt, Var2) %>%
        mutate(value = mean(value)) %>%
        dplyr::select(Var2, value, Clt, value1) %>%
        unique() %>%
        as.data.frame()
    
    if (use.sum) {
        temp_cell_score$value = temp_cell_score$value1
    }
    
    temp_cell_score <- dcast(
        temp_cell_score, Clt~Var2, 
        fun.aggregate = mean, value.var = "value",
        fill = 0
    )
    
    temp_cell_score <- as.data.frame(temp_cell_score)
    temp_cell_score <- temp_cell_score[!is.na(temp_cell_score$Clt), ]
    temp_cell_score <- temp_cell_score[, colnames(temp_cell_score) != "Clt"]
    # temp_cell_score_scale = scale(temp_cell_score)

    temp_cell_score
}


make_heatmap_seurat <- function(
    obj, cell_score, 
    samplen=200, seed=42, 
    merged = T, column_title=NULL, 
    ident = NULL) {
    temp_meta = obj@meta.data[order(obj@meta.data$Clt), ]
    temp_meta$Phase = as.character(meta[as.character(temp_meta$Cells), "Phase"])
    temp_meta$Phase[temp_meta$Phase == "G1"] = "Non-Cycling"
    
    if (merged) {
        temp_meta$Phase[temp_meta$Phase %in% c("G2M", "S")] = "Cycling"
        temp_meta$Phase = factor(temp_meta$Phase, levels=c("Non-Cycling", "Cycling"))
    } else {
        temp_meta$Phase = factor(temp_meta$Phase, levels=c("Non-Cycling", "S", "G2M"))
    }
    
    if (!is.null(ident)) {
        temp_meta = temp_meta[ident, ]
    }
    
    set.seed(seed)
    temp_meta <- temp_meta %>%
        group_by(Clt) %>%
        sample_n(samplen, replace = T) %>%
        unique() %>%
        as.data.frame()
    
    rownames(temp_meta) <- temp_meta$Cells
    temp_meta = temp_meta[order(temp_meta$Clt), ]
    
    temp_meta <- temp_meta[intersect(rownames(temp_meta), colnames(cell_score)), ]
    temp_meta = temp_meta[order(temp_meta$Phase, temp_meta$Clt, temp_meta$Batch), ]
    
    
    cell_color = c(
        wes_palette("Darjeeling1"), 
        wes_palette("Moonrise3"),
        wes_palette("Darjeeling2"),
        wes_palette("Moonrise1")
    )[1:length(unique(temp_meta$Clt))]
    
    names(cell_color) = unique(temp_meta$Clt)
    
    batch_colors = as.character(wes_palette("Moonrise3"))
    names(batch_colors) <- c("WCH", "DL", 3:5)
    
    ta = HeatmapAnnotation(
        Cycle = temp_meta$Phase,
        Cluster = temp_meta$Clt,
        Batch = temp_meta$Batch,
        col = list(
            Cluster=cell_color,
            Cycle=cell_cycle_color,
            Batch=batch_colors
        ),
        show_legend = T,
        show_annotation_name = T,
        annotation_legend_param = list(
            Cluster=list(direction = "horizontal"),
            Cycle=list(direction = "horizontal"),
            Batch=list(direction = "horizontal")
        )
    )
    
    h <- Heatmap(
        as.matrix(cell_score[, rownames(temp_meta)]),
        name = "Score",
        col = col_fun,
        border = T,
        cluster_rows = F,
        cluster_columns = F,
        show_column_names = F,
        column_split = temp_meta$Phase,
        top_annotation = ta,
        column_title = column_title,
        row_title_gp = gpar(fontfamily = "Arial Unicode MS"),
        column_title_gp = gpar(fontfamily = "Arial Unicode MS"),
        heatmap_legend_param = list(
            labels_gp = gpar(fontfamily = "Arial Unicode MS"),
            title_gp = gpar(fontfamily = "Arial Unicode MS"),
            direction = "horizontal"
        ),
        use_raster = T
    )
    
    h
}



calculate_sample_cycle_percentage <- function(obj, meta) {
    temp_meta = obj@meta.data[order(obj@meta.data$Clt), ]
    temp_meta$Phase = as.character(meta[as.character(temp_meta$Cells), "Phase"])
    temp_meta$Phase[temp_meta$Phase == "G1"] = "Non-Cycling"
    temp_meta$Phase[temp_meta$Phase %in% c("G2M", "S")] = "Cycling"
    
    temp_meta %>%
        group_by(Phase, Clt) %>%
        add_tally() %>%
        dplyr::select(Phase, Clt, n) %>% # PatientID, 
        unique() %>%
        group_by(Clt) %>% #PatientID, 
        mutate(p = n / sum(n) * 100) %>%
        as.data.frame()
}
```


### ATII

```{R}
# obj <- readRDS(full.path("03_each_cells/Batch/ATII/LUAD/seurat.rds"))

obj <- readRDS(full.path("11_CNV/each_cells/ATII/LUAD/seurat_obj.rds"))

markers <- read.csv(
    full.path("11_CNV/each_cells/ATII/LUAD/cluster_markers.csv"),   # 03_each_cells/Batch
    row.names = 1, stringsAsFactors = F
)
```



```{R fig.height=4, fig.width=6}
markers <- markers %>%
    filter(p_val_adj < 0.05) %>%
    group_by(ident) %>%
    top_n(50, wt = avg_logFC) %>%
    unique() %>%
    as.data.frame()

temp_cell_score = make_cell_score_seurat(obj, obj@meta.data, markers, scale.data = T)
obj@meta.data$Clt <- as.numeric(obj@meta.data$res.0.03) + 1
obj@meta.data$Cells <- rownames(obj@meta.data)
obj@meta.data$Batch <- ifelse(
    meta[rownames(obj@meta.data), "Batch"] == 3, "DL", "WCH"
)


# temp_sd = apply(temp_cell_score, 2, sd)
# plot(density(temp_sd))
# temp_cell_score1 <- temp_cell_score[, temp_sd > 0.4]

pdf(full.path("14_cycling/ATII_LUAD.pdf"), width = 6, height = 4)
draw(
    make_heatmap_seurat(obj, temp_cell_score, column_title = "ATII"),
    heatmap_legend_side = "bottom", annotation_legend_side = "bottom",
    merge_legend = T
)
dev.off()
```


```{R}
obj@meta.data$Clt <- obj@meta.data$Clt
perc = NULL
temp_perc = calculate_sample_cycle_percentage(obj, meta)
temp_perc$Cell = "ATII"
temp_perc$Disease = "LUAD"
perc = rbind(perc, temp_perc)
```



### Basal
```{R}
# obj <- readRDS(full.path("03_each_cells/Batch/Basal/LUSC/seurat.rds"))
obj <- readRDS(full.path("11_CNV/each_cells/Basal/LUSC/seurat_obj.rds"))

markers <- read.csv(
    full.path("11_CNV/each_cells/Basal/LUSC/cluster_markers.csv"), 
    row.names = 1, 
    stringsAsFactors = F
)
```


```{R}
markers <- markers %>%
    filter(p_val_adj < 0.05) %>%
    group_by(ident) %>%
    top_n(50, wt = avg_logFC) %>%
    unique() %>%
    as.data.frame()

temp_cell_score = make_cell_score_seurat(obj, obj@meta.data, markers, scale.data = T)

obj@meta.data$Clt <- factor(
    as.character(as.numeric(obj@meta.data$res.0.2)  + 1 ),
    levels = as.character(sort(unique(as.numeric(obj@meta.data$res.0.2) + 1)))
)

obj@meta.data$Cells <- rownames(obj@meta.data)
obj@meta.data$Batch <- ifelse(
    meta[rownames(obj@meta.data), "Batch"] == 3, "DL", "WCH"
)

pdf(full.path("14_cycling/Basal_LUSC.pdf"), width = 6, height = 4)
draw(
    make_heatmap_seurat(obj, temp_cell_score, column_title = "Basal"),
    heatmap_legend_side = "bottom", annotation_legend_side = "bottom",
    merge_legend = T
)
dev.off()
```


```{R}
temp_perc = calculate_sample_cycle_percentage(obj, meta)
temp_perc$Cell = "Basal"
temp_perc$Disease = "LUSC"
perc = rbind(perc, temp_perc)
```


### NE
```{R}
# obj <- readRDS(full.path("03_each_cells/Batch/NE/LUAD/seurat.rds"))

obj <- readRDS(full.path("11_CNV/each_cells/NE/LUAD/seurat_obj.rds"))


markers <- read.csv(
    full.path("11_CNV/each_cells/NE/LUAD/cluster_markers.csv"), 
    row.names = 1, stringsAsFactors = F
)
```


```{R}
markers <- markers %>%
    filter(p_val_adj < 0.05) %>%
    group_by(ident) %>%
    top_n(50, wt = avg_logFC) %>%
    unique() %>%
    as.data.frame()

temp_cell_score = make_cell_score_seurat(obj, obj@meta.data, markers, scale.data = T)

obj@meta.data$Clt <- as.character(as.numeric(obj@meta.data$res.0.1) + 1)
# temp_sd = apply(temp_cell_score, 2, sd)
# plot(density(temp_sd))
# temp_cell_score1 <- temp_cell_score[, temp_sd > 0.4]

pdf(full.path("14_cycling/NE_LUAD.pdf"), width = 8, height = 4)
draw(make_heatmap_seurat(obj, temp_cell_score, column_title = "NE (LUAD)"))
dev.off()
```

```{R}
temp_perc = calculate_sample_cycle_percentage(obj, meta)
temp_perc$Cell = "NE"
temp_perc$Disease = "LUAD"
perc = rbind(perc, temp_perc)
```


### Epi
```{R}
# obj <- readRDS(full.path("03_each_cells/Batch/Epi/LUAD/seurat.rds"))

obj <- readRDS(full.path("11_CNV/each_cells/Epi/LUAD/seurat_obj.rds"))

markers <- read.csv(
    full.path("11_CNV/each_cells/Epi/LUAD/cluster_markers.csv"), 
    row.names = 1, stringsAsFactors = F
)
```


```{R}
markers <- markers %>%
    filter(p_val_adj < 0.05) %>%
    group_by(ident) %>%
    top_n(20, wt = avg_logFC) %>%
    unique() %>%
    as.data.frame()

temp_cell_score = make_cell_score_seurat(obj, obj@meta.data, markers, scale.data = T)

obj@meta.data$Clt <- as.character(as.numeric(obj@meta.data$res.0.1) + 1)
# temp_sd = apply(temp_cell_score, 2, sd)
# plot(density(temp_sd))
# temp_cell_score1 <- temp_cell_score[, temp_sd > 0.4]

# pdf(full.path("14_cycling/Epi_LUAD.pdf"), width = 8, height = 4)
draw(make_heatmap_seurat(obj, temp_cell_score, column_title = "Epi (LUAD)"))
# dev.off()
```

```{R}
temp_perc = calculate_sample_cycle_percentage(obj, meta)
temp_perc$Cell = "Epi"
temp_perc$Disease = "LUAD"
perc = rbind(perc, temp_perc)
```

### Fib LUAD
```{R}
obj <- readRDS(full.path("11_CNV/each_cells/Fib/LUAD/seurat_obj.rds"))

markers <- read.csv(
    full.path("11_CNV/each_cells/Fib/LUAD/cluster_markers.csv"), 
    row.names = 1, stringsAsFactors = F
)
```


```{R}
markers <- markers %>%
    filter(p_val_adj < 0.05) %>%
    group_by(ident) %>%
    top_n(50, wt = avg_logFC) %>%
    unique() %>%
    as.data.frame()

temp_cell_score = make_cell_score_seurat(obj, obj@meta.data, markers, scale.data = T)

obj@meta.data$Clt <- as.character(as.numeric(obj@meta.data$res.0.1) + 1)
# temp_sd = apply(temp_cell_score, 2, sd)
# plot(density(temp_sd))
# temp_cell_score1 <- temp_cell_score[, temp_sd > 0.4]

pdf(full.path("14_cycling/Fib_LUAD.pdf"), width = 8, height = 4)
draw(make_heatmap_seurat(obj, temp_cell_score, column_title = "Fib (LUAD)"))
dev.off()
```

```{R}
temp_perc = calculate_sample_cycle_percentage(obj, meta)
temp_perc$Cell = "Fib"
temp_perc$Disease = "LUAD"
perc = rbind(perc, temp_perc)
```


### Fib LUSC
```{R}
obj <- readRDS(full.path("11_CNV/each_cells/Fib/LUSC/seurat_obj.rds"))

markers <- read.csv(
    full.path("11_CNV/each_cells/Fib/LUSC/cluster_markers.csv"), 
    row.names = 1, stringsAsFactors = F
)
```


```{R}
markers <- markers %>%
    filter(p_val_adj < 0.05) %>%
    group_by(ident) %>%
    top_n(50, wt = avg_logFC) %>%
    unique() %>%
    as.data.frame()

temp_cell_score = make_cell_score_seurat(obj, obj@meta.data, markers, scale.data = T)

obj@meta.data$Clt <- as.character(as.numeric(obj@meta.data$res.0.1) + 1)
# temp_sd = apply(temp_cell_score, 2, sd)
# plot(density(temp_sd))
# temp_cell_score1 <- temp_cell_score[, temp_sd > 0.4]

pdf(full.path("14_cycling/Fib_LUSC.pdf"), width = 8, height = 4)
draw(make_heatmap_seurat(obj, temp_cell_score, column_title = "Fib (LUSC)"))
dev.off()
```


```{R}
temp_perc = calculate_sample_cycle_percentage(obj, meta)
temp_perc$Cell = "Fib"
temp_perc$Disease = "LUSC"
perc = rbind(perc, temp_perc)
```


```{R}
library(openxlsx)

perc_mtx_LUAD = dcast(
    perc[perc$Disease == "LUAD", ], 
    PatientID + Phase ~ Cell + Clt, 
    fun.aggregate = mean, value.var = "n",
    fill = 0
)

perc_mtx_LUSC = dcast(
    perc[perc$Disease == "LUSC", ], 
    PatientID + Phase ~ Cell + Clt, 
    fun.aggregate = mean, value.var = "n",
    fill = 0
)


perc_perc_LUAD = dcast(
    perc[perc$Disease == "LUAD", ], 
    PatientID + Phase ~ Cell + Clt, 
    fun.aggregate = mean, value.var = "p",
    fill = 0
)

perc_perc_LUSC = dcast(
    perc[perc$Disease == "LUSC", ], 
    PatientID + Phase ~ Cell + Clt, 
    fun.aggregate = mean, value.var = "p",
    fill = 0
)


wb = createWorkbook()
addWorksheet(wb, "LUAD_num")
writeData(wb, 1, perc_mtx_LUAD)

addWorksheet(wb, "LUAD_perc")
writeData(wb, 2, perc_perc_LUAD)

addWorksheet(wb, "LUSC_num")
writeData(wb, 3, perc_mtx_LUSC)

addWorksheet(wb, "LUSC_perc")
writeData(wb, 4, perc_perc_LUSC)
```

### cycling by disease
```{r}
temp_meta = meta
temp_meta$Phase = as.character(temp_meta$Phase)
temp_meta$Phase[temp_meta$Phase == "G1"] = "Non-Cycling"
temp_meta$Phase[temp_meta$Phase %in% c("G2M", "S")] = "Cycling"

temp_meta = temp_meta %>%
    group_by(PatientID, Phase, Disease, cell_short) %>%
    add_tally() %>%
    dplyr::select(PatientID, Phase, Disease, cell_short, n) %>%
    unique() %>%
    group_by(PatientID, Disease, cell_short) %>%
    mutate(p = n / sum(n) * 100) %>%
    as.data.frame()



perc_mtx <- dcast(
    temp_meta, PatientID + Phase ~ cell_short + Disease, 
    value.var = "n", fun.aggregate = mean, fill = 0
)

perc_perc <- dcast(
    temp_meta, PatientID + Phase ~ cell_short + Disease, 
    value.var = "p", fun.aggregate = mean, fill = 0
)


addWorksheet(wb, "All_cell_num")
writeData(wb, 5, perc_mtx)

addWorksheet(wb, "All_cell_perc")
writeData(wb, 6, perc_perc)

saveWorkbook(wb, file = full.path("14_cycling/percentage_stats.xlsx"), overwrite = T)
```


### Cycling by Stage

```{r fig.height=12, fig.width=12}
temp_meta = meta
temp_meta$Phase = as.character(temp_meta$Phase)
temp_meta$Phase[temp_meta$Phase == "G1"] = "Non-Cycling"
temp_meta$Phase[temp_meta$Phase %in% c("G2M", "S")] = "Cycling"

temp <- temp_meta %>%
    group_by(cell_short, Stage, Phase) %>%
    add_tally() %>%
    dplyr::select(cell_short, Stage, Phase, n) %>%
    unique() %>%
    group_by(cell_short, Stage) %>%
    mutate(p = n / sum(n)) %>%
    as.data.frame()

ggplot(temp, aes(x=Stage, y=n, fill=Phase)) + 
    geom_bar(stat = "identity", position = position_dodge()) +
    facet_wrap(.~cell_short, scales = "free")
```


```{r fig.height=8, fig.width=12}
temp$lab = paste0(round(temp$p, 2), "\n(", temp$n, ")")

ggplot(temp, aes(x=Stage, y=p, fill=Phase, label = lab)) + 
    geom_bar(stat = "identity") +
    facet_wrap(.~cell_short, scales = "free") +
    geom_text(size = 3, position = position_stack(vjust = 0.5))
```


## Cell cycle percentage by cells

```{r}
make_percentage_barplot_by_clt <- function(obj, meta) {
    temp_meta = obj@meta.data
    temp_meta$Phase = as.character(meta[rownames(temp_meta), "Phase"])
    temp_meta$Phase[temp_meta$Phase == "G1"] = "Non-Cycling"
    temp_meta$Phase[temp_meta$Phase %in% c("G2M", "S")] = "Cycling"
    
    temp <- temp_meta %>%
        group_by(Clt, Phase) %>%
        add_tally() %>%
        dplyr::select(Clt, Phase, n) %>%
        unique() %>%
        group_by(Clt) %>%
        mutate(p = n / sum(n)) %>%
        as.data.frame()
    
    temp$lab = paste0(round(temp$p, 2), "\n(", temp$n, ")")
    
    ggplot(temp, aes(x=as.character(Clt), y=p, fill=Phase, label = lab)) + 
        geom_bar(stat = "identity") +
        geom_text(size = 3, position = position_stack(vjust = 0.5))
}
```


### ATII
```{R fig.height=3, fig.width=6}
atii <- readRDS(full.path("03_each_cells/Batch/ATII/LUAD/seurat.rds"))
atii@meta.data$Clt <- as.character(as.numeric(atii@meta.data$res.0.25) + 1)

make_percentage_barplot_by_clt(atii, meta) + labs(x="", y="", title = "ATII")
```

### Basal
```{R fig.height=3, fig.width=6}
basal <- readRDS(full.path("03_each_cells/Batch/Basal/LUSC/seurat.rds"))
basal@meta.data$Clt <- as.character(as.numeric(basal@meta.data$res.0.09) + 1)

make_percentage_barplot_by_clt(basal, meta) + labs(x="", y="", title = "Basal")
```


### NE
```{R fig.height=3, fig.width=6}
ne <- readRDS(full.path("03_each_cells/Batch/NE/LUAD/seurat.rds"))
ne@meta.data$Clt <- as.character(as.numeric(ne@meta.data$res.0.1) + 1)

make_percentage_barplot_by_clt(ne, meta) + labs(x="", y="", title = "NE")
```

### Epi
```{R fig.height=3, fig.width=6}
epi <- readRDS(full.path("03_each_cells/Batch/Epi/LUAD/seurat.rds"))
epi@meta.data$Clt <- as.character(as.numeric(epi@meta.data$res.0.1) + 1)

make_percentage_barplot_by_clt(epi, meta) + labs(x="", y="", title = "Epi")
```


### Fib LUAD
```{R fig.height=3, fig.width=6}
fib_ad <- readRDS(full.path("03_each_cells/Batch/Fib/LUAD/seurat.rds"))
fib_ad@meta.data$Clt <- as.character(as.numeric(fib_ad@meta.data$res.0.1) + 1)

make_percentage_barplot_by_clt(fib_ad, meta) + labs(x="", y="", title = "Fib (LUAD)")
```

### Fib LUSC
```{R fig.height=3, fig.width=6}
fib_sc <- readRDS(full.path("03_each_cells/Batch/Fib/LUSC/seurat.rds"))
fib_sc@meta.data$Clt <- as.character(as.numeric(fib_sc@meta.data$res.0.1) + 1)

make_percentage_barplot_by_clt(fib_sc, meta) + labs(x="", y="", title = "Fib (LUSC)")
```


# Re calcualte after inferCNV


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.