---
title: "Test_tmod"
author: "Zhang Yiming"
date: "6/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tmod)
library(openxlsx)
library(gplots)
library(dplyr)
library(doMC)
library(stringr)
library(fmsb)
```

```{r}
full.path <- function(path) {
    return(paste("/mnt/raid62/Lung_cancer_10x/02_figures_each_cell/Alveolar_II_ADC", path, sep = "/"))
}


# function to read and format the normalized_counts from sctransform
read_sctransform <- function(path="paga/normalized_counts.csv.gz") {
    print(path)
    r = gzfile(path)
    data = read.csv(r, row.names = 1)
    
    colnames(data) = gsub("\\.", "-", colnames(data), perl=F)
    colnames(data) = gsub("^X", "", colnames(data), perl=T)
    return(data)
}

obj <- readRDS(full.path("Alveolar_II.rds"))

expr = read_sctransform(full.path("paga/normalized_counts.csv.gz"))

module <- read.xlsx(full.path("mfuzz_gene_module/results.xlsx"))

diff.exp = read.xlsx(full.path("annotation_results_by_stage.xlsx"))
```

之前的预处理中已经获得基因的p.value和logFC，后续直接使用该数据

## analysis of transcriptomic responses to tuberculosis

> examples

```{r}
library(limma)
data(Egambia)
design <- cbind(Intercept=rep(1, 30), TB=rep(c(0,1), each= 15))
E <- as.matrix(Egambia[,-c(1:3)])
fit <- eBayes(lmFit(E, design))
tt <- topTable(fit, coef=2, number=Inf,
genelist=Egambia[,1:3])

head(tt)
```

```{r}
group <- rep( c("CTRL", "TB"), each=15)
showGene(E["20799",], group,
main=Egambia["20799", "GENE_SYMBOL"])
```

```{r}
l <- tt$GENE_SYMBOL
resC <- tmodCERNOtest(l)
head(resC, 15)
```

以自己的数据进行测试，计划从limma开始算，先计算I与II之间的区别，limma无法直接用在扎。所以必须修饰数据
```{r}
stagesToNum <- c("I"=1, "II"=2, "III"=3, "IV"=4)

design <- obj@meta.data[, "Stage", drop = F]
design$Stage = sapply(design$Stage, function(x){stagesToNum[x]})


E <- as.matrix(obj@raw.data[, rownames(design)])

fit <- eBayes(lmFit(E, design))

# 获取msd的值
x <- tmodLimmaTopTable(fit, coef="Stage")
```

## 检查logFC和qval以及msd的排序
```{r}
x <- x[ x$logFC.Stage > 0, ] # only to simplify the output!
x2 <- x[ order(x$msd.Stage, decreasing=T),]
```

直接输入一行list of genes理论上即可进行计算，但是跟什么比，毕竟只是一个factor，
理论上应该是数据和level的排序，影响结果。
然而如何排序，q值还是logfc？都不靠谱

```{r }
msig <- tmodImportMSigDB("/mnt/raid62/Lung_cancer_10x/00_data_ingest/00_raw_data/msigdb_v6.2.xml")
for (i in unique(module$gene_module_id)) {
    l <- module$gene[module$gene_module_id == i]
    print
    resC <- tmodCERNOtest(l[order(x2[as.character(l), "msd.Stage"], decreasing = T)], mset = msig)

    print(head(resC, 15))
}


```

> Where logFC is the log fold change, CI.L is the left boundary of the 95% confidence
interval of logFC and CI.R is the right boundary

根据MSD进行排序，logFC > 0, CI.L; logFC < 0, -CI.R

```{r}
x <- tmodLimmaTopTable(fit, coef="TB", genelist=Egambia[,1:3])
x.lfc <- x[ order(abs(x$logFC.TB), decreasing=T),]
x.qval <- x[ order(x$qval.TB),]
x.msd <- x[ order(x$msd.TB, decreasing=T),]
comparison <- list(
lfc=tmodCERNOtest(x.lfc$GENE_SYMBOL),
qval=tmodCERNOtest(x.qval$GENE_SYMBOL),
msd=tmodCERNOtest(x.msd$GENE_SYMBOL))
tmodPanelPlot(comparison)
```

```{r}
data = read.table("/mnt/raid62/Lung_cancer_10x/02_figures_each_cell/Alveolar_II_ADC/tmod_stage/stage_3_data.txt", header = F)
```

```{r}
res = tmodCERNOtest(data[data[,1] == 0, 2])msig
```

tmodLimmaDecideTests比较不同条件下，多少基因高表达，多少基因低表达，这个必须依赖limma的fit结果。然而，这个能够拿到我想要的结果么

---
### 测试下不同细胞的tmod富集效果


```{R}
obj <- readRDS("/mnt/raid62/Lung_cancer_10x/00_data_ingest/04_rds_generated/seurat_obj_selected_patients_harmony.rds")


meta <- read.csv("/mnt/raid62/Lung_cancer_10x/00_data_ingest/03_annotation_csv/20190506_meta.csv", row.names = 1)


obj@meta.data = meta
```

```{r}
saveRDS(obj, "/mnt/raid62/Lung_cancer_10x/00_data_ingest/04_rds_generated/20190610_seurat_obj_selected_patients_harmony_formatted_meta.rds")
```

```{r}
# obj <- readRDS("/mnt/raid62/Lung_cancer_10x/00_data_ingest/99_seurat_by_patients/PLC029/seurat_obj.rds")
# markers = find_markers(obj, group.by = "res.0.6", n.cores = 10)
#
# write.csv(markers, "/mnt/raid62/Lung_cancer_10x/00_data_ingest/99_seurat_by_patients/PLC029/markers.csv")

markers = read.csv("/mnt/raid62/Lung_cancer_10x/00_data_ingest/99_seurat_by_patients/PLC029/markers.csv")

temp = markers %>% filter(p_val_adj < 0.05) # %>% group_by(ident) %>% top_n(20, w = avg_logFC)

res_p29 = NULL
for (i in unique(temp$ident)) {
    temp_res = temp[temp$ident == i, ]
    temp_res = temp_res[order(temp_res$avg_logFC, decreasing = T), ]
    temp_res = tmodCERNOtest(
        temp_res$gene,
        mset = msig,
        qval=0.05
    )

    temp_res$ident = i
    res_p29 = rbind(res_p29, temp_res)
}

res_p29 = res_p29[res_p29$N1 > 0, ]
```

```{r}
markers = read.csv("/mnt/raid62/Lung_cancer_10x/00_data_ingest/99_seurat_by_patients/PLC012/markers.csv")

temp = markers %>% filter(p_val_adj < 0.05) # %>% group_by(ident) %>% top_n(50, w = avg_logFC)

res_p12 = NULL
for (i in unique(temp$ident)) {
    temp_res = temp[temp$ident == i, ]
    temp_res = temp_res[order(temp_res$avg_logFC, decreasing = T), ]
    temp_res = tmodCERNOtest(
        temp_res$gene,
        mset = msig,
        qval=0.05
    )

    if (!is.null(temp_res) && nrow(temp_res) > 0) {
        temp_res$ident = i
        res_p12 = rbind(res_p12, temp_res)
    }
}

res_p12 = res_p12[res_p12$N1 > 0, ]
```


### 从个人样本来看，所有基因是能够富集到信息的，现在在所有细胞中测试
- q < 0.05
- logFC > 0

Perform test on all data
```{r}
files = list.files(path = "/mnt/raid62/Lung_cancer_10x/02_figures_each_cell/", pattern = "annotation_results_by_stage.xlsx", recursive = T, full.names = T)

sel <- msig$MODULES$Category %in% c("H", "C5", "C2")

make_tmod_on_xlsx <- function(path, mode = "U") {
    markers = read.xlsx(path, rowNames = T)

    temp = markers # %>% filter(p_val_adj < 0.05) # %>% group_by(ident) %>% top_n(50, w = avg_logFC)

    res = NULL
    for (i in unique(temp$ident)) {
        temp_res = temp[temp$ident == i, ]
        temp_res = temp_res[order(temp_res$avg_logFC, decreasing = T), ]

        if (mode == "U") {
            temp_res = tmodUtest(
                temp_res$gene,
                mset = msig[sel],
                qval=0.05
            )
        } else if (mode == "Z") {
            temp_res = tmodZtest(
                temp_res$gene,
                mset = msig[sel],
                qval=0.05
            )
        } else {
            temp_res = tmodCERNOtest(
                temp_res$gene,
                mset = msig[sel],
                qval=0.05
            )
        }

        if (!is.null(temp_res) && nrow(temp_res) > 0) {
            temp_res$ident = i
            res = rbind(res, temp_res)
        }
    }

    res = res[res$N1 > 0, ]
    return(res)
}

registerDoMC(20)
resZ = foreach(i = files, .combine=rbind) %dopar% {
    print(i)
    temp_res = make_tmod_on_xlsx(i, mode = "Z")

    if (!is.null(temp_res) && nrow(temp_res) > 0) {
        temp_res$cell = basename(dirname(i))
    } else {
        temp_res = NULL
    }

    temp_res
}


resU = foreach(i = files, .combine=rbind) %dopar% {
    print(i)
    temp_res = make_tmod_on_xlsx(i, mode = "U")

    if (!is.null(temp_res) && nrow(temp_res) > 0) {
        temp_res$cell = basename(dirname(i))
    } else {
        temp_res = NULL
    }

    temp_res
}


resCERNO = foreach(i = files, .combine=rbind) %dopar% {
    print(i)
    temp_res = make_tmod_on_xlsx(i, mode = "C")

    if (!is.null(temp_res) && nrow(temp_res) > 0) {
        temp_res$cell = basename(dirname(i))
    } else {
        temp_res = NULL
    }

    temp_res
}



wb = createWorkbook()
addWorksheet(wb, "Utest")
writeData(wb, 1, resU)

addWorksheet(wb, "CERNOtest")
writeData(wb, 2, resCERNO)

addWorksheet(wb, "Ztest")
writeData(wb, 3, resZ)

saveWorkbook(wb, "/mnt/raid62/Lung_cancer_10x/tmod_stage/all_stage_specific_genes_different_tests.xlsx", overwrite=T)
```




Make tagcloud of Utest
```{r}
library(tagcloud)
image_dir = "/mnt/raid62/Lung_cancer_10x/tmod_stage/all_stage_specific_genes/Utest"
dir.create(image_dir, showWarnings = F, recursive = T)

foreach(i = unique(resU$cell)) %dopar% {
    temp_res = resU[resU$cell == i, ]

    out_dir = paste(image_dir, i, sep = "/")
    dir.create(out_dir, showWarnings = F)

    for (j in unique(temp_res$ident)) {
        temp = temp_res[temp_res$ident == j,]

        if (nrow(temp) <=1 ) {
            print(i)
            print(temp$Title)
            next
        }

        png(paste(out_dir, paste0(j, ".png"), sep = "_"), width = 8, height = 8, units = "in", res = 600)
        w <- -log10(temp$P.Value)
        c <- smoothPalette(temp$AUC, min=0.5)
        tags <- strmultline(temp$Title)
        tagcloud(tags, weights=w, col=c)
        mtext(paste(i, j), side=3, outer=TRUE, line=-1)
        dev.off()
    }
}
```

CERNOtest
```{r}
image_dir = "/mnt/raid62/Lung_cancer_10x/tmod_stage/all_stage_specific_genes/CERNOtest"
dir.create(image_dir, showWarnings = F, recursive = T)

foreach(i = unique(resCERNO$cell)) %dopar% {
    temp_res = resCERNO[resCERNO$cell == i, ]

    out_dir = paste(image_dir, i, sep = "/")
    # dir.create(out_dir, showWarnings = F)

    for (j in unique(temp_res$ident)) {
        temp = temp_res[temp_res$ident == j,]

        if (nrow(temp) <=1 ) {
            print(i)
            print(temp$Title)
            next
        }

        png(paste(out_dir, paste0(j, ".png"), sep = "_"), width = 8, height = 8, units = "in", res = 600)
        w <- -log10(temp$P.Value)
        c <- smoothPalette(temp$AUC, min=0.5)
        tags <- strmultline(temp$Title)
        tagcloud(tags, weights=w, col=c)
        mtext(paste(i, j), side=3, outer=TRUE, line=-1)
        dev.off()
    }
}
```


Ztest
```{r}
image_dir = "/mnt/raid62/Lung_cancer_10x/tmod_stage/all_stage_specific_genes/Ztest"
dir.create(image_dir, showWarnings = F, recursive = T)

foreach(i = unique(resZ$cell)) %dopar% {
    temp_res = resZ[resZ$cell == i, ]

    out_dir = paste(image_dir, i, sep = "/")
    # dir.create(out_dir, showWarnings = F)

    for (j in unique(temp_res$ident)) {
        temp = temp_res[temp_res$ident == j,]

        if (nrow(temp) <=1 ) {
            print(i)
            print(temp$Title)
            next
        }

        png(paste(out_dir, paste0(j, ".png"), sep = "_"), width = 8, height = 8, units = "in", res = 600)
        w <- -log10(temp$P.Value)
        c <- smoothPalette(temp$AUC, min=0.5)
        tags <- strmultline(temp$Title)
        tagcloud(tags, weights=w, col=c)
        mtext(paste(i, j), side=3, outer=TRUE, line=-1)
        dev.off()
    }
}
```


Try to make dotplot
```{r fig.height=12, fig.width=12}
image_dir = "/mnt/raid62/Lung_cancer_10x/tmod_stage/all_stage_specific_genes/dotplot"

dir.create(image_dir, showWarnings = F, recursive = T)

resU <- read.xlsx("/mnt/raid62/Lung_cancer_10x/tmod_stage/all_stage_specific_genes_different_tests.xlsx", sheet = 1)

resCERNO <- read.xlsx("/mnt/raid62/Lung_cancer_10x/tmod_stage/all_stage_specific_genes_different_tests.xlsx", sheet = 2)

resZ = read.xlsx("/mnt/raid62/Lung_cancer_10x/tmod_stage/all_stage_specific_genes_different_tests.xlsx", sheet = 3)

make_dotplot <- function(resC, resU, resZ, cell, label=NULL, pval.thr = 0.01) {
    tempC = resC[resC$cell == cell, ]
    tempU = resU[resU$cell == cell, ]
    tempZ = resZ[resZ$cell == cell, ]

    if(!is.null(label)) {
        tempC = tempC[str_detect(tempC$Title, label), ]
        tempU = tempU[str_detect(tempU$Title, label), ]
        tempZ = tempZ[str_detect(tempZ$Title, label), ]
    }

    stages = unique(c(tempC$ident, tempU$ident, tempZ$ident))

    resAll = list()
    for(i in sort(stages)) {
        resAll[[paste("CERNO", cell, i, sep = " - ")]] = tempC[tempC$ident == i, ]
        resAll[[paste("Utest", cell, i, sep = " - ")]] = tempU[tempU$ident == i, ]
        resAll[[paste("Ztest", cell, i, sep = " - ")]] = tempZ[tempZ$ident == i, ]
    }

    # tempC$Title = paste(tempC$Title, tempC$cell, tempC$ident, sep = " - ")
    # tempU$Title = paste(tempU$Title, tempU$cell, tempU$ident, sep = " - ")
    # tempZ$Title = paste(tempZ$Title, tempZ$cell, tempZ$ident, sep = " - ")
    # resAll <- list(paste(CERNO, cell, )=tempC, U=tempU, Z=tempZ)

   return(resAll)
}

cells = unique(c(resCERNO$cell, resU$cell, resZ$cell))
registerDoMC(20)
foreach (cell = cells) %dopar% {

# for (cell in intersect(intersect(resCERNO$cell, resU$cell), resZ$cell)) {
    print(cell)

    tryCatch({
        resAll = make_dotplot(resC = resCERNO, resU = resU, resZ = resZ, cell = cell, label = "^Go", pval.thr = 0.01)
        temp_summ = tmodSummary(resAll)

        height = nrow(temp_summ) / 4
        if (height < 8) {
            height = 8
        }

        png(
            paste(image_dir, paste0(cell, "_Go.png"), sep = "/"),
            width = ncol(temp_summ) + max(sapply(temp_summ$Title, str_length)) / 20,
            height = height,
            units = "in",
            res = 300
        )
        plot.new()
        tmodPanelPlot(resAll, pval.thr = 0.01, pval.thr.lower = -Inf )
        dev.off()
    }, error = function(e){})


    tryCatch({
        resAll = make_dotplot(resC = resCERNO, resU = resU, resZ = resZ, cell = cell, label = "^[^(Go)]", pval.thr = 0.05)
        temp_summ = tmodSummary(resAll)

        height = nrow(temp_summ) / 4
        if (height < 8) {
            height = 8
        }

        png(
            paste(image_dir, paste0(cell, "_Other.png"), sep = "/"),
            width = ncol(temp_summ) + max(sapply(temp_summ$Title, str_length)) / 20,
            height = height,
            units = "in",
            res = 300
        )
        plot.new()
        tmodPanelPlot(resAll, pval.thr = 0.01, pval.thr.lower = -Inf )
        dev.off()
    }, error = function(e){})
}
```

### get gene module from tmod tests, using the overlap of tmod and mfuzz
```{r}
temp = resU[resU$adj.P.Val < 0.05 & resU$AUC > 0.5, ]

msig <- tmodImportMSigDB("/mnt/raid62/Lung_cancer_10x/00_data_ingest/00_raw_data/msigdb_v6.2.xml")

head(temp)
```

```{r}

msig_dataframe = NULL
for (i in names(msig$MODULES2GENES)) {
    temp = msig[i]$GENES
    temp$MODULE = i

    msig_dataframe = rbind(msig_dataframe, temp)
}

data.frame(test="1", test1 = "2")
```


```{R}
markers <- read.xlsx("/mnt/raid62/Lung_cancer_10x/02_figures_each_cell/Alveolar_II_ADC/annotation_results_by_stage.xlsx", rowNames = T)

temp_markers = markers[markers$ident == "I", "gene"]

length(intersect(temp_markers, msig["M11585"]$GENES$ID))
```

---

### make radar plots
```{R}
library(openxlsx)
library(stringr)
library(fmsb)
library(Seurat)
library(reshape2)
library("wesanderson")

set.seed(1)

module = read.xlsx("/mnt/raid62/Lung_cancer_10x/tmod_stage/tmod_mfuzz_overlap_AUC_0.5_merged.xlsx", sheet = 2)

obj = readRDS("/mnt/raid62/Lung_cancer_10x/02_figures_each_cell/Alveolar_II/Alveolar_II.rds")
```

```{r}

format_radar_matrix <- function(object, meta) {
    mat = MinMax(object@scale.data[meta$genes, , drop = FALSE], -2.5, 2.5)
    mat = melt(as.matrix(mat))
    colnames(mat) <- c("genes", "cell", "value")
    
    expr <- merge(meta, mat, by = "genes")
    expr$PatientID = object@meta.data[expr$cell, "PatientID"]
    
    # print(head(expr))
    
    data = as.data.frame(
        expr %>% 
            dplyr::select(stage, PatientID, value) %>% 
            dplyr::group_by(stage, PatientID) %>%
            mutate(val=mean(value)) %>%
            dplyr::select(stage, PatientID, val) %>%
            unique()
        )
    
    data = dcast(data, stage~PatientID, value.var = "val")
    rownames(data) <- data$stage
    data <- data[, colnames(data) != "stage"]
    
    return(as.data.frame(t(data)))
}


make_radar_plots <- function(object, meta, group.by = "Stage", output_prefix=NULL) {
    
    current_stages = sapply(meta$stage, function(x) {
        return(str_split(x, "\\.")[[1]][2])
    })
    current_stages = intersect(object@meta.data[, group.by], current_stages)
    
    for(i in current_stages) {
        temp_meta = object@meta.data[object@meta.data[, group.by] == i, ]
        obj <- CreateSeuratObject(
            object@raw.data[, rownames(temp_meta), drop = FALSE],
            meta = temp_meta
        )
        
        obj@scale.data = object@scale.data[, rownames(temp_meta), drop = FALSE]
        
        expr = format_radar_matrix(obj, meta)    
        expr <- rbind(
            rep(ceiling(max(expr)), ncol(expr)),
            rep(floor(min(expr)), ncol(expr)),
            expr
        )
    
        # print(expr)
        
        if (!is.null(output_prefix)) {
            png(paste(output_prefix, "_", i, ".png", sep = ""), width = 12, height = 6, res = 600, units = "in")
        }
        
        par(
            mfrow=c(1, 2), 
            bty='n', 
            oma = c(0.5,0.5,0,0) + 0.1,
            mar = c(1,0,0,1) + 0.1
        )
    
        legend_labels = rownames(expr)
        legend_labels = legend_labels[3: length(legend_labels)]
        colors_border = wes_palette("Zissou1", length(legend_labels), type = "continuous")
    
        radarchart(expr, axistype=1, 
            #custom polygon
            pcol=colors_border, # pfcol=colors_in, plwd=4 , plty=1, 
            #custom the grid
            cglcol="grey", cglty=1, axislabcol="grey", 
            caxislabels=seq(expr[2, 1], expr[1, 1], (expr[1, 1] - expr[2, 1]) / 5), cglwd=0.8,
            #custom labels
            vlcex=0.8
        )
        plot(0, col="white", cex=0, axes=F, ann=FALSE)
    
        legend(
            "left",
            legend = legend_labels, 
            bty = "n", pch=20 , 
            col=colors_border , 
            text.col = "grey", 
            cex=1.2, 
            pt.cex=3,
            ncol = 2
        )
        
        if (!is.null(output_prefix)) {
            dev.off()
        }
        
        break
    }
}




for(i in unique(module$Cell_name)) {
    print(i)
    
    temp_module = module[module$Cell_name == i, ]
    
    # format meta
    meta = NULL
    for(j in 1:nrow(temp_module)) {
        stage = paste("M", temp_module[j, "Stage"], sep = ".")
        genes = str_split(temp_module[j, "Genes"], "\\|")[[1]]
        
        meta = rbind(meta, data.frame(stage=stage, genes=genes))
    }
    
    # two random select groups
    for(i in 1:2) {
        meta = rbind(meta, data.frame(
            stage = paste("R", i, sep = ""), 
            genes = sample(
                rownames(obj@raw.data)[!rownames(obj@raw.data) %in% meta$genes],
                min(100, sum(!rownames(obj@raw.data) %in% meta$genes))
            )
        ))
    
    }
    
    # format and plot
    make_radar_plots(object = obj, meta = meta, group.by = "Stage", output_prefix = NULL)

    # print(meta)
    break
}
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
